### 2022-06-03 README.Rmd
Updated readme file a little

```{r global_options, include = FALSE}
library(knitr)
options(width = 120)
opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figs/',
               include = TRUE, warning = FALSE, message = FALSE)
```

# Bremla
This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including all the data/code to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K., Rypdal, M. & Boers, N. (2022). Comprehensive uncertainty estimation of the timing of Greenland warmings of the Greenland Ice core records. Climate of the Past, x(x),x-x. doi:xxxxx

and

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Tie-point paper (title TBD)

## installation

The simplest way to install the package is to install the remotes package and run
```{r github-install, eval=FALSE}
#install.packages("remotes")
remotes::install_github("eirikmn/bremla")
```

To get started, see the documentation '?bremla' and the associated example.

The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [iceandclimate.nbi.ku.dk](Centre for Ice and Climate) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [https://www.sciencedirect.com/science/article/pii/S0277379114003485](Rasmussen et al. (2014)). Also includes the tie-point presented in [https://cp.copernicus.org/articles/14/1755/2018/](Adolphi et al. (2018)) and [https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D](Muscheler et al. (2020)). Other data and tie-points should work as well.

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if synchronized chronologies is also computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Example

```{r example}
This is a real data example which produces 5000 simulations from a synchronized time scale.
require(INLA)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
d18O = NGRIP_5cm$d18O
proxy=d18O
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2

events=list(locations = events_rasmussen$depth,
             locations_unit="depth",degree=1)
control.transition_dating=list(age.ref=age.reference,age.label="years (yb2k)")
object = bremla(formula,data,reference.label="GICC05",
                  nsims=5000,
                  events=events,
                  synchronization=list(method="adolphi"),
                  control.fit=list(method="inla"),
                  control.sim=list(synchronized=TRUE),
                  print.progress=TRUE )
summary(object)
plot(object)
```

## Attribution

This code is associated and written for the papers Myrvoll-Nilsen et al. (2022) and Myrvoll-Nilsen et al. (202x) mentioned above. Feel free to use the code, but please cite the accompanying papers.

## License

The code in this repository is made available under the terms of the MIT License. For details, see LICENSE file.

### 2022-06-09 README.Rmd
Small changes + LICENSE change
Small changes in code and changed license from MIT to GPL2 since the package depend on INLA which is GPL2

@@
 # Bremla
-This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including all the data/code to reproduce the results for
+This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for
 
@@
 ## Example
-
-```{r example}
 This is a real data example which produces 5000 simulations from a synchronized time scale.
+```{r example, message=FALSE, warning=FALSE, include=FALSE}
 require(INLA)
@@
 depth = NGRIP_5cm$depth
-d18O = NGRIP_5cm$d18O
-proxy=d18O
+proxy = NGRIP_5cm$d18O
+
 data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)

### 2022-06-09 R/plot.bremla.R

@@ -82,7 +82,7 @@ plot.bremla = function(x,
                        plot.event_age = list(age.reference=NULL,xrev=TRUE,label=NULL),
                        postscript=FALSE,
                        pdf=FALSE,
-                       prefix = "bemla.plots/figure-",
+                       prefix = "bremla.plots/figure-",
                        ...){
   if(!postscript && !pdf){
     dev=getOption("device")

### 2022-07-26 README.Rmd
Changed license and some other things
Also added Github action: R CMD check

@@
-```{r global_options, include = FALSE}
-library(knitr)
-options(width = 120)
-opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figs/',
-               include = TRUE, warning = FALSE, message = FALSE)
+---
+output: github_document
+---
+
+<!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.path = "man/figures/README-",
+  dev = "png",
+  out.width = "80%"
+)
 ```
 
-# Bremla
+# bremla
+
+<!-- badges: start -->
+[![R-CMD-check](https://github.com/eirikmn/bremla/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/eirikmn/bremla/actions/workflows/R-CMD-check.yaml)
+<!-- badges: end -->
+
 This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for
@@
 Comprehensive uncertainty estimation of the timing of Greenland warmings of the Greenland Ice core records.
-Climate of the Past, x(x),x-x.
-doi:xxxxx
+Climate of the Past, 18, 1275-1294.
+doi.org/10.5194/cp-18-1275-2022
 
@@
 The simplest way to install the package is to install the remotes package and run
-```{r github-install, eval=FALSE}
-#install.packages("remotes")
-remotes::install_github("eirikmn/bremla")
+```{r install, eval=FALSE}
+#install.packages("devtools")
+devtools::install_github("eirikmn/bremla")
 ```
@@
 
-The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [iceandclimate.nbi.ku.dk](Centre for Ice and Climate) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [https://www.sciencedirect.com/science/article/pii/S0277379114003485](Rasmussen et al.
-(2014)).
-Also includes the tie-point presented in [https://cp.copernicus.org/articles/14/1755/2018/](Adolphi et al.
-(2018)) and [https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D](Muscheler et al.
-(2020)).
+The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al.
+(2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485).
+Also includes the tie-point presented in [Adolphi et al.
+(2018)](https://cp.copernicus.org/articles/14/1755/2018/) and [Muscheler et al.
+(2020)](https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D).
 Other data and tie-points should work as well.
@@
 require(INLA)
+library(bremla)
 data("event_intervals")
@@
             locations_unit="depth",degree=1)
-control.transition_dating=list(age.ref=age.reference,age.label="years (yb2k)")
+
 object = bremla(formula,data,reference.label="GICC05",
@@
 summary(object)
-plot(object)
 ```
 
+```{r plot, message=FALSE, eval=TRUE, echo=FALSE}
+require(ggplot2)
+free_indexes = x$tie_points$free_indexes
+tie_indexes = x$tie_points$tie_indexes
+ageref_free = x$data$age[free_indexes]
+ageref_tie = x$data$age[tie_indexes]
+
+fullpd = data.frame(depth = x$data$depth[free_indexes],
+                medians=x$simulation$summary_sync$median[free_indexes]-ageref_free,
+                lower=x$simulation$summary_sync$lower[free_indexes]-ageref_free,
+                upper=x$simulation$summary_sync$upper[free_indexes]-ageref_free)
+
+gg2 = ggplot(data=fullpd,aes(x=.data$depth)) +
+geom_line(aes(y=0),color="blue",linetype="dotted",size=0.2)+
+theme_bw()+ylab("Estimated age - reference (years)")+
+xlab("NGRIP depth (m)")
+
+gg2 = gg2+geom_line(aes(y=.data$medians))+
+geom_ribbon(aes(ymin=.data$lower,ymax=.data$upper),color="red",fill="red",alpha=0.3)
+
+gg2 = gg2 + geom_segment(data=data.frame(depth=x$data$depth[tie_indexes],                                          median=x$simulation$summary_sync$median[tie_indexes]-ageref_tie,                               lower=x$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
+                    upper=x$simulation$summary_sync$upper[tie_indexes]-ageref_tie),
+                    aes(x=.data$depth,y=.data$lower,xend=.data$depth,yend=.data$upper),
+                    col="magenta")
+print(gg2)
+```   
 ## Attribution
@@
 
-The code in this repository is made available under the terms of the MIT License. For details, see LICENSE file.
+The code in this repository is made available under the terms of the GNU (version >=2) License. For details, see LICENSE.md file.

### 2022-07-26 R/bremla.R

@@ -34,6 +34,7 @@
 #' \donttest{
 #' ## Simulation example
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2
@@ -186,7 +187,8 @@ bremla <- function(formula,data,reference.label=NULL,
   }
 
 
-  if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
+  if(nsims>0 && control.sim$synchronized %in% c(TRUE,2)){
+    control.sim$nsims=nsims
     #produce samples from the chronologies
     object = bremla_synchronized_simulation(object, control.sim=control.sim,
                                             print.progress=print.progress)

### 2022-07-26 R/bremla_biased_chronologies.R

@@ -90,7 +90,7 @@ bremla_biased_chronologies = function(object,control.bias,print.progress=FALSE){
   n = nrow(object$simulation$age)
   biasparams = control.bias$biasparams
 
-  if(class(biasparams) == "numeric"){
+  if(is.null(dim(biasparams))){
     biasparams = matrix(biasparams,ncol=1)
   }
   m = ncol(biasparams)

### 2022-07-26 R/bremla_chronology_simulation.R

@@ -72,6 +72,7 @@ bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE
   #if(!is.null(control.sim))
    control.sim = set.options(control.sim,control.sim.default())
 
+   if(control.sim$synchronized==TRUE) control.sim$synchronized=2
   object$.args$control.sim = control.sim
 
   ## sample hyperparameters
@@ -186,6 +187,7 @@ bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE
     }
   }
 
+  object$.args$sim = list(unsync=TRUE)
   time.endage = Sys.time()
   if(print.progress) cat("Completed in ",difftime(time.endage,time.startage,units="secs")[[1]]," seconds!\n",sep="")
   object$time$simulation$mean = difftime(time.endmean,time.startmean,"secs")[[1]]

### 2022-07-26 R/bremla_modelfitter.R

@@ -75,6 +75,7 @@ bremla_modelfitter = function(object, control.fit,
 
   fit = lm(object$.internal$formula.ls,object$data) #fits model using least squares
 
+
   if(sum(is.na(fit$coefficients))){
     stop(paste0("least squares yields NA estimates for ",sum(is.na(fit$coefficients)),
                 " coefficients. Make sure 'formula' does not include linearly dependent variables.

### 2022-07-26 R/bremla_synchronized_simulation.R

@@ -58,7 +58,7 @@
 #' @export
 bremla_synchronized_simulation = function(object,control.sim,print.progress=FALSE){
 
-
+time.start = Sys.time()
   if(missing(control.sim)){
 
     if(!is.null(object$.args$control.sim)){
@@ -74,6 +74,7 @@ bremla_synchronized_simulation = function(object,control.sim,print.progress=FALS
   control.sim = set.options(control.sim,control.sim.default())
 
   object$.args$control.sim = control.sim
+  if(control.sim$synchronized==FALSE) control.sim$synchronized=2
 
   ## sample hyperparameters
   if(is.null(object$fitting)) stop("Fitting results not found. Run 'bremla_modelfitter' first.")
@@ -139,6 +140,7 @@ bremla_synchronized_simulation = function(object,control.sim,print.progress=FALS
     if(print.progress) cat(" completed!\n",sep="")
   }
 
+  if(print.progress) cat("Sampling fixed coefficients...",sep="")
   latentselection = object$.internal$lat.selection
   # latentselection = list()
   # reg.model = object$.args$reg.model
@@ -151,15 +153,19 @@ bremla_synchronized_simulation = function(object,control.sim,print.progress=FALS
   #   if(reg.model$psi0) latentselection[[paste0("a",i-1)]] = 1
   #   if(reg.model$psi1)latentselection[[paste0("c",i-1)]] = 1
   # }
+  timecoef.start = Sys.time()
   ncores_postsamp = max(1,object$.args$control.sim$ncores)
   latentsamples = inla.posterior.sample(nsims,object$fitting$inla$fit,
                                         selection=latentselection,verbose=FALSE,
                                         add.names=FALSE,num.threads = ncores_postsamp)
 
+  timecoef.end = Sys.time()
+  if(print.progress) cat(" completed in ",difftime(timecoef.end,timecoef.start,units="secs")[[1]]," seconds.\n",sep="")
   samples = matrix(NA,nrow=n,ncol=nsims)
 
 
-  time.start = Sys.time()
+  if(print.progress) cat("Simulating synchronized chronologies...\n",sep="")
+  timeage.start = Sys.time()
   for(r in 1:nsims){
     sigma_sample = object$simulation$sigma[r]
     phi_sample = object$simulation$phi[r]
@@ -199,11 +205,11 @@ bremla_synchronized_simulation = function(object,control.sim,print.progress=FALS
 
[…]

### 2022-07-26 R/events_depth_to_age.R

@@ -15,6 +15,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2
@@ -37,7 +38,7 @@
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
 #' synchronization=list(method="gauss")
-#' control.sim=list(synchronized=2,
+#' control.sim=list(synchronized=TRUE,
 #'                  summary=list(compute=TRUE))
 #'
 #' #simulate transition:

### 2022-07-26 R/linrampfitter.R

@@ -17,6 +17,7 @@
 #' @examples
 #'\donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2
@@ -39,7 +40,7 @@
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
 #' synchronization=list(method="gauss")
-#' control.sim=list(synchronized=2,
+#' control.sim=list(synchronized=TRUE,
 #'                  summary=list(compute=TRUE))
 #'
 #' #simulate transition:

### 2022-07-26 R/plot.bremla.R

@@ -69,8 +69,8 @@
 #' @importFrom rlang .data
 #' @importFrom ggplot2 ggplot aes geom_line geom_ribbon theme_bw xlab ylab geom_segment geom_point
 plot.bremla = function(x,
-                       plot.ls = list(fitted=TRUE,legend=NULL,residuals=TRUE,histogram=TRUE,qqplot=TRUE,acf=TRUE,xrev=FALSE,
-                                      label.fit=NULL,label.res=NULL,label.hist=NULL,label.qq=NULL,label.acf=NULL),
+                       plot.ls = list(fitted=TRUE,legend=NULL,residuals=TRUE,histogram=TRUE,qqplot=FALSE,acf=TRUE,xrev=FALSE,
+                                      label.fit=NULL,label.res=NULL,label.hist=NULL,label.qq=NULL,label.acf=NULL,merge=TRUE),
                        plot.inla.posterior = list(posteriors=TRUE,label=NULL),
                        plot.inlasims = list(plotsims=0,legend=NULL,xrev=FALSE,label=NULL),
                        plot.syncsims = list(plotsims=0,legend=NULL,xrev=FALSE,label=NULL),
@@ -115,8 +115,16 @@ plot.bremla = function(x,
 
   if(!is.null(plot.ls) && !is.null(x$fitting$LS$fit)){
     figure.count <- new.plot(postscript,pdf,prefix,figure.count,...) +1
-    par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
+    if(plot.ls$merge){
+      par(mfrow=c(2,2),mar=c(5,4,4,2)+0.1)
+    }else{
+      par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
+    }
+    plotnum=0
+
+
     if(plot.ls$fitted){
+      plotnum=plotnum+1
       xlim = range(z)
       if(plot.ls$xrev) xlim=rev(xlim)
       if(is.null(plot.ls$label.fit)){
@@ -133,6 +141,7 @@ plot.bremla = function(x,
     if(!is.null(plot.ls$legend)) legend(x=leg$x,y=leg$y,legend=leg$legend,col=leg$col,lty=leg$lty,cex=leg$cex,pch=leg$pch,lwd=leg$lwd,pt.cex=leg$pt.cex,bty=leg$bty)
 
     if(plot.ls$residuals){
+      plotnum=plotnum+1
       if(is.null(plot.ls$label.res)){
[…]

### 2022-07-26 R/summary.bremla.R

@@ -71,7 +71,7 @@ summary.bremla = function(object,
 
   if(!is.null(object$time$simulation$total)){
     cpu=as.numeric(c(cpu,round(object$time$simulation$total,digits=digits)))
-    cpu.navn=c(cpu.navn,"Chronology sampling")
+    cpu.navn=c(cpu.navn,"Chron. sampling")
   }
   if(!is.null(object$linramp) && !is.null(object$event_dating)){
     cpu=as.numeric(c(cpu,round(object$time$t1_and_ramp + object$time$event_age$total,digits=digits)))
@@ -124,11 +124,20 @@ summary.bremla = function(object,
       colnames(hypers) = c("mean","sd","quant0.025","quant0.25","quant0.5","quant0.75","quant0.975")
       rownames(hypers) = c("sigma_epsilon","phi1","phi2")
     }
-    maxlength=2048L
-    formulastring = format(object$.args$formula.ls)
-    if(sum(nchar(formulastring)) > maxlength){
-      formulastring = paste0( substr(deparse(object$.args$formula.ls),1L,maxlength),"...")
+
+    formulastring = format(object$.args$formula.input)
+
+    if(!is.null(object$.args$events)){
+      if(object$.args$events$fill_formula){
+        formulastring = paste0(formulastring," + psi_fill(degree=",object$.args$events$degree,
+                               ", n_events=",object$.args$events$nevents,")")
+      }
     }
+
+    # formulastring = format(object$.args$formula.ls)
+    # if(sum(nchar(formulastring)) > maxlength){
+    #   formulastring = paste0( substr(deparse(object$.args$formula.ls),1L,maxlength),"...")
+    # }
     fit.arg = list(formula = formulastring,noise=object$.args$control.fit$noise,
                    nevents=object$.args$events$nevents,
                    method = object$.args$control.fit$method)
@@ -153,7 +162,7 @@ summary.bremla = function(object,
                      method=object$tie_points$method,
                      nsims=object$tie_points$nsims,
                      locations=object$tie_points$locations,
-                     locations_unit=object$tie_points$unit)
+                     locations_unit=object$tie_points$locations_unit)
     ut = c(ut,list(tiepoints=tiepoints))
   }
 
[…]

### 2022-07-26 README.Rmd
Adjustments to readme and dependencies in DESCRIPTION

@@
 This is a real data example which produces 5000 simulations from a synchronized time scale.
-```{r example, message=FALSE, warning=FALSE, include=FALSE}
+```{r example, message=FALSE, warning=FALSE, include=TRUE}
 require(INLA)
@@
 ```
+
 ## Attribution

### 2022-07-26 README.Rmd
Added seeds on examples

@@
 
-object = bremla(formula,data,reference.label="GICC05",
+results = bremla(formula,data,reference.label="GICC05",
                 nsims=5000,
@@
                 control.fit=list(method="inla"),
-                control.sim=list(synchronized=TRUE),
-                print.progress=TRUE )
-summary(object)
+                control.sim=list(synchronized=TRUE) )
+summary(results)
 ```
@@
 require(ggplot2)
-free_indexes = x$tie_points$free_indexes
-tie_indexes = x$tie_points$tie_indexes
-ageref_free = x$data$age[free_indexes]
-ageref_tie = x$data$age[tie_indexes]
+free_indexes = results$tie_points$free_indexes
+tie_indexes = results$tie_points$tie_indexes
+ageref_free = results$data$age[free_indexes]
+ageref_tie = results$data$age[tie_indexes]
 
-fullpd = data.frame(depth = x$data$depth[free_indexes],
-                medians=x$simulation$summary_sync$median[free_indexes]-ageref_free,
-                lower=x$simulation$summary_sync$lower[free_indexes]-ageref_free,
-                upper=x$simulation$summary_sync$upper[free_indexes]-ageref_free)
+fullpd = data.frame(depth = results$data$depth[free_indexes],
+                medians=results$simulation$summary_sync$median[free_indexes]-ageref_free,
+                lower=results$simulation$summary_sync$lower[free_indexes]-ageref_free,
+                upper=results$simulation$summary_sync$upper[free_indexes]-ageref_free)
 
@@
 
-gg2 = gg2 + geom_segment(data=data.frame(depth=x$data$depth[tie_indexes],                                          median=x$simulation$summary_sync$median[tie_indexes]-ageref_tie,                               lower=x$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
-                    upper=x$simulation$summary_sync$upper[tie_indexes]-ageref_tie),
+gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=x$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
+                   upper=results$simulation$summary_sync$upper[tie_indexes]-ageref_tie),
                    aes(x=.data$depth,y=.data$lower,xend=.data$depth,yend=.data$upper),

### 2022-07-26 R/bremla_chronology_simulation.R

@@ -17,6 +17,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/bremla_modelfitter.R

@@ -16,6 +16,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/bremla_prepare.R

@@ -38,6 +38,7 @@
 #'
 #' @examples
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/bremla_synchronized_simulation.R

@@ -15,6 +15,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/helpfulfunctions.R

@@ -124,6 +124,7 @@ which.index = function(events, record){ ## Finds which indices of 'record' that 
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2
@@ -459,6 +460,7 @@ adolphi_tiepoint_simmer = function(nsims=10000,tieshifts = numeric(5), plotdens=
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/plot.bremla.R

@@ -26,6 +26,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/print.bremla.R

@@ -13,6 +13,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-07-26 R/rgeneric.uneven.AR1.R

@@ -15,6 +15,7 @@
 #'
 #' @examples
 #' \donttest{
+#' set.seed(1)
 #' n=500
 #' timepoints = 1:n/n #should be scaled so that timepoints is between 0 and 1
 #'                    #(you can always transform back later, but this helps

### 2022-07-26 R/summary.bremla.R

@@ -13,6 +13,7 @@
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2
@@ -268,6 +269,7 @@ return(ut)
 #' @examples
 #' \donttest{
 #' require(stats)
+#' set.seed(1)
 #' n <- 1000
 #' phi <- 0.8
 #' sigma <- 1.2

### 2022-08-23 inst/Examples/Paper_results/Paper_results_1.Rmd
Added some more features and Rmd files to generate the plots and results from both papers

---
output: github_document
---

<!-- Paper_results.md is generated from paper_results.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/results-",
  dev = "png",
  out.width = "80%"
)
```

# bremla

This file includes the data and code needed to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K., Rypdal, M. & Boers, N. (2022). Comprehensive uncertainty estimation of the timing of Greenland warmings of the Greenland Ice core records. Climate of the Past, 18, 1275-1294. doi.org/10.5194/cp-18-1275-2022

# installation

The simplest way to install the package is to install the devtools package and run
```{r install, eval=FALSE}
#install.packages("devtools")
devtools::install_github("eirikmn/bremla")
```

# Myrvoll-Nilsen et al. (2022)

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r bremla, eval=TRUE}
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
n=length(diff(age))

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
     xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
formula = dy~-1+depth2+proxy

nsims=3000

results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=FALSE) ,
                print.progress=TRUE
                )

```

The results from the linear fit can be extracted by running:

```{r ls, eval=TRUE}
fit = results$fitting$LS$fit

{
  par(mfrow=c(2,1),mar=c(5,4,4,2)+0.1)
plot(results$data$depth,results$data$dy,type="l",xlab="Depth (m)",ylab="Layers per 5cm",main="(a) Least squares fit",xlim=rev(range(results$data$depth))) + lines(results$data$depth,fit$fitted.values,col="red")   
abline(v=events$locations,col="gray",lwd=0.8)
legend(1770,6.35,legend=c("Layer increments","Regression fit","Climate transitions"),
       col=c("black","red","gray"),lty=c(1,1,NA),cex=0.9,pch=c(NA,NA,"|"))
plot(results$data$depth,fit$residuals,type="l",xlab="Depth (m)",ylab="Residual errors per 5cm",main="(b) Least squares residuals",xlim=rev(range(results$data$depth)))
}
{
layout(mat=matrix(c(1,3,2,3),nrow=2))
  par(mar=c(5,3.5,4,2)+0.1)
hist(fit$residuals,xlab="Residual errors per 5cm",ylab="Density",main="(a) Histogram", col="orange")
{qqnorm(fit$residuals,main="(b) Q-Q Plot")
qqline(fit$residuals)}
acf(fit$residuals, lag.max = 20,main="(c) Autocorrelation function")
}
par(mfrow=c(1,1))
```

To compare the final age-depth uncertainties across the iid, AR(1) and AR(2) models run:
```{r iid_ar2, eval=TRUE}
resultsiid = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="iid"),
                control.sim=list(synchronized=FALSE) 
                )

resultsar2 = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar2"),
                control.sim=list(synchronized=FALSE) 
                )

gicc05 = results$original.chron$age[1+1:n]
{
plot(depth[1+1:n],resultsar2$simulation$summary$upper-gicc05,type="l",col="red", xlab="Depth (m)", ylab="Simulated time scale - GICC05 (years)",ylim=c(-260,260),xlim=rev(range(depth)), lwd=2)
lines(depth[1+1:n],results$simulation$summary$upper-gicc05,col="blue",lwd=2)
lines(depth[1+1:n],resultsiid$simulation$summary$upper-gicc05,col="black",lwd=2)
lines(depth[1+1:n],results$simulation$summary$mean-gicc05,col="gray",lwd=2)
abline(h=0,lty=3)
lines(depth[1+1:n],results$simulation$summary$lower-gicc05,col="blue",lwd=2)
lines(depth[1+1:n],resultsiid$simulation$summary$lower-gicc05,col="black",lwd=2)
lines(depth[1+1:n],resultsar2$simulation$summary$lower-gicc05,col="red",lwd=2)
legend(1640,275,legend=c("Mean","iid CI","AR(1) CI", "AR(2) CI"),
       col=c("gray","black","blue","red"),lty=c(1,1,1,1),cex=0.75)
}
```

To generate the plot to compare different parameters for additional stochastic bias use the following code.

```{r biased, eval=TRUE}

results_biased = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=FALSE),
                control.bias=list(bias.model="uniform",nsims=nsims,
                                  biasparams = matrix(c(0.98,1.02,0.96,1.04),nrow=2)
                                  ),
                print.progress=TRUE
                )

{
  plot(depth[1+1:n],results_biased$biases$bias1$quant0.975-gicc05,type="l",col="blue",lty=2, xlab="depth", ylab="Simulated time scale - GICC05 (years)",xlim = rev(range(depth)), ylim = c(-2700,2700))
  lines(depth[1+1:n], results_biased$biases$bias2$quant0.975-gicc05,col="blue",lty=3)
  lines(depth[1+1:n], results$simulation$summary$upper-gicc05,col="blue",lty=1)
  abline(h=0,lty=3,col="gray")
  lines(depth[1+1:n], results$simulation$summary$lower-gicc05,col="blue",lty=1)
  lines(depth[1+1:n], results_biased$biases$bias1$quant0.025-gicc05,col="blue",lty=2)
  lines(depth[1+1:n], results_biased$biases$bias2$quant0.025-gicc05,col="blue",lty=3)
  lines(depth[1+1:n], NGRIP_5cm$MCE[1+1:n],col="black",lwd=2)
  legend(1660,2700,legend=c("MCE","Unbiased","2% bias rate", "4% bias rate"),
       col=c("black","blue","blue","blue"),lty=c(1,1,2,3),lwd=c(2,1,1,1),cex=0.65)
}

```

## Abrupt warming transitions

We also import table XXXX from Myrvoll-Nilsen et al. (2022) which gives suitable data windows for each abrupt warming transition that is analysed. This was found by varying the start and end point over a 2-dimensional grid and picking those where the noise in the linear ramp model yielded the lowest amplitude.

```{r events_individual, eval=TRUE}

eventnumber=13 #number between 1 and 29. specifies which transition to consider

#load data window and specifics to transition
lowerints = which.index(event_intervals$depth_int_lower.m, depth[2:length(depth)])
upperints = which.index(event_intervals$depth_int_upper.m, depth[2:length(depth)])
depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
age.reference = event_intervals$GICC_age.yb2k[eventnumber]
interval = lowerints[eventnumber]:upperints[eventnumber]

transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                      str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])

control.linramp=list(proxy=proxy,
                     interval=interval,
                     interval.unit="index",
                     depth.reference=event_intervals$NGRIP_depth_m[eventnumber]
                     )

results_event = linrampfitter(results,control.linramp)

control.transition_dating = list(label="Simulated transition")
results_event = events_depth_to_age(results_event,
                                    control.transition_dating=control.transition_dating)
{
  layout(mat=matrix(c(1,3,2,3),nrow=2))

  plot(depth[interval],proxy[interval],type="l",col="gray",
     main=paste0("(a) ",transitionlabel," linear ramp fit"),
     xlab="Depth (m)", ylab=paste(expression(delta^18,"O (permil)")),
     xlim = rev(range(depth[interval]))
     )
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$mean,lwd=1.5)
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.025,lwd=1.5,col="red")
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.975,lwd=1.5,col="red")
abline(v=depth.reference,lty=3,lwd=0.7,col="black")

ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt0=results_event$linramp$param$t0$marg.t0
normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt1 = results_event$linramp$param$t1$marginal
normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)

plot(results_event$linramp$param$t0$marg.t0,type="l",xlab="Depth (m)",ylab="Density",
     main="(b) Marginal posterior of onset depth", 
     xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
abline(v=depth.reference,lty=3)
abline(v=results_event$linramp$param$t0$mean,lwd=1.5)
abline(v=results_event$linramp$param$t0$q0.025,col="gray")
abline(v=results_event$linramp$param$t0$q0.975,col="gray")

hist(results_event$event_dating$samples,main="(c) Histogram of onset age",
     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40)
abline(v=age.reference,lty=3)
abline(v=results_event$event_dating$mean,lwd=2)

par(mar=c(5,3.5,4,2)+0.1)
}

```

We want to compare the dating onsets with those given by Rasmussen et al. (2014), Capron () and Buizert (). We start by importing those values

```{r events_import, eval=TRUE}
library(readxl)
rasmussen_depths = event_intervals$NGRIP_depth_m
rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`

buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
                            col_types=c("text","numeric","numeric","numeric","numeric"))
buizert_depths = buizert_onsets$Depth; buizert_ages = buizert_onsets$Age
#buizert_depths[c(2:6,8:9,12:13,15:17,19:21)]
buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]

capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                           sheet="d18O",n_max=25)
capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`

capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                                sheet="Ca2+",n_max=24)
capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`

capind = numeric(29); buiind = numeric(29)

capind = c(NA,2,3, 4,5,6,NA,NA,7,8,NA,9,10,11,NA,NA,NA,NA,NA,12,13,14,NA,NA,15,NA,NA,16,17)
buiind = c(NA,2,NA,3,4,6,NA,NA,8,9,NA,11,12,13,NA,NA,NA,NA,NA,15,16,17,NA,NA,19,NA,NA,20,21)
```

We then generate a model fit using log-Ca as proxy

```{r events_imputedust, eval=TRUE}
dust = NGRIP_5cm$logdust
library(zoo)
dust_filled = na.approx(dust) #Impute missing data using linear interpolation
logdust = log(dust_filled)

data2=data
data2$proxy = logdust
results_dust = bremla(formula,data2,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=FALSE) 
                )
```

We then estimate the dating uncertainty for all 29 events

```{r events_all, evals=TRUE}
agesimmatrix_dust = matrix(NA,29,nsims)
agesimmatrix_d18O = matrix(NA,29,nsims)
depthlist_dust = c()
depthlist_d18O = c()
do.plot.rampfit = FALSE
do.plot.agehist = FALSE
depthstats = as.data.frame(matrix(NA,7,29))
colnames(depthstats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")
agestats = as.data.frame(matrix(NA,7,29))
colnames(agestats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")

# below is some fiddling with optimization parameters to improve the linear ramp fit for 
# some of the more challenging fits

steplengths = rep(0.005,29) 
steplengths[c(5,10)]=0.0001 #sometimes changing the steplength in the INLA optimization procedure can improve convergence (default is 0.005)
#compute.dust = rep(TRUE,29); compute.dust[c(8,27)]=FALSE #
#dtparams1 = rep(1,29); dtparams2 = rep(0.2,29) #prior-parameters for transition length
#t0params1 = rep(1,29); t0params2 = rep(0.2,29) #prior-parameters for transition onset
#dtparams1[10] = 20; dtparams2[10] = 0.1 #
#t0params1[10] = 120; t0params2[10] = 20

for(eventnumber in 1:29){
  if(eventnumber == 10){ #one of the transitions struggled to find good initial value 
    opt.params = c(40,NA,NA,NA) #for onset depth
    priorparams=NULL
  }else{
    opt.params = NULL
    priorparams=NULL
  }

  lowerints = which.index(event_intervals$depth_int_lower.m, results$data$depth)
  upperints = which.index(event_intervals$depth_int_upper.m, results$data$depth)
  interval_d18O = (lowerints[eventnumber]):upperints[eventnumber]+1

  lowerints = which.index(event_intervals$depth_int_lower.m, results_dust$data$depth)
  upperints = which.index(event_intervals$depth_int_upper.m, results_dust$data$depth)
  interval_dust = lowerints[eventnumber]:upperints[eventnumber]+1

  depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
  age.reference = event_intervals$GICC_age.yb2k[eventnumber]

  control.linramp = list(proxy=proxy,interval=interval_d18O,interval.unit="index",
                         depth.ref=depth.reference, imp.fit=imp.fits[eventnumber],
                         h=steplengths[eventnumber],
                         opt.params=opt.params,
                         priorparams=priorparams,
                         silent=2L,
                         label=paste0(eventnumber," (d18O):", event_intervals[eventnumber,2])
                         )

  control.linramp_dust = list(proxy=logdust,interval=interval_dust,interval.unit="index",
                         depth.ref=depth.reference, imp.fit=imp.fits[eventnumber],
                         h=steplengths[eventnumber],
                         opt.params=opt.params,
                         priorparams=priorparams,
                         silent=2L,
                         label=paste0(eventnumber," (logCa):", event_intervals[eventnumber,2])
                         )

  temp_d18O = linrampfitter(results,control.linramp)
  if(do.plot.rampfit){
    {
      plot(depth[interval_d18O],proxy[interval_d18O],type="l",col="gray",
          xlim=rev(range(depth[interval_d18O])), xlab="Depth (m)",ylab=expression(paste(delta^18,"O (permil)")),
          main=paste0(eventnumber," (d18O): ", event_intervals[eventnumber,2]))
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$mean)
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.025,col="red")
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.975,col="red")
      abline(v=rasmussen_depths[eventnumber],lty=3)

      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt0=temp_d18O$linramp$param$t0$marg.t0
      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
      lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt1 = temp_d18O$linramp$param$t1$marginal
      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
      lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
    }
  }
  temp_dust = linrampfitter(results_dust,control.linramp_dust)
  if(do.plot.rampfit){
    {
      plot(temp_dust$linramp$data$x,temp_dust$linramp$data$y,type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
          main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
      # plot(depth[interval_d18O],logdust[interval_d18O],type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
      #     main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$mean)
      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.025,col="red")
      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.975,col="red")
      abline(v=rasmussen_depths[eventnumber],lty=3)

      ybottom = min(temp_dust$linramp$data$y)
      ytop = min(temp_dust$linramp$linrampfit$q0.025)
      margt0=temp_dust$linramp$param$t0$marg.t0
      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
      lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
      ybottom = min(temp_dust$linramp$data$y)
      ytop = min(temp_dust$linramp$linrampfit$q0.025)
      margt1 = temp_dust$linramp$param$t1$marginal
      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
      lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
    }
  }

  control.transition_dating=list(label="d18O")
  agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)

  agesimmatrix_d18O[eventnumber,] = agetemp_d18O$event_dating$samples 
  if(do.plot.agehist){
    hist(agesimmatrix_d18O[eventnumber,],freq=0,breaks=50,col="orange",main=paste0("d18O: ",eventnumber),xlab="Age (yb2k)")
    abline(v=age.reference,col="blue")
  }
  control.transition_dating_dust=list(label="dust")
  agetemp_dust = events_depth_to_age(temp_d18O,control.transition_dating_dust)
  agesimmatrix_dust[eventnumber,] = agetemp_dust$event_dating$samples
   if(do.plot.agehist){
    hist(agesimmatrix_dust[eventnumber,],freq=0,breaks=50,col="orange",main=paste0("dust: ",eventnumber),xlab="Age (yb2k)")
     abline(v=age.reference,col="blue")
  }

  #compute summary statistics
  depthstats[1,eventnumber] = event_intervals$NGRIP_depth_m[eventnumber]
  depthstats[2,eventnumber] = agetemp_d18O$linramp$param$t0$mean
  depthstats[3,eventnumber] = agetemp_d18O$linramp$param$t0$q0.025
  depthstats[4,eventnumber] = agetemp_d18O$linramp$param$t0$q0.975
  depthstats[5,eventnumber] = agetemp_dust$linramp$param$t0$mean
  depthstats[6,eventnumber] = agetemp_dust$linramp$param$t0$q0.025
  depthstats[7,eventnumber] = agetemp_dust$linramp$param$t0$q0.975

  densage1 = density(agesimmatrix_d18O[eventnumber,]); densage1 = data.frame(x=densage1$x,y=densage1$y)
  densage2 = density(agesimmatrix_dust[eventnumber,]); densage2 = data.frame(x=densage2$x,y=densage2$y)
  zage1 = inla.zmarginal(densage1,silent=TRUE)
  zage2 = inla.zmarginal(densage2,silent=TRUE)

  agestats[1,eventnumber] = event_intervals$GICC_age.yb2k[eventnumber]
  agestats[2,eventnumber] = mean(agesimmatrix_d18O[eventnumber])
  agestats[3,eventnumber] = zage1$quant0.025
  agestats[4,eventnumber] = zage1$quant0.975
  agestats[5,eventnumber] = mean(agesimmatrix_dust[eventnumber])
  agestats[6,eventnumber] = zage1$quant0.025
  agestats[7,eventnumber] = zage1$quant0.975

}

```

## Results

To plot all posteriors with the comparison onsets run the code below.

```{r events_plot, eval=TRUE}
library(stringr)
par(mfrow=c(5,6),mar=c(2,2,1.5,1))
for(i in 1:29){
  dens1 = density(agesimmatrix_dust[i,]); dens1 = data.frame(x=dens1$x,y=dens1$y)#/max(dens1$y))
  dens2 = density(agesimmatrix_d18O[i,]); dens2 = data.frame(x=dens2$x,y=dens2$y)#/max(dens2$y))
  xlim = range(dens1$x,dens2$x);ylim = range(dens1$y,dens2$y)
  #par(mar=c(2,2,1.5,1))

  {
    plot(dens1,type="l",col=1,xlab="Onset year (b2k)",ylab="Density",
       main=str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]),
       xlim=xlim,ylim=ylim)
  #Axis(side=1,labels=FALSE,at=dens1$x,tck=0.1)

  lines(dens2,type="l",col="gray")

  if(FALSE){#95% credible intervals
    abline(v=mean(agesimmatrix_dust[i,]),col=1)
    zdens1 = inla.zmarginal(dens1,silent = TRUE)
    abline(v=c(zdens1$quant0.025,zdens1$quant0.975),col=1,lty=3,lwd=0.8)
    abline(v=mean(agesimmatrix_d18O[i,]),col=2)
    zdens2 = inla.zmarginal(dens2,silent = TRUE)
    abline(v=c(zdens2$quant0.025,zdens2$quant0.975),col=2,lty=3,lwd=0.8)
  }

  abline(v=event_intervals$GICC_age.yb2k[i],col="blue")  
  if(!is.na(buiind[i])){
    abline(v=buizert_onsets$Age[buiind[i]],col="green")
    #abline(v=buizert_ages[buiind[i]],col="green")
  }

  # abline(v=capron_ages_NGRIP_d18O,col="orange")
  # abline(v=capron_ages_NGRIP_dust,col="orange",lty=3)
  if(!is.na(capind[i])){
    abline(v=capron_ages_NGRIP_d18O[capind[i]],col="red")
    abline(v=capron_ages_NGRIP_dust[capind[i]],col="pink",lty=1)
  }
  }

}
par(mar=c(0,0,0,0));plot(-1,axes=FALSE,xlab="",ylab="",xlim=c(0,1),ylim=c(0,1))
legend(0,1,legend=c("d18O onset posterior","Ca2+ onset posterior",
                    "Rasmussen onset","Buizert onset",
                    "Capron d18O onset","Capron Ca2+ onset"),
       col=c("black","gray", "blue", "green", "red", "pink"),
       lty=c(1,1,1,1,1,1), cex=0.65,bty="n")

par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
rownames=c()
for(i in 1:29){
  rownames=c(rownames,str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]))
}

```

To format the summary statistics of the onset depth and age into a data.frame object, run the code below.

```{r events_table, eval=TRUE}

colnames = c("d18O_mean","d18O_q0.025","d18O_q0.975",
             "Ca_mean","Ca_q0.025","Ca_q0.975")
depthresults = t(depthstats)[,2:7]
colnames(depthresults)=colnames
rownames(depthresults)=rownames

ageresults = t(agestats[2:7,])
colnames(ageresults)=colnames
rownames(ageresults)=rownames

```

## Attribution

This code is associated and written for the papers Myrvoll-Nilsen et al. (2022) and Myrvoll-Nilsen et al. (202x) mentioned above. Feel free to use the code, but please cite the accompanying papers.

## License

The code in this repository is made available under the terms of the GNU (version >=2) License. For details, see LICENSE.md file.

### 2022-08-23 R/bremla_modelfitter.R

@@ -147,7 +147,7 @@ bremla_modelfitter = function(object, control.fit,
       phi.ls = object$fitting$LS$params$phi
       initialmodes = c(initialmodes, log( (1+phi)/(1-phi) ))
 
-    }else if(tolower(object$.args$noise) %in% c(2,"ar2","ar(2)")){
+    }else if(tolower(object$.args$control.fit$noise) %in% c(2,"ar2","ar(2)")){
       phi1.ls = object$fitting$LS$params$phi1
       phi2.ls = object$fitting$LS$params$phi2
 

### 2022-08-23 R/bremla_synchronized_simulation.R

@@ -205,7 +205,7 @@ time.start = Sys.time()
     #mu_amidb = (free_mu - solve(Qa)%*%Qab%*%(skewsamples[r]-tie_mu))[,1]
 
     #samples[free_indexes,r] = Qsimmer(1,Qa,mu_amidb)
-    if((r %% 1000) == 0){
+    if(print.progress && (r %% 1000 == 0)){
       cat("Synchronous age simulation ",r,"/",nsims,". Elapsed time: ",difftime(Sys.time(),timeage.start,units="secs")[[1]]," seconds...","\n",sep="")
     }
   }

### 2022-08-23 R/helpfulfunctions.R

@@ -86,6 +86,8 @@ linramprev = function(t,t0=0,dt=1,y0=0,dy=1){
 #'
 #' @param events Vector describing the values of interest (here: either depth or age of events)
 #' @param record Vector we want to search (here: either full depth or age record)
+#' @param suppress Boolean. Set to \code{TRUE} to suppress warnings that events fall outside designated
+#' interval.
 #'
 #' @return Returns a vector of indices for which elements of \code{record} best matches the elements of \code{events}-
 #' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
@@ -93,11 +95,14 @@ linramprev = function(t,t0=0,dt=1,y0=0,dy=1){
 #' @keywords indices
 #'
 #' @export
-which.index = function(events, record){ ## Finds which indices of 'record' that are located closest to the 'event's
+which.index = function(events, record,suppress=TRUE){ ## Finds which indices of 'record' that are located closest to the 'event's
   eventindexes = numeric(length(events))
   for(i in 1:length(events)){
     if(events[i] < min(record) || events[i] > max(record)){ #Gives NA if located outside range of 'record'
-      warning(paste0("Event ",i,", located at ",events[i]," is outside the interval covered by 'record' (",min(record),", ",max(record),"). The event will be omitted!"))
+      if(!suppress){
+        warning(paste0("Event ",i,", located at ",events[i]," is outside the interval covered by 'record' (",min(record),", ",max(record),"). The event will be omitted!\n"))
+      }
+
       eventindexes[i] = NA
     }else{
       eventindexes[i] = which(abs(events[i]-record) == min(abs(events[i]-record)))
@@ -197,11 +202,12 @@ bremla_simulationsummarizer = function(object,sync=TRUE,print.progress=FALSE){
         lower[i] = samples[i,1]
         upper[i] = samples[i,1]
       }else{
-        dens = density(samples[i,])
+        dens = density(samples[i,]); dens = data.frame(x=dens$x,y=dens$y)
 
+        hpds = suppressWarnings( inla.hpdmarginal(0.95,inla.smarginal(dens)) )
 
-        lower[i] = inla.hpdmarginal(0.95,dens)[1]
-        upper[i] = inla.hpdmarginal(0.95,dens)[2]
+        lower[i] = hpds[1]
+        upper[i] = hpds[2]
       }
     }
   }else{
@@ -518,6 +524,14 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
[…]

### 2022-08-23 R/linrampfitter.R

@@ -93,6 +93,10 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(is.null(object$simulation$age) && is.null(object$simulation$age_sync))
     stop("Chronologies not found. Run 'bremla_synchronized_simulation' or 'bremla_chronology_simulation' first!")
   interval = control.linramp$interval
+  if(is.null(interval)){
+    interval = 1:length(control.linramp$proxy)
+    control.linramp$interval.unit = "index"
+  }
 
 
 
@@ -106,18 +110,19 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
     interval = range(interval)[1]:range(interval)[2]
   }
 
+  n=length(interval)
   if(print.progress) cat("Initializing linear ramp fit using INLA.\n",sep="")
   df_event = data.frame(xx=rev(object$data$depth[interval]),
                         yy=rev(control.linramp$proxy[interval])) #reverse: Want depth axis representing old->new
 
-  ## default initial values for optim function
-  if(is.null(control.linramp$opt.params)){
-    optparams = c(round(length(interval)/2),round(length(interval)/10),df_event$yy[1],df_event$yy[length(interval)]-df_event$yy[1])
-  }else{
-    optparams = control.linramp$opt.params
-  }
 
-  n=length(interval)
+  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
+  ypointsmin = min(df_event$yy)
+  ypointsmax = max(df_event$yy-ypointsmin)
+  #ypointsmax=1; ypointsmin=0
+  ypoints =(df_event$yy-ypointsmin)/ypointsmax
+
+
   df0=data.frame(y=df_event$yy,x=df_event$xx)
 
   t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
@@ -127,54 +132,101 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(print.progress) cat("Using 'optim' to find initial positions for hyperparameters in INLA.\n",sep="")
   ## for stability the x-axis is transformed to values ranging from 1:n.
 
-  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
+
   df0$time = timepoints
+  df0$proxy = ypoints
 
[…]

### 2022-09-29 inst/Examples/Paper_results/Paper_results_2.Rmd
Minor changes

---
output: github_document
---

<!-- Paper_results_2.md is generated from paper_results_2.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/results2-",
  dev = "png",
  out.width = "80%"
)
```

# bremla

This file includes the data and code needed to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Tie-point paper (title TBD)

# Myrvoll-Nilsen et al. (202x)

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r initialize, eval=FALSE}
#library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
n=length(diff(age))

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
     xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
formula = dy~-1+depth2+proxy

```

We also use tie-points described by probability distributions obtained from Adolphi et al. (2018) and Muscheler et al. (2020). These can be imported as follows.

```{r adolphi, eval=TRUE}

adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = TRUE,
                            x.ref=c(11050,12050,13050,22050,42050))

```

The model can be fitted using the \texttt{bremla} function, which estimates the posterior marginal distributions of the hyperparameters and also performs simulations of chronologies.

```{r hyperpars, eval=FALSE}
nsims=10000

results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=FALSE) ,
                print.progress=TRUE
                )

{

  par(mfrow=c(1,2))
  plot(results$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(paste(sigma[epsilon])),ylab="Density")

  abline(v=results$fitting$inla$hyperparameters$results$sigma_epsilon$mean)
  abline(v=c(results$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
             results$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),
             col="red")

  plot(results$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density")
  abline(v=results$fitting$inla$hyperparameters$results$phi$mean)
  abline(v=c(results$fitting$inla$hyperparameters$results$phi$quant0.025,
             results$fitting$inla$hyperparameters$results$phi$quant0.975),
             col="red")
 par(mfrow=c(1,1)) 
}

```

The fitted model to the layer increments and the simulated chronologies (as compared to the GICC05 time scale) can be plotted as follows.

```{r unconstrained}
{
  library(ggplot2)

  dsims = colDiffs(rbind(matrix(rep(results$initial$age,ncol(results$simulation$age)),nrow=1),results$simulation$age))
  dmeans = rowMeans(dsims)
  dsd = rowSds(dsims)
ggd = data.frame(y = results$data$dy,depth=results$data$depth,
                   mean=dmeans,
                   lower = dmeans-1.96*dsd,
                   upper = dmeans+1.96*dsd)
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+ #xlim(rev(range(results$data$depth)))+
    geom_line(aes(y=y),col="gray")+
    geom_line(aes(y=mean),col="blue",size=0.3)+
    geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.05)+
    xlab("Depth (m)")+ylab("Layers per 5cm")
 print(ggp)

 #ggsave("fitdiff-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}

{
  gicc05=results$original.chron$age[1+1:n]
  ggd = data.frame(depth=results$data$depth,
                   mean=results$simulation$summary$mean-gicc05,
                   lower=results$simulation$summary$lower-gicc05,
                   upper=results$simulation$summary$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")
  print(ggp)

  #ggsave("unsync-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}
```

## Fixed tie-points

In the paper we demonstrate our model first using the Adolphi tie-points fixed to be equal to the mode of the probability distribution. We omit the first tie-point since it falls outside of our considered interval. The constrained chronologies can be computed as follows.

```{r fixed, eval=TRUE}

fixedsims = cbind(#rep(inla.mmarginal(adolphipdfs$tie1),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie2),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie3),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie4),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie5),nsims)
                  )
synchronization = list(samples = fixedsims,nsims=nsims,
                       locations=c(#1492.50,
                                   1502.80,1532.70,1767.25,2132.40),
                       locations_unit="depth")

results_fixed = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=synchronization,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=TRUE) ,
                print.progress=TRUE
                )

{
  gicc05=results_fixed$original.chron$age[1+1:n]
  ggd = data.frame(depth=results_fixed$data$depth,
                   mean=results_fixed$simulation$summary_sync$mean-gicc05,
                   lower=results_fixed$simulation$summary_sync$lower-gicc05,
                   upper=results_fixed$simulation$summary_sync$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")
  print(ggp)

  #ggsave("fixed-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}

```

Synchronizing our chronologies with the Adolphi tie-points where the uncertainties is incorporated cam be done using Monte Carlo simulations. All of this can be done using the bremla function.

```{r sync, eval = TRUE}
library(ggplot2)
synchronization = list(method="adolphi",nsims=nsims,
                       locations=c(#1492.50,
                                   1502.80,1532.70,1767.25,2132.40),
                       locations_unit="depth")

results_sync = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=synchronization,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=TRUE,summary=list(CI.type="hpd")) ,
                print.progress=TRUE
                )

tiepointranges = as.data.frame(matrix(NA,nrow=5,ncol=3))

gicc05=results_sync$original.chron$age[1+1:n]
gicc05tie = gicc05[results_sync$tie_points$locations_indexes]

colnames(tiepointranges)=c("lower","upper","depth")
for(i in 2:5){
  tiepointranges[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",i)]])-gicc05tie[i]
}
tiepointranges[,3] = depth[results_sync$tie_points$locations_indexes]

{

  ggd = data.frame(depth=results_sync$data$depth,
                   mean=results_sync$simulation$summary_sync$mean-gicc05,
                   lower=results_sync$simulation$summary_sync$lower-gicc05,
                   upper=results_sync$simulation$summary_sync$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
    geom_linerange(data=tiepointranges,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
  print(ggp)

  #ggsave("sync-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}

```

# Applications

## Abrupt warmings

The onset of abrupt warmings is dated in two stages that are combined to express the complete dating uncertainty, pertaining to the uncertainty originating from both the age-depth relationship as well as estimating the depth in the record which represent the onset of the transition. The latter is estimated using a linear ramp model as described in Erhardt et al. (2019). The linear ramp fit can be performed and plotted as follows.

```{r events_individual, eval=FALSE}

eventnumber=13 #number between 1 and 29. specifies which transition to consider

#load data window and specifics to transition
lowerints = which.index(event_intervals$depth_int_lower.m, depth[2:length(depth)])
upperints = which.index(event_intervals$depth_int_upper.m, depth[2:length(depth)])
depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
age.reference = event_intervals$GICC_age.yb2k[eventnumber]
interval = lowerints[eventnumber]:upperints[eventnumber]

transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                      str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])

control.linramp=list(proxy=proxy,
                     interval=interval,
                     interval.unit="index",
                     depth.reference=event_intervals$NGRIP_depth_m[eventnumber]
                     )

results_event = linrampfitter(results_sync,control.linramp)

control.transition_dating = list(label="Simulated transition")
results_event = events_depth_to_age(results_event,
                                    control.transition_dating=control.transition_dating)
{
  layout(mat=matrix(c(1,2,1,3),nrow=2))
  par(mar=c(5,4,4,2)+0.1)

  plot(depth[interval],proxy[interval],type="l",col="gray",
     main=paste0("(a) ",transitionlabel," linear ramp fit"),
     xlab="Depth (m)", ylab=expression(paste(delta^18,"O (permil)")),
     xlim = rev(range(depth[interval]))
     )
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$mean,lwd=1.5)
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.025,lwd=1.5,col="red")
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.975,lwd=1.5,col="red")
abline(v=depth.reference,lty=3,lwd=0.7,col="black")

ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt0=results_event$linramp$param$t0$marg.t0
normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt1 = results_event$linramp$param$t1$marginal
normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)

plot(results_event$linramp$param$t0$marg.t0,type="l",xlab=expression(paste("Onset depth (m)")),ylab="Density",
     main="(b) Marginal posterior of onset depth")#, 
     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
#abline(v=depth.reference,lty=3)
abline(v=results_event$linramp$param$t0$mean,lwd=1.5)
abline(v=results_event$linramp$param$t0$q0.025,col="red")
abline(v=results_event$linramp$param$t0$q0.975,col="red")

plot(results_event$linramp$param$t1$marginal,type="l",xlab=expression(paste("End point depth (m)")),ylab="Density",
     main="(c) Marginal posterior of end point depth")#, 
     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
#abline(v=depth.reference,lty=3)
abline(v=results_event$linramp$param$t1$mean,lwd=1.5)
abline(v=results_event$linramp$param$t1$q0.025,col="red")
abline(v=results_event$linramp$param$t1$q0.975,col="red")

par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)

hist(results_event$event_dating$samples,main=paste0(transitionlabel," Onset age"),
     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40, freq=0)
abline(v=age.reference,lty=3)
abline(v=results_event$event_dating$mean,lwd=2)
abline(v=results_event$event_dating$q0.025,lwd=2,col="red")
abline(v=results_event$event_dating$q0.975,lwd=2,col="red")

}

```

We want to compare the dating onsets with those given by Rasmussen et al. (2014), Capron (2021) and Buizert et al. (2015). We start by importing those values

```{r events_import, eval=FALSE}
library(readxl)
rasmussen_depths = event_intervals$NGRIP_depth_m
rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`

buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
                            col_types=c("text","numeric","numeric","numeric","numeric"))
buizert_depths = buizert_onsets$Depth; buizert_ages = buizert_onsets$Age
#buizert_depths[c(2:6,8:9,12:13,15:17,19:21)]
buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]

capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                           sheet="d18O",n_max=25)
capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`

capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                                sheet="Ca2+",n_max=24)
capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`

capind = numeric(29); buiind = numeric(29)

capind = c(NA,2,3, 4,5,6,NA,NA,7,8,NA,9,10,11,NA,NA,NA,NA,NA,12,13,14,NA,NA,15,NA,NA,16,17)
buiind = c(NA,2,NA,3,4,6,NA,NA,8,9,NA,11,12,13,NA,NA,NA,NA,NA,15,16,17,NA,NA,19,NA,NA,20,21)
```

We then estimate the dating uncertainty for all 29 events
```{r}
agesimmatrix_d18O = matrix(NA,29,nsims)
depthlist_d18O = c()
do.plot.rampfit = TRUE
do.plot.agehist = FALSE
depthstats = as.data.frame(matrix(NA,4,29))
colnames(depthstats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")
agestats = as.data.frame(matrix(NA,4,29))
colnames(agestats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")

steplengths = rep(0.01,29)
steplengths[21]=0.001 #sometimes changing the steplength slightly might improve convergence (default is 0.01)

for(eventnumber in 1:29){
  transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                      str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])
  lowerints = which.index(event_intervals$depth_int_lower.m, results_sync$data$depth)
  upperints = which.index(event_intervals$depth_int_upper.m, results_sync$data$depth)
  interval = lowerints[eventnumber]:upperints[eventnumber]

  depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
  age.reference = event_intervals$GICC_age.yb2k[eventnumber]

  control.linramp = list(proxy=proxy,interval=interval,interval.unit="index",
                         depth.ref=depth.reference, 
                         label=paste0(eventnumber," (d18O):", event_intervals[eventnumber,2])
                         )

  temp_d18O = linrampfitter(results_sync,control.linramp)
  if(do.plot.rampfit){
    {plot(depth[interval],proxy[interval],type="l",col="gray",
          xlim=rev(range(depth[interval])), xlab="Depth (m)",ylab=expression(paste(delta^18,"O (permil)")),
          main=paste0(eventnumber,": ", transitionlabel))
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$mean)
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.025,col="red")
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.975,col="red")
      abline(v=rasmussen_depths[eventnumber],lty=3)

      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt0=temp_d18O$linramp$param$t0$marg.t0
      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
      lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt1 = temp_d18O$linramp$param$t1$marginal
      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
      lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
    }
  }

  control.transition_dating=list(label="d18O")
  agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)

  agesimmatrix_d18O[eventnumber,] = agetemp_d18O$event_dating$samples 
  if(do.plot.agehist){
    hist(agesimmatrix_d18O[eventnumber,],freq=0,breaks=50,col="orange",main=paste0("Onset age: ",transitionlabel),xlab="Age (yb2k)")
    abline(v=age.reference,col="blue")
  }

  #Store summary statistics
  depthstats[1,eventnumber] = event_intervals$NGRIP_depth_m[eventnumber]
  depthstats[2,eventnumber] = agetemp_d18O$linramp$param$t0$mean
  depthstats[3,eventnumber] = agetemp_d18O$linramp$param$t0$q0.025
  depthstats[4,eventnumber] = agetemp_d18O$linramp$param$t0$q0.975

  densage = density(agesimmatrix_d18O[eventnumber,]); densage = data.frame(x=densage$x,y=densage$y)

  zage = inla.zmarginal(densage,silent=TRUE)

  agestats[1,eventnumber] = event_intervals$GICC_age.yb2k[eventnumber]
  agestats[2,eventnumber] = mean(agesimmatrix_d18O[eventnumber])
  agestats[3,eventnumber] = zage$quant0.025
  agestats[4,eventnumber] = zage$quant0.975

}

```

```{r}

{
par(mfrow=c(5,6),mar=c(2,2,1.5,1))
for(i in 1:29){
  dens = density(agesimmatrix_d18O[i,])
  dens = data.frame(x=dens$x,y=dens$y)#/max(dens2$y))
  xlim = range(dens$x,event_intervals$GICC_age.yb2k[i]);ylim = range(dens$y)

  if(!is.na(buiind[i])) xlim=range(xlim,buizert_onsets$Age[buiind[i]])
  if(!is.na(capind[i])) xlim=range(xlim,capron_ages_NGRIP_d18O[capind[i]])
  #par(mar=c(2,2,1.5,1))

  plot(dens,type="l",col=1,xlab="Onset year (b2k)",ylab="Density",
       main=str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]),
       xlim=xlim,ylim=ylim)
  #Axis(side=1,labels=FALSE,at=dens1$x,tck=0.1)

  #lines(dens,type="l",col="gray")
  if(FALSE){#95% credible intervals
    abline(v=mean(agesimmatrix_dust[i,]),col=1)
    zdens1 = inla.zmarginal(dens1,silent = TRUE)
    abline(v=c(zdens1$quant0.025,zdens1$quant0.975),col=1,lty=3,lwd=0.8)
    abline(v=mean(agesimmatrix_d18O[i,]),col=2)
    zdens2 = inla.zmarginal(dens2,silent = TRUE)
    abline(v=c(zdens2$quant0.025,zdens2$quant0.975),col=2,lty=3,lwd=0.8)
  }

  abline(v=event_intervals$GICC_age.yb2k[i],col="blue")  
  if(!is.na(buiind[i])){
    abline(v=buizert_onsets$Age[buiind[i]],col="green")
    #abline(v=buizert_ages[buiind[i]],col="green")
  }

  # abline(v=capron_ages_NGRIP_d18O,col="orange")
  # abline(v=capron_ages_NGRIP_dust,col="orange",lty=3)
  if(!is.na(capind[i])){
    abline(v=capron_ages_NGRIP_d18O[capind[i]],col="red")
  }

}
par(mar=c(0,0,0,0));plot(-1,axes=FALSE,xlab="",ylab="",xlim=c(0,1),ylim=c(0,1))
legend(0,1,legend=c("Onset posterior distribution",#"Ca2+ onset posterior",
                    "Rasmussen onset","Buizert onset",
                    "Capron onset"),#"Capron Ca2+ onset"),
       col=c("black", "blue", "green", "red"),
       lty=c(1,1,1,1,1,1), cex=1,bty="n")

par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)

}

rownames=c()
for(i in 1:29){
  rownames=c(rownames,str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]))
}

#colnames = colnames(depthstats)[1:4]
#colnames = c("name","d18O_mean","d18O_q0.025","d18O_q0.975")
depthresults = t(depthstats)[,2:4]
colnames(depthresults)=c("Z* mean", "Z* q0.025","Z* q0.975")

ageresults = t(agestats[2:4,])
colnames(ageresults)=c("Y* mean", "Y* q0.025","Y* q0.975")
allresults = cbind(depthresults,ageresults)
rownames(allresults)=rownames

latexstr = "Event & Z^* mean (m) & Z^* 95\\% CI (m) & Y^* mean (yb2k) & Y^* 95\\% CI (yb2k) \\\\ \n"
for(i in 1:29){
  latexstr = paste0(latexstr,rownames[i], "& ", format(round(allresults[i,1],2),nsmall=2)," & (", format(round(allresults[i,2],2),nsmall=2), ",
         ",format(round(allresults[i,3],2),nsmall=2),") & ", format(round(allresults[i,4],2),nsmall=2)," & (",format(round(allresults[i,5],2),nsmall=2),", ",format(round(allresults[i,6],2),nsmall=2),") \\\\ \n"
         )
}
cat(latexstr)
```

# Other stuff

To get started, see the documentation '?bremla' and the associated example.

The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). Also includes the tie-point presented in [Adolphi et al. (2018)](https://cp.copernicus.org/articles/14/1755/2018/) and [Muscheler et al. (2020)](https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D). Other data and tie-points should work as well.

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if synchronized chronologies is also computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Simulation example

To verify our model we can test it on simulated data where the true parameters are known.

```{r simex}
rm(list=ls())
require(stats)
library(ggplot2)
set.seed(123)

n=1000
y0 = 1000
z0 = 1200
phi=0.8; sigma=1.2

depth = z0+seq(from=0.5,500,length.out=1000)
dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
events=list(locations=c(1,250,750,1000))
a1 = c(-rep(1,249)*15,numeric(751))
a2 = c(numeric(249),rep(1,500)*50,numeric(251))
a3 = c(numeric(749),-rep(1,251)*110)
c1 = c(depth[1:249]/60,numeric(751))
c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
c3 = c(numeric(749),depth[1:251]/10)
dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5

age = y0+cumsum(dy)

df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth^2); data=rbind(c(y0,z0,0,0,0),df)

formula=dy~-1+proxy#+depth2

nsims = 10000

synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))

res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             events=list(locations=c(1,250,750,1000),locations_unit="index"),
             control.fit=list(noise="ar1"),
             synchronization=synchronization,
             control.sim = list(synchronized=TRUE,
                                nsims=nsims))

par(mfrow=c(1,2))
plot(res$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(sigma),ylab="Density",main=expression(paste("(a) Posterior distribution of ",sigma)))
abline(v=res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
           res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),lwd=1.5,col="red")

plot(res$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density",main=expression(paste("(b) Posterior distribution of ",phi)))
abline(v=res$fitting$inla$hyperparameters$results$phi$mean,lwd=1.5)
abline(v=c(res$fitting$inla$hyperparameters$results$phi$quant0.025,
           res$fitting$inla$hyperparameters$results$phi$quant0.975),lwd=1.5,col="red")
par(mfrow=c(1,1))

#plot results

  dsims = colDiffs(rbind(matrix(rep(res$initial$age,ncol(res$simulation$age)),nrow=1), res$simulation$age))
  dmeans = rowMeans(dsims)
  dsd = rowSds(dsims)
ggd = data.frame(y = res$data$dy,depth=res$data$depth,
                   mean=dmeans,
                   lower = dmeans-1.96*dsd,
                   upper = dmeans+1.96*dsd)
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+ #xlim(rev(range(results$data$depth)))+
    geom_line(aes(y=y),col="gray")+
    geom_line(aes(y=mean),col="blue",size=0.8)+
    geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.8)+
    xlab("Depth (m)")+ylab("Layers per 5cm")+ggtitle("(a) Fitted layer increment model")

  ggd2 = data.frame(mean=res$simulation$summary_sync$mean-res$data$age,
                    lower=res$simulation$summary_sync$lower-res$data$age,
                    upper=res$simulation$summary_sync$upper-res$data$age,
                    depth=res$data$depth
                    )

 reference=res$original.chron$age[1+1:n]
 reftie = reference[res$tie_points$locations_indexes]

ggdtie = data.frame(mean=synchronization$params$mean-reftie,
                    upper=synchronization$params$mean+1.96*synchronization$params$sd-reftie,
                    lower=synchronization$params$mean-1.96*synchronization$params$sd-reftie,
                    locations = res$data$depth[synchronization$locations])
ggp2 = ggplot(data=ggd2,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=lower,ymax=upper),color="red",fill="red",alpha=0.3)+
  geom_line(aes(y=mean),color="blue") +
  geom_linerange(data=ggdtie,aes(x=locations,ymin=lower,ymax=upper),color="magenta")+
  xlab("Depth (m)") + ylab("Deviation from simulated reference (years)")+
  ggtitle("(b) Simulated chronologies")

library(gridExtra)

ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))

str = ""
for(i in 1:length(res$fitting$inla$fit$summary.fixed$mean)){
  str=paste0(str," & & ",format(res$fitting$inla$fit$summary.fixed$mean[i],digits=7,scientific=FALSE), " & (",
             format(res$fitting$inla$fit$summary.fixed$`0.025quant`[i],digits=7,scientific=FALSE),", ",
             format(res$fitting$inla$fit$summary.fixed$`0.975quant`[i],digits=7,scientific=FALSE),") \\\\ \n")
}

cat(str)  
```

## Real data example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2

events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=list(method="adolphi"),
                control.fit=list(method="inla"),
                control.sim=list(synchronized=TRUE) )
summary(results)
```

```{r plot, message=FALSE, eval=TRUE, echo=FALSE}
require(ggplot2)
free_indexes = results$tie_points$free_indexes
tie_indexes = results$tie_points$tie_indexes
ageref_free = results$data$age[free_indexes]
ageref_tie = results$data$age[tie_indexes]

fullpd = data.frame(depth = results$data$depth[free_indexes],
                medians=results$simulation$summary_sync$median[free_indexes]-ageref_free,
                lower=results$simulation$summary_sync$lower[free_indexes]-ageref_free,
                upper=results$simulation$summary_sync$upper[free_indexes]-ageref_free)

gg2 = ggplot(data=fullpd,aes(x=depth)) +
geom_line(aes(y=rep(0,nrow(fullpd))),color="blue",linetype="dotted",size=0.2)+
theme_bw()+ylab("Estimated age - reference (years)")+
xlab("NGRIP depth (m)")

gg2 = gg2+geom_line(aes(y=medians))+
geom_ribbon(aes(ymin=.data$lower,ymax=.data$upper),color="red",fill="red",alpha=0.3)

gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=results$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
                   upper=results$simulation$summary_sync$upper[tie_indexes]-ageref_tie),
                   aes(x=.data$depth,y=.data$lower,xend=.data$depth,yend=.data$upper),
                   col="magenta")
print(gg2)
```

## Attribution

This code is associated and written for the papers Myrvoll-Nilsen et al. (2022) and Myrvoll-Nilsen et al. (202x) mentioned above. Feel free to use the code, but please cite the accompanying papers.

## License

The code in this repository is made available under the terms of the GNU (version >=2) License. For details, see LICENSE.md file.

### 2022-09-29 R/bremla.R

@@ -175,17 +175,6 @@ bremla <- function(formula,data,reference.label=NULL,
 
   }
 
-
-  if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
-    control.sim$nsims=nsims
-    #produce samples from the chronologies
-    object = bremla_chronology_simulation(object, control.sim=control.sim,
-                                          print.progress=print.progress)
-
-    #compute posterior marginal mean, quantiles and other summary statistics
-    #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=FALSE,print.progress=print.progress)
-
-  }
   if(!is.null(synchronization)){
     synchronization$nsims=nsims
     ##format and or simulate tie-points
@@ -195,15 +184,32 @@ bremla <- function(formula,data,reference.label=NULL,
   }
 
 
-  if(nsims>0 && control.sim$synchronized %in% c(TRUE,2)){
-    control.sim$nsims=nsims
-    #produce samples from the chronologies
-    object = bremla_synchronized_simulation(object, control.sim=control.sim,
+  if(!is.null(control.sim)){
+    if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
+      control.sim$nsims=nsims
+      #produce samples from the chronologies
+      object = bremla_chronology_simulation(object, control.sim=control.sim,
                                             print.progress=print.progress)
 
-    #compute posterior marginal mean, quantiles and other summary statistics
-    #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=TRUE,print.progress=print.progress)
+      #compute posterior marginal mean, quantiles and other summary statistics
+      #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=FALSE,print.progress=print.progress)
+
+    }
+
+
+
+    if(nsims>0 && control.sim$synchronized %in% c(TRUE,2)){
+      control.sim$nsims=nsims
+      #produce samples from the chronologies
+      object = bremla_synchronized_simulation(object, control.sim=control.sim,
+                                              print.progress=print.progress)
+
+      #compute posterior marginal mean, quantiles and other summary statistics
+      #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=TRUE,print.progress=print.progress)
+    }
+
   }
+
 
 
   #if control.transition_dating list object (containing specifications) is included, perform dating estimation

### 2022-09-29 R/bremla_prepare.R

@@ -108,7 +108,7 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
     age=data$age
   }else if("y" %in% colnames(data)){
     age=data$y
-    colnames(data$y)="age"
+    colnames(data)[which("y"==colnames(data))]="age"
   }else{
     stop("Could not find 'age' in data.frame. Stopping!")
   }
@@ -136,8 +136,10 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
       eventindexes = which.index (events$locations, z)
     }else if(tolower(events$locations_unit) %in% c("age","time","y")){
       eventindexes = which.index (events$locations, y)
+    }else{
+      eventindexes = events$locations
     }
-    eventindexes = unique(c(1,eventindexes[!is.na(eventindexes)]))
+    eventindexes = unique(c(1,eventindexes[!is.na(eventindexes)],length(y)))
     nevents = length(eventindexes)
   }else{
     nevents=0
@@ -157,43 +159,49 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
         if(events$degree>=1) formulastring=paste0(formulastring," + psi1_",i-1)
         if(events$degree==2) formulastring=paste0(formulastring," + psi2_",i-1)
       }
-      formulastring = paste0(formulastring, " + psi0_",nevents)
-      if(events$degree>=1) formulastring=paste0(formulastring," + psi1_",nevents)
-      if(events$degree==2) formulastring=paste0(formulastring," + psi2_",nevents)
+      # formulastring = paste0(formulastring, " + psi0_",nevents)
+      # if(events$degree>=1) formulastring=paste0(formulastring," + psi1_",nevents)
+      # if(events$degree==2) formulastring=paste0(formulastring," + psi2_",nevents)
     }
     if(events$fill_data){
       for(i in 2:nevents){
+        if(i==nevents){
+          subtract=0
+        }else{
+          subtract=1
+        }
         konst = numeric(n-1)
-        konst[eventindexes[i-1]:(eventindexes[i]-1)] = 1
+        konst[eventindexes[i-1]:(eventindexes[i]-subtract)] = 1
         data_obj[[paste0("psi0_",i-1)]] = konst
[…]

### 2022-09-29 R/default_arguments.R

@@ -130,6 +130,7 @@ control.sim.default <- function(){
 #'   distribution with mean and standard deviations supplied by user (TODO).
 #'   If \code{"adolphi"} is used, then \code{locations}, \code{locations_unit} is
 #'   filled out accordingly.
+#'   \item \code{params} List containing parameters used to simulate tie-points.
 #'   \item \code{x.ref} Numeric giving a reference for the x-axis (given by \code{locations_unit})
 #'   that will be included in future \code{plot}.
 #' }
@@ -148,6 +149,7 @@ synchronization.default <- function(){
       samples=NULL,
       nsims=10000,
       method="adolphis",
+      params = NULL,
       x.ref = NULL
     )
   )

### 2022-09-29 R/helpfulfunctions.R

@@ -524,13 +524,7 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     if(ncol(synchronization$samples) != length(synchronization$locations))
       warning("number of columns in 'samples' must correspond to length of 'locations'!")
     samples = synchronization$samples
-    if(synchronization$locations_unit %in% c("depth","z")){
-      locations_indexes = which.index(synchronization$locations,object$data$depth)
-    }else if(synchronization$locations_unit %in% c("age","y")){
-      locations_indexes = which.index(synchronization$locations,object$data$age)
-    }else if(synchronization$locations_unit %in% c("indexes","i")){
-      locations_indexes=synchronization$locations
-    }
+
 
   }else{
     if(synchronization$method=="adolphi"){
@@ -543,22 +537,37 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
       samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
                                         tieshifts=tieshifts,...)
     }else if(synchronization$method %in% c("normal","gaussian","gauss")){
-      synchronization$locations = object$data$depth[c(100,400,700)]
+      #synchronization$locations = object$data$depth[c(100,400,700)]
 
       ntie = length(synchronization$locations)
-      ## temporary
+      means = synchronization$params$mean
+      sds = synchronization$params$sd
       samples = matrix(NA,nrow=synchronization$nsims,ncol=ntie)
-      loc.ind = which.index(synchronization$locations,object$data$depth)
-      meanvek = object$data$age[loc.ind]+c(1,-100,300)
-      sdvek = object$data$age[loc.ind]/meanvek[1]*(1:3)*50
+
+      #loc.ind = which.index(synchronization$locations,object$data$depth)
+
       for(i in 1:ntie){
-        samples[,i] = rnorm(synchronization$nsims,mean=meanvek[i],sd=sdvek[i])
+        samples[,i] = rnorm(synchronization$nsims,mean=means[i],sd=sds[i])
       }
-      locations_indexes=loc.ind
+      ## temporary
+
+      # meanvek = object$data$age[loc.ind]+c(1,-100,300)
+      # sdvek = object$data$age[loc.ind]/meanvek[1]*(1:3)*50
+      # for(i in 1:ntie){
+      #   samples[,i] = rnorm(synchronization$nsims,mean=meanvek[i],sd=sdvek[i])
+      # }
+      #locations_indexes=loc.ind
     }
 
 
 
+  }
[…]

### 2022-09-29 inst/Examples/Paper_results/Paper_results_2.Rmd
Fixed bug in tiepointsimmer
Issue arose from adding tie-point-parameter specification

@@
 We then estimate the dating uncertainty for all 29 events
-```{r}
+```{r compare_transitions_1}
 agesimmatrix_d18O = matrix(NA,29,nsims)
@@
 
-```{r}
+```{r compare_transitions_2}
 

### 2022-09-29 R/bremla.R

@@ -56,7 +56,8 @@
 #'
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
-#' synchronization=list(method="gauss")
+#' synchronization=list(locations=depth[c(100,400,700)],locations_unit="depth",method="gauss",
+#'         params=list(mean=age[c(100,400,700)]+c(20,5,-20),sd=c(20,10,50)))
 #' control.sim=list(synchronized=2,
 #'                  summary=list(compute=TRUE))
 #'

### 2022-09-29 R/linrampfitter.R

@@ -39,7 +39,8 @@
 #'
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
-#' synchronization=list(method="gauss")
+#' synchronization=list(locations=depth[c(100,400,700)],locations_unit="depth",method="gauss",
+#'         params=list(mean=age[c(100,400,700)]+c(20,5,-20),sd=c(20,10,50)))
 #' control.sim=list(synchronized=TRUE,
 #'                  summary=list(compute=TRUE))
 #'

### 2022-09-29 R/plot.bremla.R

@@ -48,7 +48,7 @@
 #'
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
-#' control.sim=list(synchronized=2,
+#' control.sim=list(synchronized=FALSE,
 #'                  summary=list(compute=TRUE))
 #'
 #' object = bremla_prepare(formula,data,nsims=5000,reference.label="simulated timescale",

### 2022-09-29 R/print.bremla.R

@@ -35,7 +35,7 @@
 #'
 #' events=list(locations=c(1210,1220,1240))
 #' control.fit = list(ncores=2,noise="ar1")
-#' control.sim=list(synchronized=2,
+#' control.sim=list(synchronized=FALSE,
 #'                  summary=list(compute=TRUE))
 #'
 #' object = bremla_prepare(formula,data,nsims=5000,reference.label="simulated timescale",

### 2022-09-29 inst/Examples/Paper_results/Paper_results_2.Rmd
Revert "Minor changes"
This reverts commit fd599223a308a663a715e9ed2b04772279b7bd5d.

--- 2022-09-29 inst/Examples/Paper_results/Paper_results_2.Rmd
@@
 These are included in the package and can be imported as follows.
-```{r initialize, eval=FALSE}
+```{r install, eval=FALSE}
 #library(bremla)
@@
 hist(results_event$event_dating$samples,main=paste0(transitionlabel," Onset age"),
-     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40, freq=0)
+     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40)
 abline(v=age.reference,lty=3)
@@
 
-## Simulation example
-
-To verify our model we can test it on simulated data where the true parameters are known.
-
-```{r simex}
-rm(list=ls())
-require(stats)
-library(ggplot2)
-set.seed(123)
-
-n=1000
-y0 = 1000
-z0 = 1200
-phi=0.8; sigma=1.2
-
-depth = z0+seq(from=0.5,500,length.out=1000)
-dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
-proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
-events=list(locations=c(1,250,750,1000))
-a1 = c(-rep(1,249)*15,numeric(751))
-a2 = c(numeric(249),rep(1,500)*50,numeric(251))
-a3 = c(numeric(749),-rep(1,251)*110)
-c1 = c(depth[1:249]/60,numeric(751))
-c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
-c3 = c(numeric(749),depth[1:251]/10)
-dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5
-
-age = y0+cumsum(dy)
-
-df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth^2); data=rbind(c(y0,z0,0,0,0),df)
-
-formula=dy~-1+proxy#+depth2
-
-nsims = 10000
-
-synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))
-
-res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
-             events=list(locations=c(1,250,750,1000),locations_unit="index"),
-             control.fit=list(noise="ar1"),
-             synchronization=synchronization,
-             control.sim = list(synchronized=TRUE,
-                                nsims=nsims))
-
-par(mfrow=c(1,2))
-plot(res$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(sigma),ylab="Density",main=expression(paste("(a) Posterior distribution of ",sigma)))
-abline(v=res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,lwd=1.5)
-abline(v=c(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
-           res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),lwd=1.5,col="red")
-
-plot(res$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density",main=expression(paste("(b) Posterior distribution of ",phi)))
-abline(v=res$fitting$inla$hyperparameters$results$phi$mean,lwd=1.5)
-abline(v=c(res$fitting$inla$hyperparameters$results$phi$quant0.025,
-           res$fitting$inla$hyperparameters$results$phi$quant0.975),lwd=1.5,col="red")
-par(mfrow=c(1,1))
-
-#plot results
-
-  dsims = colDiffs(rbind(matrix(rep(res$initial$age,ncol(res$simulation$age)),nrow=1), res$simulation$age))
-  dmeans = rowMeans(dsims)
-  dsd = rowSds(dsims)
-ggd = data.frame(y = res$data$dy,depth=res$data$depth,
-                   mean=dmeans,
-                   lower = dmeans-1.96*dsd,
-                   upper = dmeans+1.96*dsd)
-  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+ #xlim(rev(range(results$data$depth)))+
-    geom_line(aes(y=y),col="gray")+
-    geom_line(aes(y=mean),col="blue",size=0.8)+
-    geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.8)+
-    xlab("Depth (m)")+ylab("Layers per 5cm")+ggtitle("(a) Fitted layer increment model")
-
-  ggd2 = data.frame(mean=res$simulation$summary_sync$mean-res$data$age,
-                    lower=res$simulation$summary_sync$lower-res$data$age,
-                    upper=res$simulation$summary_sync$upper-res$data$age,
-                    depth=res$data$depth
-                    )
-
- reference=res$original.chron$age[1+1:n]
- reftie = reference[res$tie_points$locations_indexes]
-
-ggdtie = data.frame(mean=synchronization$params$mean-reftie,
-                    upper=synchronization$params$mean+1.96*synchronization$params$sd-reftie,
-                    lower=synchronization$params$mean-1.96*synchronization$params$sd-reftie,
-                    locations = res$data$depth[synchronization$locations])
-ggp2 = ggplot(data=ggd2,aes(x=depth)) + theme_bw()+
-  geom_ribbon(aes(ymin=lower,ymax=upper),color="red",fill="red",alpha=0.3)+
-  geom_line(aes(y=mean),color="blue") +
-  geom_linerange(data=ggdtie,aes(x=locations,ymin=lower,ymax=upper),color="magenta")+
-  xlab("Depth (m)") + ylab("Deviation from simulated reference (years)")+
-  ggtitle("(b) Simulated chronologies")
-
-library(gridExtra)
-
-ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
-
-str = ""
-for(i in 1:length(res$fitting$inla$fit$summary.fixed$mean)){
-  str=paste0(str," & & ",format(res$fitting$inla$fit$summary.fixed$mean[i],digits=7,scientific=FALSE), " & (",
-             format(res$fitting$inla$fit$summary.fixed$`0.025quant`[i],digits=7,scientific=FALSE),", ",
-             format(res$fitting$inla$fit$summary.fixed$`0.975quant`[i],digits=7,scientific=FALSE),") \\\\ \n")
-}
-
-cat(str)  
-```
-
-## Real data example
+## Example
 This is a real data example which produces 5000 simulations from a synchronized time scale.

### 2022-09-29 R/bremla.R

@@ -175,6 +175,17 @@ bremla <- function(formula,data,reference.label=NULL,
 
   }
 
+
+  if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
+    control.sim$nsims=nsims
+    #produce samples from the chronologies
+    object = bremla_chronology_simulation(object, control.sim=control.sim,
+                                          print.progress=print.progress)
+
+    #compute posterior marginal mean, quantiles and other summary statistics
+    #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=FALSE,print.progress=print.progress)
+
+  }
   if(!is.null(synchronization)){
     synchronization$nsims=nsims
     ##format and or simulate tie-points
@@ -184,32 +195,15 @@ bremla <- function(formula,data,reference.label=NULL,
   }
 
 
-  if(!is.null(control.sim)){
-    if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
-      control.sim$nsims=nsims
-      #produce samples from the chronologies
-      object = bremla_chronology_simulation(object, control.sim=control.sim,
+  if(nsims>0 && control.sim$synchronized %in% c(TRUE,2)){
+    control.sim$nsims=nsims
+    #produce samples from the chronologies
+    object = bremla_synchronized_simulation(object, control.sim=control.sim,
                                             print.progress=print.progress)
 
-      #compute posterior marginal mean, quantiles and other summary statistics
-      #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=FALSE,print.progress=print.progress)
-
-    }
-
-
-
-    if(nsims>0 && control.sim$synchronized %in% c(TRUE,2)){
-      control.sim$nsims=nsims
-      #produce samples from the chronologies
-      object = bremla_synchronized_simulation(object, control.sim=control.sim,
-                                              print.progress=print.progress)
-
-      #compute posterior marginal mean, quantiles and other summary statistics
-      #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=TRUE,print.progress=print.progress)
-    }
-
+    #compute posterior marginal mean, quantiles and other summary statistics
+    #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=TRUE,print.progress=print.progress)
   }
-
 
 
   #if control.transition_dating list object (containing specifications) is included, perform dating estimation

### 2022-09-29 R/helpfulfunctions.R

@@ -524,7 +524,13 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     if(ncol(synchronization$samples) != length(synchronization$locations))
       warning("number of columns in 'samples' must correspond to length of 'locations'!")
     samples = synchronization$samples
-
+    if(synchronization$locations_unit %in% c("depth","z")){
+      locations_indexes = which.index(synchronization$locations,object$data$depth)
+    }else if(synchronization$locations_unit %in% c("age","y")){
+      locations_indexes = which.index(synchronization$locations,object$data$age)
+    }else if(synchronization$locations_unit %in% c("indexes","i")){
+      locations_indexes=synchronization$locations
+    }
 
   }else{
     if(synchronization$method=="adolphi"){
@@ -537,37 +543,22 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
       samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
                                         tieshifts=tieshifts,...)
     }else if(synchronization$method %in% c("normal","gaussian","gauss")){
-      #synchronization$locations = object$data$depth[c(100,400,700)]
+      synchronization$locations = object$data$depth[c(100,400,700)]
 
       ntie = length(synchronization$locations)
-      means = synchronization$params$mean
-      sds = synchronization$params$sd
-      samples = matrix(NA,nrow=synchronization$nsims,ncol=ntie)
-
-      #loc.ind = which.index(synchronization$locations,object$data$depth)
-
-      for(i in 1:ntie){
-        samples[,i] = rnorm(synchronization$nsims,mean=means[i],sd=sds[i])
-      }
       ## temporary
-
-      # meanvek = object$data$age[loc.ind]+c(1,-100,300)
-      # sdvek = object$data$age[loc.ind]/meanvek[1]*(1:3)*50
-      # for(i in 1:ntie){
-      #   samples[,i] = rnorm(synchronization$nsims,mean=meanvek[i],sd=sdvek[i])
-      # }
-      #locations_indexes=loc.ind
+      samples = matrix(NA,nrow=synchronization$nsims,ncol=ntie)
+      loc.ind = which.index(synchronization$locations,object$data$depth)
+      meanvek = object$data$age[loc.ind]+c(1,-100,300)
+      sdvek = object$data$age[loc.ind]/meanvek[1]*(1:3)*50
+      for(i in 1:ntie){
[…]

### 2022-09-29 R/default_arguments.R
temp

@@ -162,7 +162,8 @@ synchronization.default <- function(){
 #'    \item \code{proxy} Numeric giving the proxy data to perform the linear ramp model fit.
 #'    Default value is \code{NULL} meaning that it must be provided by the user.
 #'    \item \code{interval} Interval describing the window of \code{proxy} to perform the
-#'    linear ramp model fit. Default value is \code{NULL} so it must be given by the user.
+#'    linear ramp model fit. Default value is \code{NULL} which will tell the \code{linramp}
+#'    function to use \code{1:length(proxy)}.
 #'    \item \code{interval.unit} Character giving which axis is used for interval. Can be
 #'    \code{"depth"} (default), \code{"age"} or \code{"index"}.
 #'    \item \code{depth.label} Character describing what label to place on x-axis (depth) in a
@@ -178,10 +179,14 @@ synchronization.default <- function(){
 #'    compute transition end point \code{t1} if \code{compute.t1} is \code{TRUE}.
 #'    \item \code{h} Step length of optimization procedure in the \code{inla}-program.
 #'    Default value is \code{0.01}.
+#'    \item \code{imp.fit} Boolean indicaating whether initial values for \code{inla} be computed using \code{optim}.
 #'    \item \code{ncores} Integer giving the number of cores used in the \code{inla}-program.
 #'    For \code{rgeneric} models which are used here, stability can sometimes be
 #'    improved if the number of cores is set to \code{1}, which is why it is the
 #'    default here.
+#'    \item \code{silent}. If equal to \code{TRUE} or \code{1L}, then \code{inla} will run in "silent" mode,
+#'    If equal to \code{2L} then error messages will also be suppressed.
+#'    \item \code{priorparams} List object containing values for the priors for the hyperparameters.
 #'    \item \code{verbose} Boolean stating whether or not the \code{inla}-program should
 #'    run in verbose mode. Default \code{FALSE}.
 #' }
@@ -202,8 +207,11 @@ control.linramp.default <- function(){
       opt.params=NULL, #function will recognize this
       compute.t1=TRUE,
       nsims=30000,
-      h=0.01,
+      h=0.005,
+      imp.fit=FALSE,
       ncores=1,
+      silent=FALSE,
+      priorparams = NULL,
       verbose=FALSE
       )
     )

### 2022-09-29 R/helpfulfunctions.R

@@ -1,4 +1,5 @@
 <<<<<<< HEAD
+<<<<<<< HEAD
 #' Mean vector from fixed effects
 #'
 #' Gives the mean number of layer differences per depth from the fixed effects.
@@ -666,6 +667,8 @@ Qymaker = function(Qx){
   return(t(S)%*%Qx%*%S)
 }
 =======
+=======
+>>>>>>> parent of 67d821f (Revert "Added some more features and Rmd files to generate the plots and results from both papers")
 #' Mean vector from fixed effects
 #'
 #' Gives the mean number of layer differences per depth from the fixed effects.
@@ -1192,7 +1195,17 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     if(ncol(synchronization$samples) != length(synchronization$locations))
       warning("number of columns in 'samples' must correspond to length of 'locations'!")
     samples = synchronization$samples
+<<<<<<< HEAD
 
+=======
+    if(synchronization$locations_unit %in% c("depth","z")){
+      locations_indexes = which.index(synchronization$locations,object$data$depth)
+    }else if(synchronization$locations_unit %in% c("age","y")){
+      locations_indexes = which.index(synchronization$locations,object$data$age)
+    }else if(synchronization$locations_unit %in% c("indexes","i")){
+      locations_indexes=synchronization$locations
+    }
+>>>>>>> parent of 67d821f (Revert "Added some more features and Rmd files to generate the plots and results from both papers")
 
   }else{
     if(synchronization$method=="adolphi"){
@@ -1205,6 +1218,7 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
       samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
                                         tieshifts=tieshifts,...)
     }else if(synchronization$method %in% c("normal","gaussian","gauss")){
+<<<<<<< HEAD
       #synchronization$locations = object$data$depth[c(100,400,700)]
 
       ntie = length(synchronization$locations)
@@ -1225,11 +1239,26 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
       #   samples[,i] = rnorm(synchronization$nsims,mean=meanvek[i],sd=sdvek[i])
       # }
       #locations_indexes=loc.ind
+=======
+      synchronization$locations = object$data$depth[c(100,400,700)]
+
[…]

### 2022-09-29 R/linrampfitter.R

@@ -97,6 +97,10 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(is.null(object$simulation$age) && is.null(object$simulation$age_sync))
     stop("Chronologies not found. Run 'bremla_synchronized_simulation' or 'bremla_chronology_simulation' first!")
   interval = control.linramp$interval
+  if(is.null(interval)){
+    interval = 1:length(control.linramp$proxy)
+    control.linramp$interval.unit = "index"
+  }
 
 
 
@@ -110,18 +114,19 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
     interval = range(interval)[1]:range(interval)[2]
   }
 
+  n=length(interval)
   if(print.progress) cat("Initializing linear ramp fit using INLA.\n",sep="")
   df_event = data.frame(xx=rev(object$data$depth[interval]),
                         yy=rev(control.linramp$proxy[interval])) #reverse: Want depth axis representing old->new
 
-  ## default initial values for optim function
-  if(is.null(control.linramp$opt.params)){
-    optparams = c(round(length(interval)/2),round(length(interval)/10),df_event$yy[1],df_event$yy[length(interval)]-df_event$yy[1])
-  }else{
-    optparams = control.linramp$opt.params
-  }
 
-  n=length(interval)
+  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
+  ypointsmin = min(df_event$yy)
+  ypointsmax = max(df_event$yy-ypointsmin)
+  #ypointsmax=1; ypointsmin=0
+  ypoints =(df_event$yy-ypointsmin)/ypointsmax
+
+
   df0=data.frame(y=df_event$yy,x=df_event$xx)
 
   t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
@@ -131,63 +136,111 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(print.progress) cat("Using 'optim' to find initial positions for hyperparameters in INLA.\n",sep="")
   ## for stability the x-axis is transformed to values ranging from 1:n.
 
-  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
+
   df0$time = timepoints
+  df0$proxy = ypoints
 
[…]

### 2022-09-29 inst/Examples/Paper_results/Paper_results_1.Rmd
Revert back to something that was working

@@
 data("NGRIP_5cm")
-
 age = NGRIP_5cm$age
@@
 n=length(diff(age))
-
 data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
@@
             locations_unit="depth",degree=1)
-
 plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
@@
 formula = dy~-1+depth2+proxy
-
 nsims=3000
-
 results = bremla(formula,data,reference.label="GICC05",
@@
                 )
-
 ```
@@
 fit = results$fitting$LS$fit
-
 {
@@
                 )
-
 resultsar2 = bremla(formula,data,reference.label="GICC05",
@@
                 )
-
 gicc05 = results$original.chron$age[1+1:n]
@@
 ```{r biased, eval=TRUE}
-
 results_biased = bremla(formula,data,reference.label="GICC05",
@@
                 )
-
 {
@@
 }
-
 ```
@@
 ```{r events_individual, eval=TRUE}
-
 eventnumber=13 #number between 1 and 29. specifies which transition to consider
-
 #load data window and specifics to transition
@@
 interval = lowerints[eventnumber]:upperints[eventnumber]
-
 transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                       str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])
-
 control.linramp=list(proxy=proxy,
@@
                      )
-
 results_event = linrampfitter(results,control.linramp)
-
 control.transition_dating = list(label="Simulated transition")
@@
 abline(v=depth.reference,lty=3,lwd=0.7,col="black")
-
 ybottom = min(results_event$linramp$data$y)
@@
 lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
-
 plot(results_event$linramp$param$t0$marg.t0,type="l",xlab="Depth (m)",ylab="Density",
@@
 abline(v=results_event$linramp$param$t0$q0.975,col="gray")
-
 hist(results_event$event_dating$samples,main="(c) Histogram of onset age",
@@
 abline(v=results_event$event_dating$mean,lwd=2)
-
 par(mar=c(5,3.5,4,2)+0.1)
 }
-
 ```
@@
 rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`
-
 buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
@@
 buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]
-
 capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
@@
 capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`
-
 capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
@@
 capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`
-
 capind = numeric(29); buiind = numeric(29)
-
 capind = c(NA,2,3, 4,5,6,NA,NA,7,8,NA,9,10,11,NA,NA,NA,NA,NA,12,13,14,NA,NA,15,NA,NA,16,17)
@@
 logdust = log(dust_filled)
-
 data2=data
@@
 colnames(agestats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")
-
 # below is some fiddling with optimization parameters to improve the linear ramp fit for 
 # some of the more challenging fits
-
 steplengths = rep(0.005,29) 
@@
 #t0params1[10] = 120; t0params2[10] = 20
-
 for(eventnumber in 1:29){
@@
        lty=c(1,1,1,1,1,1), cex=0.65,bty="n")
-
 par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
@@
 ```{r events_table, eval=TRUE}
-
 colnames = c("d18O_mean","d18O_q0.025","d18O_q0.975",
@@
 rownames(depthresults)=rownames
-
 ageresults = t(agestats[2:7,])
@@
 rownames(ageresults)=rownames
-
 ```

### 2022-09-29 inst/Examples/Paper_results/Paper_results_2.Rmd
Revert back to something that was working

@@
 These are included in the package and can be imported as follows.
-```{r install, eval=FALSE}
+```{r initialize, eval=FALSE}
 #library(bremla)
@@
 formula = dy~-1+depth2+proxy
-
 ```
@@
 ```{r adolphi, eval=TRUE}
-
 adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = TRUE,
                             x.ref=c(11050,12050,13050,22050,42050))
-
 ```
@@
 nsims=10000
-
 results = bremla(formula,data,reference.label="GICC05",
@@
 }
-
 {
@@
 ```{r fixed, eval=TRUE}
-
 fixedsims = cbind(#rep(inla.mmarginal(adolphipdfs$tie1),nsims),
@@
                        locations_unit="depth")
-
 results_fixed = bremla(formula,data,reference.label="GICC05",
@@
                        locations_unit="depth")
-
 results_sync = bremla(formula,data,reference.label="GICC05",
@@
                 )
-
 tiepointranges = as.data.frame(matrix(NA,nrow=5,ncol=3))
-
 gicc05=results_sync$original.chron$age[1+1:n]
 gicc05tie = gicc05[results_sync$tie_points$locations_indexes]
-
 colnames(tiepointranges)=c("lower","upper","depth")
@@
 tiepointranges[,3] = depth[results_sync$tie_points$locations_indexes]
-
 {
@@
 ```{r events_individual, eval=FALSE}
-
 eventnumber=13 #number between 1 and 29. specifies which transition to consider
-
 #load data window and specifics to transition
@@
                      )
-
 results_event = linrampfitter(results_sync,control.linramp)
-
 control.transition_dating = list(label="Simulated transition")
@@
 lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
-
 plot(results_event$linramp$param$t0$marg.t0,type="l",xlab=expression(paste("Onset depth (m)")),ylab="Density",
@@
 abline(v=results_event$linramp$param$t0$q0.975,col="red")
-
 plot(results_event$linramp$param$t1$marginal,type="l",xlab=expression(paste("End point depth (m)")),ylab="Density",
@@
 abline(v=results_event$linramp$param$t1$q0.975,col="red")
-
 par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
-
 hist(results_event$event_dating$samples,main=paste0(transitionlabel," Onset age"),
-     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40)
+     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40, freq=0)
 abline(v=age.reference,lty=3)
@@
 abline(v=results_event$event_dating$q0.975,lwd=2,col="red")
-
 }
-
 ```
@@
 colnames(agestats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")
-
 steplengths = rep(0.01,29)
 steplengths[21]=0.001 #sometimes changing the steplength slightly might improve convergence (default is 0.01)
-
 for(eventnumber in 1:29){
@@
 }
-
 ```
 
-```{r}
-
+```{r compare_transitions_2}
 {
@@
        lty=c(1,1,1,1,1,1), cex=1,bty="n")
-
 par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
-
 }
-
 rownames=c()
@@
 }
-
 #colnames = colnames(depthstats)[1:4]
@@
 colnames(depthresults)=c("Z* mean", "Z* q0.025","Z* q0.975")
-
 ageresults = t(agestats[2:4,])
@@
 rownames(allresults)=rownames
-
 latexstr = "Event & Z^* mean (m) & Z^* 95\\% CI (m) & Y^* mean (yb2k) & Y^* 95\\% CI (yb2k) \\\\ \n"
@@
 
-## Example
+## Simulation example
+
+To verify our model we can test it on simulated data where the true parameters are known.
+
+```{r simex}
+rm(list=ls())
+require(stats)
+library(ggplot2)
+set.seed(123)
+n=1000
+y0 = 1000
+z0 = 1200
+phi=0.8; sigma=1.2
+depth = z0+seq(from=0.5,500,length.out=1000)
+dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
+proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
+events=list(locations=c(1,250,750,1000))
+a1 = c(-rep(1,249)*15,numeric(751))
+a2 = c(numeric(249),rep(1,500)*50,numeric(251))
+a3 = c(numeric(749),-rep(1,251)*110)
+c1 = c(depth[1:249]/60,numeric(751))
+c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
+c3 = c(numeric(749),depth[1:251]/10)
+dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5
+age = y0+cumsum(dy)
+df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth^2); data=rbind(c(y0,z0,0,0,0),df)
+formula=dy~-1+proxy#+depth2
+nsims = 10000
+synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))
+res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
+             events=list(locations=c(1,250,750,1000),locations_unit="index"),
+             control.fit=list(noise="ar1"),
+             synchronization=synchronization,
+             control.sim = list(synchronized=TRUE,
+                                nsims=nsims))
+par(mfrow=c(1,2))
+plot(res$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(sigma),ylab="Density",main=expression(paste("(a) Posterior distribution of ",sigma)))
+abline(v=res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,lwd=1.5)
+abline(v=c(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
+           res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),lwd=1.5,col="red")
+plot(res$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density",main=expression(paste("(b) Posterior distribution of ",phi)))
+abline(v=res$fitting$inla$hyperparameters$results$phi$mean,lwd=1.5)
+abline(v=c(res$fitting$inla$hyperparameters$results$phi$quant0.025,
+           res$fitting$inla$hyperparameters$results$phi$quant0.975),lwd=1.5,col="red")
+par(mfrow=c(1,1))
+#plot results
+
+  dsims = colDiffs(rbind(matrix(rep(res$initial$age,ncol(res$simulation$age)),nrow=1), res$simulation$age))
+  dmeans = rowMeans(dsims)
+  dsd = rowSds(dsims)
+ggd = data.frame(y = res$data$dy,depth=res$data$depth,
+                   mean=dmeans,
+                   lower = dmeans-1.96*dsd,
+                   upper = dmeans+1.96*dsd)
+  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+ #xlim(rev(range(results$data$depth)))+
+    geom_line(aes(y=y),col="gray")+
+    geom_line(aes(y=mean),col="blue",size=0.8)+
+    geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.8)+
+    xlab("Depth (m)")+ylab("Layers per 5cm")+ggtitle("(a) Fitted layer increment model")
+
+  ggd2 = data.frame(mean=res$simulation$summary_sync$mean-res$data$age,
+                    lower=res$simulation$summary_sync$lower-res$data$age,
+                    upper=res$simulation$summary_sync$upper-res$data$age,
+                    depth=res$data$depth
+                    )
+
+ reference=res$original.chron$age[1+1:n]
+ reftie = reference[res$tie_points$locations_indexes]
+ggdtie = data.frame(mean=synchronization$params$mean-reftie,
+                    upper=synchronization$params$mean+1.96*synchronization$params$sd-reftie,
+                    lower=synchronization$params$mean-1.96*synchronization$params$sd-reftie,
+                    locations = res$data$depth[synchronization$locations])
+ggp2 = ggplot(data=ggd2,aes(x=depth)) + theme_bw()+
+  geom_ribbon(aes(ymin=lower,ymax=upper),color="red",fill="red",alpha=0.3)+
+  geom_line(aes(y=mean),color="blue") +
+  geom_linerange(data=ggdtie,aes(x=locations,ymin=lower,ymax=upper),color="magenta")+
+  xlab("Depth (m)") + ylab("Deviation from simulated reference (years)")+
+  ggtitle("(b) Simulated chronologies")
+library(gridExtra)
+ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+str = ""
+for(i in 1:length(res$fitting$inla$fit$summary.fixed$mean)){
+  str=paste0(str," & & ",format(res$fitting$inla$fit$summary.fixed$mean[i],digits=7,scientific=FALSE), " & (",
+             format(res$fitting$inla$fit$summary.fixed$`0.025quant`[i],digits=7,scientific=FALSE),", ",
+             format(res$fitting$inla$fit$summary.fixed$`0.975quant`[i],digits=7,scientific=FALSE),") \\\\ \n")
+}
+cat(str)  
+```
+
+## Real data example
 This is a real data example which produces 5000 simulations from a synchronized time scale.
@@
 proxy = NGRIP_5cm$d18O
-
 data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
 formula = dy~-1+depth2
-
 events=list(locations = events_rasmussen$depth,
             locations_unit="depth",degree=1)
-
 results = bremla(formula,data,reference.label="GICC05",
@@
 ageref_tie = results$data$age[tie_indexes]
-
 fullpd = data.frame(depth = results$data$depth[free_indexes],
@@
                 upper=results$simulation$summary_sync$upper[free_indexes]-ageref_free)
-
 gg2 = ggplot(data=fullpd,aes(x=depth)) +
@@
 xlab("NGRIP depth (m)")
-
 gg2 = gg2+geom_line(aes(y=medians))+
 geom_ribbon(aes(ymin=.data$lower,ymax=.data$upper),color="red",fill="red",alpha=0.3)
-
 gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=results$simulation$summary_sync$lower[tie_indexes]-ageref_tie,

### 2022-09-29 R/default_arguments.R

@@ -182,6 +182,8 @@ synchronization.default <- function(){
 #'    For \code{rgeneric} models which are used here, stability can sometimes be
 #'    improved if the number of cores is set to \code{1}, which is why it is the
 #'    default here.
+#'    \item \code{silent} If equal to \code{TRUE} or \code{1L}, then \code{inla} will run in "silent" mode,
+#'    If equal to \code{2L} then error messages will also be suppressed.
 #'    \item \code{verbose} Boolean stating whether or not the \code{inla}-program should
 #'    run in verbose mode. Default \code{FALSE}.
 #' }
@@ -204,6 +206,7 @@ control.linramp.default <- function(){
       nsims=30000,
       h=0.01,
       ncores=1,
+      silent=FALSE,
       verbose=FALSE
       )
     )

### 2022-09-29 R/helpfulfunctions.R

@@ -86,6 +86,7 @@ linramprev = function(t,t0=0,dt=1,y0=0,dy=1){
 #'
 #' @param events Vector describing the values of interest (here: either depth or age of events)
 #' @param record Vector we want to search (here: either full depth or age record)
+#' @param suppress Boolean. Set to \code{TRUE} to suppress warnings that events fall outside designated interval.
 #'
 #' @return Returns a vector of indices for which elements of \code{record} best matches the elements of \code{events}-
 #' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
@@ -93,11 +94,14 @@ linramprev = function(t,t0=0,dt=1,y0=0,dy=1){
 #' @keywords indices
 #'
 #' @export
-which.index = function(events, record){ ## Finds which indices of 'record' that are located closest to the 'event's
+which.index = function(events, record,suppress=TRUE){ ## Finds which indices of 'record' that are located closest to the 'event's
   eventindexes = numeric(length(events))
   for(i in 1:length(events)){
     if(events[i] < min(record) || events[i] > max(record)){ #Gives NA if located outside range of 'record'
-      warning(paste0("Event ",i,", located at ",events[i]," is outside the interval covered by 'record' (",min(record),", ",max(record),"). The event will be omitted!"))
+      if(!suppress){
+        warning(paste0("Event ",i,", located at ",events[i]," is outside the interval covered by 'record' (",min(record),", ",max(record),"). The event will be omitted!"))
+      }
+
       eventindexes[i] = NA
     }else{
       eventindexes[i] = which(abs(events[i]-record) == min(abs(events[i]-record)))

### 2022-09-30 inst/Examples/Paper_results/Paper_results_1.Rmd
Reimplementing features (part 1)

@@
 
-We also import table XXXX from Myrvoll-Nilsen et al.
+We also import table D1 from Myrvoll-Nilsen et al.
 (2022) which gives suitable data windows for each abrupt warming transition that is analysed.
@@
 rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`
-buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
-                            col_types=c("text","numeric","numeric","numeric","numeric"))
+# buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
+#                             col_types=c("text","numeric","numeric","numeric","numeric"))
 buizert_depths = buizert_onsets$Depth; buizert_ages = buizert_onsets$Age
@@
 buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]
-capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
-                           sheet="d18O",n_max=25)
+# capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
+#                            sheet="d18O",n_max=25)
 capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`
-capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
-                                sheet="Ca2+",n_max=24)
+# capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
+#                                 sheet="Ca2+",n_max=24)
 capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`
@@
   control.linramp = list(proxy=proxy,interval=interval_d18O,interval.unit="index",
-                         depth.ref=depth.reference, imp.fit=imp.fits[eventnumber],
+                         depth.ref=depth.reference, #imp.fit=imp.fits[eventnumber],
                          h=steplengths[eventnumber],
@@
   control.linramp_dust = list(proxy=logdust,interval=interval_dust,interval.unit="index",
-                         depth.ref=depth.reference, imp.fit=imp.fits[eventnumber],
+                         depth.ref=depth.reference, #imp.fit=imp.fits[eventnumber],
                          h=steplengths[eventnumber],

### 2022-09-30 R/bremla_synchronized_simulation.R

@@ -205,7 +205,7 @@ time.start = Sys.time()
     #mu_amidb = (free_mu - solve(Qa)%*%Qab%*%(skewsamples[r]-tie_mu))[,1]
 
     #samples[free_indexes,r] = Qsimmer(1,Qa,mu_amidb)
-    if((r %% 1000) == 0){
+    if(print.progress && (r %% 1000) == 0){
       cat("Synchronous age simulation ",r,"/",nsims,". Elapsed time: ",difftime(Sys.time(),timeage.start,units="secs")[[1]]," seconds...","\n",sep="")
     }
   }

### 2022-09-30 R/buizert_onsets.R

@@ -1,0 +1,26 @@
+#' Buizert transition onsets
+#'
+#' Transition onsets of abrupt warmings in the northern hemisphere. Taken from
+#' Table 1 of Buizert et al. (2015). This data set is only used to compare
+#' our transition dating estimation in "Examples/Paper_results/Paper_results_1.Rmd"
+#' and "Examples/Paper_results/Paper_results_2.Rmd".
+#'
+#' @format A \code{tbl} \code{data.frame} with 22 rows and 5 columns:
+#' \describe{
+#'   \item{Transition}{Name of the transition.}
+#'   \item{Depth}{NGRIP depth of the onset of the transition (in meters).}
+#'   \item{Age}{NGRIP depth of the onset of the transition (in years before present).}
+#'   \item{`Age uncertainty`}{The dating uncertainty of the transition onset (in years).}
+#'   \item{Midpoint}{The midpoint of the transition (in years).}
+#'   }
+#' @references Buizert, C., Cuffey, K. M., Severinghaus, J. P., Baggenstos, D.,
+#' Fudge, T. J., Steig, E. J., Markle, B. R., Winstrup, M., Rhodes, R. H.,
+#' Brook, E. J., Sowers, T. A., Clow, G. D., Cheng, H., Edwards, R. L.,
+#' Sigl, M., McConnell, J. R., and Taylor, K. C.: The WAIS Divide deep ice core
+#' WD2014 chronology – Part 1: Methane synchronization (68–31 ka BP) and the gas
+#' age–ice age difference, Clim. Past, 11, 153–173, https://doi.org/10.5194/cp-11-153-2015,
+#' 2015.
+
+
+
+"buizert_onsets"

### 2022-09-30 R/capron_onsets_NGRIP_dust.R

@@ -1,0 +1,22 @@
+#' Capron transition onsets (dust)
+#'
+#' Abrupt warmings observed in the log(Ca2+) record of the NGRIP ice core. Taken from
+#' Supplementary data 4 in Capron et al. (2021). This data set is only used to compare
+#' our transition dating estimation in "Examples/Paper_results/Paper_results_1.Rmd"
+#' and "Examples/Paper_results/Paper_results_2.Rmd".
+#'
+#' @format A \code{tbl data.frame} with 25 rows and 16 columns. The following columns
+#' are relevant to analyses carried out in this package:
+#' \describe{
+#'   \item{`NGRIP dust`}{Name of the transition.}
+#'   \item{`t1 (50\%)`}{The median of the transition onset in years before present
+#'   (used as point estimate).}
+#'   \item{`d1 (50\%)`}{The median of the transition onset depth in meters.}
+#'   }
+#' @references Capron, E., Rasmussen, S.O., Popp, T.J. et al. The anatomy of past
+#' abrupt warmings recorded in Greenland ice. Nat Commun 12, 2106 (2021).
+#' https://doi.org/10.1038/s41467-021-22241-w
+
+
+
+"capron_onsets_NGRIP_dust"

### 2022-09-30 inst/Examples/Paper_results/Paper_results_1.Rmd
Reimplementing features (part 2)
Added rescaling of linear ramp data (so as to avoid having to retune priors for dust proxy)

@@
 
-  control.linramp = list(proxy=proxy,interval=interval_d18O,interval.unit="index",
+  control.linramp_d18O = list(proxy=proxy,interval=interval_d18O,interval.unit="index",
                          depth.ref=depth.reference, #imp.fit=imp.fits[eventnumber],
@@
                          priorparams=priorparams,
+                         rescale.y.factor=1,
+                         imp.fit=TRUE,
                          silent=2L,
@@
                          priorparams=priorparams,
+                         rescale.y.factor=-20,
+                         imp.fit=TRUE,
                          silent=2L,
@@
 
-  temp_d18O = linrampfitter(results,control.linramp)
+  temp_d18O = linrampfitter(results,control.linramp_d18O)
   if(do.plot.rampfit){
@@
 
-  control.transition_dating=list(label="d18O")
+  control.transition_dating=list(label="d18O",sync=FALSE)
   agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)
@@
   }
-  control.transition_dating_dust=list(label="dust")
+  control.transition_dating_dust=list(label="dust",sync=FALSE)
   agetemp_dust = events_depth_to_age(temp_d18O,control.transition_dating_dust)

### 2022-09-30 R/default_arguments.R

@@ -176,6 +176,9 @@ synchronization.default <- function(){
 #'    should be computed. Default \code{TRUE}.
 #'    \item \code{nsims} Integer giving how many Monte Carlo samples should be used to
 #'    compute transition end point \code{t1} if \code{compute.t1} is \code{TRUE}.
+#'    \item \code{rescale.y.factor} If proxies (y-axis) should be rescaled, this can be done using
+#'    this command. \code{y_new = y_old*rescale.y.factor}.
+#'    \item \code{imp.fit} Boolean indicaating whether initial values for \code{inla} be computed using \code{optim}.
 #'    \item \code{h} Step length of optimization procedure in the \code{inla}-program.
 #'    Default value is \code{0.01}.
 #'    \item \code{ncores} Integer giving the number of cores used in the \code{inla}-program.
@@ -204,6 +207,8 @@ control.linramp.default <- function(){
       opt.params=NULL, #function will recognize this
       compute.t1=TRUE,
       nsims=30000,
+      rescale.y.factor=FALSE,
+      imp.fit=TRUE,
       h=0.01,
       ncores=1,
       silent=FALSE,

### 2022-09-30 R/linrampfitter.R

@@ -93,7 +93,10 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(is.null(object$simulation$age) && is.null(object$simulation$age_sync))
     stop("Chronologies not found. Run 'bremla_synchronized_simulation' or 'bremla_chronology_simulation' first!")
   interval = control.linramp$interval
-
+  if(is.null(interval)){
+    interval = 1:length(control.linramp$proxy)
+    control.linramp$interval.unit = "index"
+  }
 
 
   if(control.linramp$interval.unit == "depth"){
@@ -106,6 +109,7 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
     interval = range(interval)[1]:range(interval)[2]
   }
 
+  n=length(interval)
   if(print.progress) cat("Initializing linear ramp fit using INLA.\n",sep="")
   df_event = data.frame(xx=rev(object$data$depth[interval]),
                         yy=rev(control.linramp$proxy[interval])) #reverse: Want depth axis representing old->new
@@ -118,6 +122,10 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   }
 
   n=length(interval)
+
+  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
+
+
   df0=data.frame(y=df_event$yy,x=df_event$xx)
 
   t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
@@ -127,7 +135,6 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(print.progress) cat("Using 'optim' to find initial positions for hyperparameters in INLA.\n",sep="")
   ## for stability the x-axis is transformed to values ranging from 1:n.
 
-  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
   df0$time = timepoints
 
   ## Finding initial values in INLA optimization procedure by first using a simple optimization
@@ -281,6 +288,354 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   object$linramp$.args = list(nsims=control.linramp$nsims,
                               depth.reference=control.linramp$depth.reference,
                               label=control.linramp$depth.label)
+  object$time$linramp=time.total
+
[…]

### 2022-09-30 R/plot.bremla.R

@@ -480,25 +480,31 @@ plot.bremla = function(x,
     }else{
       plot.label=x$linramp$.args$label
     }
-    plot(xval,x$linramp$data$y,type="l",lwd=1.25,col="gray",xlim=xlim,ylim=yrange,xlab="Depth (m)",ylab=expression(paste(delta^18, "O (permil)")),main=plot.label)
+
+    ybottom = min(x$linramp$data$y)
+    ytop = min(x$linramp$linrampfit$q0.025)
+    epsilon = (ytop-ybottom)*0.3
+    if(plot.linramp$show.t0 && !is.null(x$linramp$param)){
+
+      margt0=x$linramp$param$t0$marg.t0
+      #normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
+      normt0.y = (margt0[,2]-min(margt0[,2]))/diff(range(margt0[,2]))*(ytop-ybottom)+ybottom-epsilon
+    }
+    if(plot.linramp$show.t0 && !is.null(x$linramp$param$t1)){
+      # ybottom = min(x$linramp$data$y)
+      # ytop = min(x$linramp$linrampfit$q0.025)
+      margt1 = x$linramp$param$t1$marginal
+      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom-epsilon
+    }
+    yrange2=range(yrange,normt0.y,normt1.y)
+    plot(xval,x$linramp$data$y,type="l",lwd=1.25,col="gray",xlim=xlim,ylim=yrange2,xlab="Depth (m)",
+         ylab=expression(paste(delta^18, "O (permil)")),main=plot.label)
     lines(xval,x$linramp$linrampfit$q0.025,col="red",lwd=2)
     lines(xval,x$linramp$linrampfit$q0.975,col="red",lwd=2)
     lines(xval,x$linramp$linrampfit$mean,col="black",lwd=2)
 
-    if(plot.linramp$show.t0 && !is.null(x$linramp$param)){
-      ybottom = min(x$linramp$data$y)
-      ytop = min(x$linramp$linrampfit$q0.025)
-      margt0=x$linramp$param$t0$marg.t0
-      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
[…]

### 2022-09-30 R/rgeneric.uneven.AR1.R

@@ -153,6 +153,7 @@ rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to defin
       tslutt=get("tslutt",envir)
       tstart=get("tstart",envir)
       ystart=get("ystart",envir)
+      # rescale.y=get("rescale.y",envir)
     }
     params = interpret.theta()
 
@@ -160,10 +161,17 @@ rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to defin
 
     lprior = dnorm(theta[1],mean=round(0.5*(tslutt+tstart)),sd=50,log=TRUE) #t0
     lprior = lprior + dgamma(exp(theta[2]),shape=1.0,rate=0.02,log=TRUE) + theta[2] #dt
-    lprior = lprior + dnorm(theta[3],mean=ystart,sd=5,log=TRUE) #y0
-    lprior = lprior + dnorm(theta[4],mean=0,sd=10.0,log=TRUE) #dy
-    lprior = lprior + dgamma(exp(theta[5]),2.5,rate = 0.15,log=TRUE) + theta[6] #tau/rho
-    lprior = lprior + dgamma(exp(theta[6]),2,rate = 0.15,log=TRUE) + theta[6] #sigma/kappa
+    # if(rescale.y){
+    #   lprior = lprior + dnorm(theta[3],mean=ystart,sd=0.025,log=TRUE) #y0
+    #   lprior = lprior + dnorm(theta[4],mean=1-ystart,sd=0.01,log=TRUE) #dy
+    #   lprior = lprior + dgamma(exp(theta[5]),1,rate = 3,log=TRUE) + theta[6] #tau/rho
+    #   lprior = lprior + dgamma(exp(theta[6]),0.1,rate = 1,log=TRUE) + theta[6] #sigma/kappa
+    # }else{
+      lprior = lprior + dnorm(theta[3],mean=ystart,sd=5,log=TRUE) #y0
+      lprior = lprior + dnorm(theta[4],mean=0,sd=10.0,log=TRUE) #dy
+      lprior = lprior + dgamma(exp(theta[5]),2.5,rate = 0.15,log=TRUE) + theta[6] #tau/rho
+      lprior = lprior + dgamma(exp(theta[6]),2,rate = 0.15,log=TRUE) + theta[6] #sigma/kappa
+    # }
     return (lprior)
   }
 

### 2022-09-30 inst/Examples/Paper_results/Paper_results_1.Rmd
Minor changes in linrampfitter.R
if NA is inserted into opt.params, then replace with default value

@@
 depthlist_d18O = c()
-do.plot.rampfit = FALSE
+do.plot.rampfit = TRUE
 do.plot.agehist = FALSE
@@
     {
+      ybottom = min(temp_d18O$linramp$data$y)
+      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
+      epsilon = (ytop-ybottom)*0.3
+
+      margt0=temp_d18O$linramp$param$t0$marg.t0
+      normt0.y = (margt0[,2]-min(margt0[,2]))/diff(range(margt0[,2]))*(ytop-ybottom)+ybottom-epsilon
+      margt1 = temp_d18O$linramp$param$t1$marginal
+      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom-epsilon
+
+      yrange = range(temp_d18O$linramp$data$y,temp_d18O$linramp$linrampfit$q0.025,
+                     temp_d18O$linramp$linrampfit$q0.975,
+                     normt0.y,normt1.y)
+
       plot(depth[interval_d18O],proxy[interval_d18O],type="l",col="gray",
           xlim=rev(range(depth[interval_d18O])), xlab="Depth (m)",ylab=expression(paste(delta^18,"O (permil)")),
-          main=paste0(eventnumber," (d18O): ", event_intervals[eventnumber,2]))
+          main=paste0(eventnumber," (d18O): ", event_intervals[eventnumber,2]),
+          ylim=yrange)
       lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$mean)
@@
 
-      ybottom = min(temp_d18O$linramp$data$y)
-      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
-      margt0=temp_d18O$linramp$param$t0$marg.t0
-      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
       lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
-      ybottom = min(temp_d18O$linramp$data$y)
-      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
-      margt1 = temp_d18O$linramp$param$t1$marginal
-      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
       lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
+      # ybottom = min(temp_d18O$linramp$data$y)
+      # ytop = min(temp_d18O$linramp$linrampfit$q0.025)
+      # margt0=temp_d18O$linramp$param$t0$marg.t0
+      # normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
+      # lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
+      # ybottom = min(temp_d18O$linramp$data$y)
+      # ytop = min(temp_d18O$linramp$linrampfit$q0.025)
+      # margt1 = temp_d18O$linramp$param$t1$marginal
+      # normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
+      # lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
     }
@@
     {
-      plot(temp_dust$linramp$data$x,temp_dust$linramp$data$y,type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
-          main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
-      # plot(depth[interval_d18O],logdust[interval_d18O],type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
-      #     main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
-      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$mean)
-      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.025,col="red")
-      lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.975,col="red")
+      ybottom = min(temp_dust$linramp$data$y)
+      ytop = min(temp_dust$linramp$linrampfit$q0.025)
+      epsilon = (ytop-ybottom)*0.3
+
+      margt0=temp_dust$linramp$param$t0$marg.t0
+      normt0.y = (margt0[,2]-min(margt0[,2]))/diff(range(margt0[,2]))*(ytop-ybottom)+ybottom-epsilon
+      margt1 = temp_dust$linramp$param$t1$marginal
+      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom-epsilon
+
+      yrange = range(temp_dust$linramp$data$y,temp_dust$linramp$linrampfit$q0.025,
+                     temp_dust$linramp$linrampfit$q0.975,
+                     normt0.y,normt1.y)
+
+      plot(temp_dust$linramp$data$x,temp_dust$linramp$data$y,type="l",col="gray",
+          xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
+          main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]),
+          ylim=yrange)
+      lines(temp_d18O$linramp$data$x,temp_dust$linramp$linrampfit$mean)
+      lines(temp_d18O$linramp$data$x,temp_dust$linramp$linrampfit$q0.025,col="red")
+      lines(temp_d18O$linramp$data$x,temp_dust$linramp$linrampfit$q0.975,col="red")
       abline(v=rasmussen_depths[eventnumber],lty=3)
 
-      ybottom = min(temp_dust$linramp$data$y)
-      ytop = min(temp_dust$linramp$linrampfit$q0.025)
-      margt0=temp_dust$linramp$param$t0$marg.t0
-      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
       lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
-      ybottom = min(temp_dust$linramp$data$y)
-      ytop = min(temp_dust$linramp$linrampfit$q0.025)
-      margt1 = temp_dust$linramp$param$t1$marginal
-      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
       lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
+
+      # plot(temp_dust$linramp$data$x,temp_dust$linramp$data$y,type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
+      #     main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
+      # # plot(depth[interval_d18O],logdust[interval_d18O],type="l",col="gray",xlim=rev(range(depth[interval_dust])), xlab="Depth (m)",ylab=expression(paste("log(Ca)","")),
+      # #     main=paste0(eventnumber," (logdust): ", event_intervals[eventnumber,2]))
+      # lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$mean)
+      # lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.025,col="red")
+      # lines(temp_dust$linramp$data$x,temp_dust$linramp$linrampfit$q0.975,col="red")
+      # abline(v=rasmussen_depths[eventnumber],lty=3)
+      # 
+      # ybottom = min(temp_dust$linramp$data$y)
+      # ytop = min(temp_dust$linramp$linrampfit$q0.025)
+      # margt0=temp_dust$linramp$param$t0$marg.t0
+      # normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
+      # lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
+      # ybottom = min(temp_dust$linramp$data$y)
+      # ytop = min(temp_dust$linramp$linrampfit$q0.025)
+      # margt1 = temp_dust$linramp$param$t1$marginal
+      # normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
+      # lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
     }

### 2022-10-10 README.Rmd
Adding Github action for R CMD check when INLA is deactivated
also added some other features, including user-specification of priors for linear ramp model (experimental)

@@
 [![R-CMD-check](https://github.com/eirikmn/bremla/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/eirikmn/bremla/actions/workflows/R-CMD-check.yaml)
+[![License: GPL v2](https://img.shields.io/badge/License-GPL_v2-blue.svg)](https://www.gnu.org/licenses/old-licenses/gpl-2.0.en.html)
 <!-- badges: end -->
@@
 
-The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al.
+The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](https://www.iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al.
 (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485).
@@
 
-gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=x$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
+gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=results$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
                    upper=results$simulation$summary_sync$upper[tie_indexes]-ageref_tie),

### 2022-10-10 R/bremla.R

@@ -32,6 +32,7 @@
 #' @keywords bremla
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' ## Simulation example
 #' require(stats)
 #' set.seed(1)
@@ -83,8 +84,8 @@
 #' summary(object)
 #' plot(object)
 #'
-#'
-#'
+#' }
+#' if(inlaloader()){
 #' ### Real data example ###
 #' require(stringr)
 #' data("event_intervals")
@@ -132,6 +133,7 @@
 #'                    print.progress=TRUE )
 #'   summary(object)
 #'   plot(object)
+#'   }
 #' }
 #' @export
 #' @import matrixStats

### 2022-10-10 R/bremla_biased_chronologies.R

@@ -15,6 +15,7 @@
 #' @keywords bremla bias
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -51,6 +52,7 @@
 #'              print.progress=TRUE)
 #' summary(object)
 #' plot(object)
+#' }
 #' }
 #' @export
 #' @importFrom stats runif

### 2022-10-10 R/bremla_chronology_simulation.R

@@ -16,6 +16,7 @@
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -50,11 +51,10 @@
 #' object = bremla_chronology_simulation(object, print.progress=TRUE)
 #' summary(object)
 #' plot(object)
+#' }
 #'}
 #'
 #' @export
-#' @import INLA
-#' @importFrom INLA inla.hyperpar.sample inla.tmarginal inla.zmarginal inla.ar.pacf2phi
 #' @importFrom stats acf arima arima.sim as.formula rnorm
 #' @importFrom parallel detectCores
 bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE){
@@ -92,18 +92,18 @@ bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE
 
 
     if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]))
     }else if (tolower(noise) %in% c(1,"ar1","ar(1)")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]), phi=hypersamples[,2])
 
     }else if (tolower(noise) %in% c(2,"ar2","ar(2)")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       p=2
-      hypersamplesar2 = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamplesar2 = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       phii = hypersamplesar2[, 2L:(2L+(p-1L))]
-      phis = apply(phii, 1L, inla.ar.pacf2phi)
+      phis = apply(phii, 1L, INLA::inla.ar.pacf2phi)
[…]

### 2022-10-10 R/bremla_modelfitter.R

@@ -15,6 +15,7 @@
 #' @keywords bremla fitting
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -45,8 +46,8 @@
 #' summary(object)
 #' plot(object)
 #' }
+#' }
 #' @export
-#' @importFrom INLA inla inla.tmarginal inla.zmarginal inla.ar.pacf2phi
 #' @importFrom stats acf arima density lm sd
 
 bremla_modelfitter = function(object, control.fit,
@@ -162,7 +163,7 @@ bremla_modelfitter = function(object, control.fit,
     my.control.fixed$prec=1
 
     ## fit using INLA
-    inlafit = inla(object$formula, family="gaussian",data=object$data,
+    inlafit = INLA::inla(object$formula, family="gaussian",data=object$data,
                    control.family=list(hyper=list(prec=list(initial=12, fixed=TRUE))) ,
                    #control.fixed=my.control.fixed,
                    num.threads = object$.args$control.fit$ncores,
@@ -185,22 +186,22 @@ bremla_modelfitter = function(object, control.fit,
     }
 
     ## extract posteriors for hyperparameters
-    posterior_sigma = inla.tmarginal(function(x)1/sqrt(x),inlafit$marginals.hyperpar$`Precision for idy`); zmarg_sigma=inla.zmarginal(posterior_sigma,silent=TRUE)
+    posterior_sigma = INLA::inla.tmarginal(function(x)1/sqrt(x),inlafit$marginals.hyperpar$`Precision for idy`); zmarg_sigma=INLA::inla.zmarginal(posterior_sigma,silent=TRUE)
     object$fitting$inla$hyperparameters = list(posteriors=list(sigma_epsilon=posterior_sigma))
     object$fitting$inla$hyperparameters$results$sigma_epsilon = zmarg_sigma
     if(tolower(object$.args$control.fit$noise)%in% c(1,"ar1","ar(1)") ){
-      posterior_phi = inlafit$marginals.hyperpar$`Rho for idy`; zmarg_phi = inla.zmarginal(posterior_phi,silent=TRUE)
+      posterior_phi = inlafit$marginals.hyperpar$`Rho for idy`; zmarg_phi = INLA::inla.zmarginal(posterior_phi,silent=TRUE)
[…]

### 2022-10-10 R/bremla_synchronized_simulation.R

@@ -14,6 +14,7 @@
 #' computed or related to the simulation procedure in \code{object\$simulation}.
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -55,6 +56,7 @@
 #' object = bremla_synchronized_simulation(object, print.progress=TRUE)
 #' summary(object)
 #' plot(object)
+#' }
 #' }
 #' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
 #' @seealso \code{\link{bremla_chronology_simulation}} \code{\link{bremla}}
@@ -126,18 +128,18 @@ time.start = Sys.time()
 
 
     if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]))
     }else if (tolower(noise) %in% c(1,"ar1","ar(1)")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]), phi=hypersamples[,2])
 
     }else if (tolower(noise) %in% c(2,"ar2","ar(2)")){
-      hypersamples = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       p=2
-      hypersamplesar2 = inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      hypersamplesar2 = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       phii = hypersamplesar2[, 2L:(2L+(p-1L))]
-      phis = apply(phii, 1L, inla.ar.pacf2phi)
+      phis = apply(phii, 1L, INLA::inla.ar.pacf2phi)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]),phi1=phis[1,],phi2=phis[2,])
     }
 
@@ -160,7 +162,7 @@ time.start = Sys.time()
   # }
   timecoef.start = Sys.time()
[…]

### 2022-10-10 R/default_arguments.R

@@ -64,6 +64,7 @@ control.fit.default <- function(){
     noise="ar1",
     method="inla",
     verbose=FALSE,
+    log.theta.prior=NULL,
     ncores=1,
     transform="identity"
     )
@@ -77,8 +78,8 @@ control.fit.default <- function(){
 #'   \item{\code{synchronized} }{Boolean indicating whether or not synchronization is to be performed.
 #'   Default value is \code{TRUE}, but this is set to \code{FALSE} if supplied tie-points are not found.}
 #'   \item{\code{nsims} }{Integer giving how many simulations should be generated. Default value is \code{10000}.}
-#'   \item{\code{ncores} }{Integer which sets the number of cores to be used in \code{\link{inla.posterior.sample}.}
-#'   Default value is \code{2}.}
+#'   \item{\code{ncores} }{Integer which sets the number of cores to be used to sample from
+#'   the joint posterior obtained by INLA. Default value is \code{2}.}
 #'   \item{\code{store.everything} }{Boolean indicating whether or not to store superfluous information
 #'   such as the simulated mean vectors (from fixed effects). Default \code{FALSE}.}
 #'   \item{\code{summary} }{List object with two items
@@ -182,7 +183,11 @@ synchronization.default <- function(){
 #'    compute transition end point \code{t1} if \code{compute.t1} is \code{TRUE}.
 #'    \item \code{rescale.y.factor} If proxies (y-axis) should be rescaled, this can be done using
 #'    this command. \code{y_new = y_old*rescale.y.factor}.
-#'    \item \code{imp.fit} Boolean indicaating whether initial values for \code{inla} be computed using \code{optim}.
+#'    \item \code{imp.fit} Boolean indicating whether initial values for \code{inla} be computed using \code{optim}.
+#'   \item{\code{log.theta.prior}}{Function argument which returns the logarithmic value of the joint
+#'   prior of the rescaled hyperparameters \code{theta} used in the linear ramp model.
+#'   If \code{NULL} (default) then default priors  will be assumed (see code in
+#'   \code{rgeneric.r} file).}
 #'    \item \code{h} Step length of optimization procedure in the \code{inla}-program.
 #'    Default value is \code{0.01}.
 #'    \item \code{ncores} Integer giving the number of cores used in the \code{inla}-program.
@@ -213,6 +218,7 @@ control.linramp.default <- function(){
       nsims=30000,
       rescale.y.factor=1,
[…]

### 2022-10-10 R/events_depth_to_age.R

@@ -14,6 +14,7 @@
 #' @keywords bremla dating transition
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -65,11 +66,9 @@
 #' object = events_depth_to_age(object,print.progress=TRUE)
 #' summary(object)
 #' plot(object)
-#'
+#'}
 #' }
 #' @export
-#' @import INLA
-#' @importFrom INLA inla.rgeneric.define inla.emarginal inla.smarginal
 #' @importFrom stats optim
 #'
 events_depth_to_age = function(object, control.transition_dating,print.progress=FALSE){
@@ -125,7 +124,7 @@ events_depth_to_age = function(object, control.transition_dating,print.progress=
               object$.args$synchronization$nsims)
   if(nsims > ncol(object$simulation$age))
     stop(paste0("Not enough samples from age-depth model (",ncol(object$simulation$age),", ",nsims, "needed)"))
-  z.sample = inla.rmarginal(nsims,object$linramp$param$t0$marg.t0)
+  z.sample = INLA::inla.rmarginal(nsims,object$linramp$param$t0$marg.t0)
 
   if(is.null(object$simulation)){
     warning("Cannot find simulations from age-depth model in 'agedepthmodel'!")
@@ -161,8 +160,8 @@ events_depth_to_age = function(object, control.transition_dating,print.progress=
 
 
 
-  dens = density(ysamps); Y0marg= cbind(dens$x,dens$y); colnames(Y0marg) = c("x","y"); Y0marg = inla.smarginal(Y0marg)
-  z.Y0 = inla.zmarginal(Y0marg,silent=TRUE)
+  dens = density(ysamps); Y0marg= cbind(dens$x,dens$y); colnames(Y0marg) = c("x","y"); Y0marg = INLA::inla.smarginal(Y0marg)
+  z.Y0 = INLA::inla.zmarginal(Y0marg,silent=TRUE)
 
   object$event_dating = list(samples = ysamps, mean = z.Y0$mean,sd=z.Y0$sd,q0.025=z.Y0$quant0.025,q0.5=z.Y0$quant0.5,q0.975=z.Y0$quant0.975)
 

### 2022-10-10 R/helpfulfunctions.R

@@ -127,6 +127,7 @@ which.index = function(events, record,suppress=TRUE){ ## Finds which indices of 
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -162,11 +163,10 @@ which.index = function(events, record,suppress=TRUE){ ## Finds which indices of 
 #' object = bremla_simulationsummarizer(object,sync=FALSE,print.progress=TRUE)
 #' summary(object)
 #' plot(object)
-#'
+#' }
 #' }
 #' @export
-#' @importFrom matrixStats rowMeans2 rowSds rowMedians
-#' @importFrom INLA inla.hpdmarginal
+
 #' @importFrom stats median
 bremla_simulationsummarizer = function(object,sync=TRUE,print.progress=FALSE){
 
@@ -204,8 +204,8 @@ bremla_simulationsummarizer = function(object,sync=TRUE,print.progress=FALSE){
         dens = density(samples[i,])
 
 
-        lower[i] = inla.hpdmarginal(0.95,dens)[1]
-        upper[i] = inla.hpdmarginal(0.95,dens)[2]
+        lower[i] = INLA::inla.hpdmarginal(0.95,dens)[1]
+        upper[i] = INLA::inla.hpdmarginal(0.95,dens)[2]
       }
     }
   }else{
@@ -343,8 +343,10 @@ skewsampler = function(n,mode=0,sdL=1,sdU=1,log=FALSE,plothist=list(compute=FALS
 #' @param x.ref numeric. Gives reference value on the x-axis when distributions are plotted.
 #' @return returns a list containing the pdfs for each tie-point.
 #' @examples
+#' if(inlaloader()){
 #' adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050))
 #' plot(adolphipdfs$tie1,type="l",xlab="Time (yb2k)",ylab="Density",main="Tie-point #1")
+#' }
 #' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
 #' @references Adolphi, F., Bronk Ramsey, C., Erhardt, T., Edwards, R. L., Cheng, H., Turney, C. S. M., Cooper, A., Svensson, A., Rasmussen, S. O., Fischer, H., and Muscheler, R. (2018).
[…]

### 2022-10-10 R/linrampfitter.R

@@ -16,6 +16,7 @@
 #' @keywords bremla linear_ramp
 #' @examples
 #'\donttest{
+#'if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -64,12 +65,11 @@
 #' object = linrampfitter(object,print.progress=TRUE)
 #' summary(object)
 #' plot(object)
-#'
+#'}
 #'}
 #'
 #' @export
-#' @import INLA numDeriv
-#' @importFrom INLA inla.rgeneric.define inla.emarginal inla.smarginal
+#' @import numDeriv
 #' @importFrom stats optim
 #' @importFrom numDeriv grad
 #'
@@ -176,10 +176,15 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(print.progress) cat("Fitting linear ramp model in INLA using rgeneric model specification...\n",sep="")
   ## creating linear ramp INLA model using rgeneric framework. Requires further specification, see "rgeneric.uneven" function
   time.startinla = Sys.time()
-  model.rgeneric = inla.rgeneric.define(rgeneric.uneven.AR1,n=n,tstart=timepoints[1],tslutt=timepoints[n],ystart=y[1],timepoints = timepoints)
+  model.rgeneric = INLA::inla.rgeneric.define(rgeneric.uneven.AR1,n=n,
+                                              tstart=timepoints[1],
+                                              tslutt=timepoints[n],
+                                              ystart=y[1],
+                                              timepoints = timepoints,
+                                              log.theta.prior=control.linramp$log.theta.prior)
   formula = y ~ -1+ f(idx, model=model.rgeneric)
 
-  r = inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=as.integer(df0$time)),
+  r = INLA::inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=as.integer(df0$time)),
            control.mode=list(theta=init,
[…]

### 2022-10-10 R/plot.bremla.R

@@ -25,6 +25,7 @@
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -59,11 +60,10 @@
 #' object = bremla_chronology_simulation(object, print.progress=TRUE)
 #' summary(object)
 #' plot(object)
-#'
+#'}
 #' }
 #'
 #' @export
-#' @importFrom INLA inla inla.tmarginal inla.zmarginal inla.ar.pacf2phi
 #' @importFrom grDevices dev.cur dev.new dev.off
 #' @importFrom graphics abline hist legend lines par
 #' @importFrom stats qqline qqnorm

### 2022-10-10 R/print.bremla.R

@@ -12,6 +12,7 @@
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -46,6 +47,7 @@
 #' object = bremla_chronology_simulation(object, print.progress=TRUE)
 #' summary(object)
 #' plot(object)
+#' }
 #' }
 #' @export
 #' @method print bremla

### 2022-10-10 R/rgeneric.uneven.AR1.R

@@ -10,17 +10,18 @@
 #'
 #' @return When used an an input argument for \code{inla.rgeneric.define} this will return a model eligible to be used within the R-INLA framework (see example).
 #' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
-#' @seealso \code{\link{inla},\link{linrampfitter}}
+#' @seealso \code{\link{linrampfitter}}
 #' @keywords bremla rgeneric inla
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' set.seed(1)
 #' n=300
 #' timepoints = 1:n
 #' require(stats)
-#' require(INLA)
 #' require(numDeriv)
+#' require(INLA)
 #' sigma=1
 #' noise = stats::arima.sim(model=list(ar=c(0.2)),n=n,sd=sqrt(1-0.2^2))*sigma
 #'
@@ -63,7 +64,8 @@
 #'                    tstart=timepoints[1],
 #'                    tslutt=timepoints[n],
 #'                    ystart=y[1],
-#'                    timepoints = timepoints)
+#'                    timepoints = timepoints,
+#'                    log.theta.prior=NULL)
 #' formula = y ~ -1+ f(idx, model=model.rgeneric)
 #'
 #' result = inla(formula,family="gaussian", data=data.frame(y=y,idx=as.integer(1:n)),
@@ -83,7 +85,7 @@
 #' tau.mean = inla.emarginal(function(x)exp(x),result$marginals.hyperpar$`Theta5 for idx`)
 #' plot(timepoints,y,type="l",col="gray",xlab="Time",ylab="Observation")
 #' lines(timepoints,linramp(timepoints,t0=t0.mean,dt=dt.mean,y0=y0.mean,dy=dy.mean))
-#'
+#'}
 #'}
 #'
 #'
@@ -91,7 +93,6 @@
 #' @import Matrix
 #' @importFrom stats dgamma dnorm
 #' @importFrom Matrix sparseMatrix
-#' @importFrom stats
 
 rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to define the linear ramp model
   cmd = c("graph", "Q","mu", "initial", "log.norm.const", "log.prior", "quit"),
@@ -190,25 +191,32 @@ rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to defin
[…]

### 2022-10-10 R/summary.bremla.R

@@ -12,6 +12,7 @@
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -47,7 +48,7 @@
 #' summary(object)
 #' plot(object)
 #' }
-#'
+#'}
 #' @export
 #' @method summary bremla
 summary.bremla = function(object,
@@ -268,6 +269,7 @@ return(ut)
 #'
 #' @examples
 #' \donttest{
+#' if(inlaloader()){
 #' require(stats)
 #' set.seed(1)
 #' n <- 1000
@@ -302,6 +304,7 @@ return(ut)
 #' object = bremla_chronology_simulation(object, print.progress=TRUE)
 #' summary_object = summary(object)
 #' print(summary_object)
+#' }
 #' }
 #'
 #' @export

### 2022-10-25 README.Rmd
some bug fixes
also added more plots for the markdown file to reproduce results from paper

@@
 
-This Repository contains the Bremla package for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for
+This repository contains the bremla package (**B**ayesian **Re**gression **M**odeling of **L**ayer-counted **a**rchives) for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for
 
@@
 Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000.
-More if synchronized chronologies is also computed.
+More if both synchronized and unsynchronized chronologies computed.
 This will require a significant amount of memory.
@@
 ## Example
-This is a real data example which produces 5000 simulations from a synchronized time scale.
+This is a real data example for the GICC05 chronology of the NGRIP ice core record which produces 5000 simulations from a time scale synchronized using the Adolphi tie-points.
 ```{r example, message=FALSE, warning=FALSE, include=TRUE}

### 2022-10-25 inst/Examples/Paper_results/Paper_results_2.Rmd
some bug fixes
also added more plots for the markdown file to reproduce results from paper

@@
 
+## Illustration example
+
+The data used to generate the illustrative examples in the paper is as follows.
+
+```{r illustrationexample}
+rm(list=ls())
+require(stats)
+library(ggplot2)
+set.seed(123)
+n=800
+y0 = 0
+z0 = 0
+phi=0.8; sigma=1.2
+depth = z0+seq(from=1,to=n,length.out=n)
+
+covariate = arima.sim(model=list(ar=c(0.99)),n=n)
+noise = sigma*arima.sim(model=list(ar=c(phi)),n=n,sd=sqrt(1-phi^2))
+
+dy = 10 + depth*5/n+covariate+noise
+y = y0+cumsum(dy)
+
+data=data.frame(age=c(y0,y),depth=c(z0,depth),dy=c(0,as.numeric(dy)),covariate=c(0,covariate))
+#data=rbind(c(y0,z0,0,0),df)
+formula=dy~ 1 + depth + covariate 
+nsims = 10000
+
+# Unsynchronized
+res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
+             control.fit=list(noise="ar1"),
+             x.label="Depth",
+             y.label="Ensemble age - observed age",
+             control.sim = list(synchronized=FALSE,
+                                nsims=nsims))
+
+library(ggplot2)
+orichron = res$data$age
+ggd = data.frame(depth=res$data$depth,
+                 mean = res$simulation$summary$mean-orichron,
+                 lower = res$simulation$summary$lower-orichron,
+                 upper = res$simulation$summary$upper-orichron)
+ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)
+# plot(res)
+#ggsave("illustrative-unsync.eps",device=cairo_ps,width=4800,height=2000,units="px")
+
+# Fixed tie-points
+ntie = 3
+tiesims = matrix(NA,nrow=nsims,ncol=ntie)
+tiedepths = c(100,300,600)
+tieoffset = c(-50,0,100)
+#tieage = c(450,3000,8500)
+tieage = y[tiedepths]+tieoffset
+
+for(i in 1:ntie){
+  tiesims[,i] = rep(tieage[i],nsims)
+}
+
+res2 = tiepointsimmer(res,synchronization = list(locations=tiedepths,locations_unit="depth",
+                       samples = tiesims)
+               )
+
+res2 = bremla_synchronized_simulation(res2,control.sim=list(synchronized=TRUE))
+
+ggd = data.frame(depth=res2$data$depth,
+                 mean = res2$simulation$summary_sync$mean-orichron,
+                 lower = res2$simulation$summary_sync$lower-orichron,
+                 upper = res2$simulation$summary_sync$upper-orichron)
+
+oritie = orichron[tiedepths]
+ggd2 = data.frame(depth=res2$tie_points$locations,
+                  mean=res2$simulation$summary_sync$mean[tiedepths]-oritie)
+
+ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)+
+  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="magenta",size=1.2)
+#plot(res)
+#ggsave("illustrative-fixedsync.eps",device=cairo_ps,width=4800,height=2000,units="px")
+
+tieoffset = c(-50,0,100)
+tiemeans = y[tiedepths]+tieoffset
+#tiemeans = c(450,3000,8500)
+tiesds = c(10,20,30)
+res3 = tiepointsimmer(res,synchronization = list(locations=tiedepths,
+                                                locations_unit="depth",
+                                                method="normal",
+                                                params=list(mean=tiemeans,
+                                                            sd=tiesds))
+               )
+res3 = bremla_synchronized_simulation(res3,control.sim=list(synchronized=TRUE))
+ggd = data.frame(depth=res3$data$depth,
+                 mean = res3$simulation$summary_sync$mean-orichron,
+                 lower = res3$simulation$summary_sync$lower-orichron,
+                 upper = res3$simulation$summary_sync$upper-orichron)
+
+oritie = orichron[tiedepths]
+ggd2 = data.frame(depth=res3$tie_points$locations,
+                  lower=res3$simulation$summary_sync$lower[tiedepths]-oritie,
+                  upper=res3$simulation$summary_sync$upper[tiedepths]-oritie)
+
+ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)+
+  geom_segment(data=ggd2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
+               col="magenta",size=0.7)
+
+#plot(res3)
+#ggsave("illustrative-sync.eps",device=cairo_ps,width=4800,height=2000,units="px")
+
+```
+
 ## Simulation example

### 2022-10-25 R/bremla.R

@@ -8,6 +8,8 @@
 #' The first row is used for initializing 'age' and 'depth', hence the rest can be set to NA.
 #' Covariates related to the events (psi) can be filled in automatically by specifying the degree and events in \code{events} argument.
 #' @param reference.label Character label of reference timescale. Used in \code{\link{plot},\link{summary}}.
+#' @param x.label Character label for the x-axis (depth).
+#' @param y.label Character label for the y-axis (age).
 #' @param nsims Number of chronologies to be simulated.
 #' @param events List object describing locations, degree and other informations about
 #' the specific climatic periods used in the predictor. If used must include at least
@@ -138,6 +140,8 @@
 #' @export
 #' @import matrixStats
 bremla <- function(formula,data,reference.label=NULL,
+                            x.label=NULL,
+                            y.label=NULL,
                             nsims=10000,
                             events=NULL,
                             synchronization=NULL,
@@ -157,6 +161,8 @@ bremla <- function(formula,data,reference.label=NULL,
   #prepare bremla object by formatting dataset, writing formulastrings etc
   object = bremla_prepare(formula, data, ##data must include 'depth' and 'age' in data or input
                           reference.label=reference.label,
+                          x.label = x.label,
+                          y.label = y.label,
                           nsims=nsims,
                           events=events,
                           synchronization=synchronization,
@@ -181,17 +187,19 @@ bremla <- function(formula,data,reference.label=NULL,
 
   }
 
+  if(!is.null(control.sim)){
+    if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
+      control.sim$nsims=nsims
+      #produce samples from the chronologies
+      object = bremla_chronology_simulation(object, control.sim=control.sim,
+                                            print.progress=print.progress)
 
-  if(nsims>0 && control.sim$synchronized %in% c(FALSE,2)){
-    control.sim$nsims=nsims
-    #produce samples from the chronologies
-    object = bremla_chronology_simulation(object, control.sim=control.sim,
-                                          print.progress=print.progress)
+      #compute posterior marginal mean, quantiles and other summary statistics
+      #object = bremla_simulationsummarizer(object,CI.type=CI.type,sync=FALSE,print.progress=print.progress)
 
-    #compute posterior marginal mean, quantiles and other summary statistics
[…]

### 2022-10-25 R/bremla_modelfitter.R

@@ -68,7 +68,41 @@ bremla_modelfitter = function(object, control.fit,
     }else{
         stop("Could not find 'control.fit'. Stopping.")
     }
+  }else{
+    if(!is.null(control.fit$noise)){
+      if(control.fit$noise != object$.args$control.fit$noise){
+        object = bremla_prepare(formula=object$.args$formula.input,
+                                data=object$.args$data.input,
+                                nsims=object$.args$nsims,
+                                reference.label = object$.args$reference.label,
+                                x.label = object$.args$x.label,
+                                y.label = object$.args$y.label,
+                                events = object$.args$events,
+                                synchronization=object$.args$synchronization,
+                                control.fit=control.fit,
+                                control.sim = object$.args$control.sim,
+                                control.linramp = object$.args$control.linramp,
+                                control.transition_dating = object$.args$control.transition_dating,
+                                control.bias=object$.args$control.bias)
+      }
+    }
+    # }else{
+    #   object = bremla_prepare(formula=object$.args$formula.input,
+    #                           data=object$.args$data.input,
+    #                           nsims=object$.args$nsims,
+    #                           reference.label = object$.args$reference.label,
+    #                           x.label = object$.args$x.label,
+    #                           y.label = object$.args$y.label,
+    #                           events = object$.args$events,
+    #                           synchronization=object$.args$synchronization,
+    #                           control.fit=control.fit,
+    #                           control.sim = object$.args$control.sim,
+    #                           control.linramp = object$.args$control.linramp,
+    #                           control.transition_dating = object$.args$control.transition_dating,
+    #                           control.bias=object$.args$control.bias)
+    # }
   }
+
 
 
 
@@ -85,7 +119,7 @@ bremla_modelfitter = function(object, control.fit,
   time.start = Sys.time()
 
   formula = object$formula
-  if(print.progress) cat("Performing least squares fit...",sep="")
+    if(print.progress) cat("Performing least squares fit...",sep="")
 
   fit = lm(object$.internal$formula.ls,object$data) #fits model using least squares
 

### 2022-10-25 R/bremla_prepare.R

@@ -11,6 +11,8 @@
 #' covariates (phi) can be filled out using 'events$fill_data'.
 #' @param nsims Number of chronologies to be simulated.
 #' @param reference.label Character label of reference timescale. Used in \code{\link{plot},\link{summary}}.
+#' @param x.label Character label for the x-axis (depth).
+#' @param y.label Character label for the y-axis (age).
 #' @param events List object describing the specifics of the climate transitions
 #' used in 'formula'. Must include an item called \code{locations}. See \code{?events.default} for details.
 #' @param synchronization List object describing specifics related to tie-points
@@ -82,6 +84,7 @@
 #' summary(object)
 #' @export
 bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
+                          x.label=NULL,y.label=NULL,
                           events = NULL,
                           synchronization=NULL,
                           control.fit=NULL,
@@ -143,12 +146,14 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
     #eventindexes = unique(c(1,eventindexes[!is.na(eventindexes)]))
     eventindexes = unique(c(1,eventindexes[!is.na(eventindexes)],length(y)))
     nevents = length(eventindexes)-1
+
+    events$nevents=nevents
+    events$eventindexes=eventindexes
   }else{
     nevents=0
     eventindexes=NULL
   }
-  events$nevents=nevents
-  events$eventindexes=eventindexes
+
 
 
 
@@ -241,6 +246,8 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
                     responsename=responsename,
                     data.input=data,
                     reference.label=reference.label,
+                    x.label=x.label,
+                    y.label=y.label,
                     events=events,
                     synchronization=synchronization,
                     control.fit=control.fit,

### 2022-10-25 R/helpfulfunctions.R

@@ -20,7 +20,7 @@ meanmaker = function(coefs,reg.model,data){
   fitted = numeric(n)
   for (i in 1:length(coefs)){
     name = names(reg.model)[i]
-    if(name == "`(Intercept)`"){
+    if(name == "`(Intercept)`" || name == "(Intercept)"){
       fitted = fitted + coefs[i]
     }else{
       fitted = fitted + coefs[i]*data[[name]]
@@ -530,11 +530,20 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     if(ncol(synchronization$samples) != length(synchronization$locations))
       warning("number of columns in 'samples' must correspond to length of 'locations'!")
     samples = synchronization$samples
+    if(tolower(synchronization$locations_unit) %in% c("depth","z")){
+      locations_indexes = which.index(synchronization$locations,object$data$depth)
+    }else if(tolower(synchronization$locations_unit) %in% c("age","time","y")){
+      locations_indexes = which.index(synchronization$locations,object$data$age)
+    }else{
+      locations_indexes = synchronization$locations
+    }
   }else{
     if(synchronization$method=="adolphi"){
       synchronization$locations = c(11050,12050,13050,22050,42050)
       synchronization$locations_unit="age"
       locations_indexes = which.index(synchronization$locations,object$data$age)
+      locations_indexes = locations_indexes[!is.na(locations_indexes)]
+      synchronization$locations = synchronization$locations[!is.na(locations_indexes)]
       tieshifts = synchronization$locations
       synchronization$x.ref=tieshifts
 
@@ -570,6 +579,7 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
 
   }
 
+  n = nrow(object$data)
   object$tie_points = list(samples=samples,
                            locations=synchronization$locations,
                            locations_unit=synchronization$locations_unit,
@@ -577,7 +587,11 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
                            nsims=synchronization$nsims,
                            x.ref=synchronization$x.ref,
                            locations_indexes=locations_indexes,
-                           tie_n=length(synchronization$locations))
+                           tie_indexes = locations_indexes,
+                           free_indexes = (1:n)[-locations_indexes],
+                           free_n=n-length(synchronization$locations),
+                           tie_n=length(synchronization$locations)
+                           )
 
   time.total = difftime(Sys.time(), time.start,units="secs")[[1]]
 

### 2022-10-25 R/plot.bremla.R

@@ -258,10 +258,23 @@ plot.bremla = function(x,
         fullpd = data.frame(depth = x$data$depth)
       }
 
+      if(!is.null(x$.args$x.label)){
+        x.label = x$.args$x.label
+      }else{
+        x.label="Depth"
+      }
+
+      if(!is.null(x$.args$y.label)){
+        y.label = x$.args$y.label
+      }else{
+        y.label="Age - original time scale"
+      }
+
       gg2 = ggplot(data=fullpd,aes(x=.data$depth)) +
         geom_line(aes(y=0),color="blue",linetype="dotted",size=0.2)+
-        theme_bw()+ylab("Estimated age - reference (years)")+
-        xlab("NGRIP depth (m)")
+        theme_bw()+
+        ylab(y.label)+
+        xlab(x.label)
 
       if(!is.null(x$simulation$summary)){
         gg2 = gg2 +
@@ -374,10 +387,23 @@ plot.bremla = function(x,
         fullpd = data.frame(depth = x$data$depth[free_indexes])
       }
 
+      if(!is.null(x$.args$x.label)){
+        x.label = x$.args$x.label
+      }else{
+        x.label="Depth"
+      }
+
+      if(!is.null(x$.args$y.label)){
+        y.label = x$.args$y.label
+      }else{
+        y.label="Age - original time scale"
+      }
+
       gg2 = ggplot(data=fullpd,aes(x=.data$depth)) +
         geom_line(aes(y=0),color="blue",linetype="dotted",size=0.2)+
-        theme_bw()+ylab("Estimated age - reference (years)")+
-        xlab("NGRIP depth (m)")
+        theme_bw()+
+        ylab(y.label)+
+        xlab(x.label)
 
       if(!is.null(x$simulation$summary_sync)){
         gg2 = gg2+geom_line(aes(y=.data$medians))+
@@ -394,7 +420,7 @@ plot.bremla = function(x,
                                    col="magenta")
         }else{
           gg2 = gg2 + geom_point(data=data.frame(tiedepths = x$data$depth[tie_indexes],
-                                                 tiemid = x$simulation$summary_sync$median),
+                                                 tiemid = x$simulation$summary_sync$median[tie_indexes]-ageref_tie),
                                  aes(x=.data$tiedepths,y=.data$tiemid),
                                  color="magenta")
         }

### 2022-10-31 inst/Examples/Paper_results/Paper_results_1.Rmd
Changes to markdown files for reproducing results
and for using some october github action minutes

@@
 library(bremla)
+library(dplyr)
+library(ggplot2)
 data("event_intervals")
@@
             locations_unit="depth",degree=1)
+
  plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
@@
                 )
+
 ```

### 2022-10-31 inst/Examples/Paper_results/Paper_results_2.Rmd
Changes to markdown files for reproducing results
and for using some october github action minutes

@@
             locations_unit="depth",degree=1)
-plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
-     xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
+# plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
+     # xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
 formula = dy~-1+depth2+proxy
@@
 ```{r adolphi, eval=TRUE}
-adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = TRUE,
-                            x.ref=c(11050,12050,13050,22050,42050))
+library(ggplot2)
+x.ref=c(11050,12050,13050,22050,42050)
+adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = FALSE,
+                            x.ref=x.ref)
+lower = c(-77.53, -30.92, -12.37, 336.19, 101.51)+x.ref
+upper = c(-41.09, 11.94, 21.99, 714.03, 509.08)+x.ref
+means = c(-59.95, -7.04, 3.34, 521.44, 275.24)+x.ref
+medians = c(-60.10, -6.42, 2.64, 532.36, 243.57)+x.ref
+gg1 = ggplot(data=adolphipdfs$tie1,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+  ggtitle("Tie-point 1") + ylab("Density") + 
+  geom_vline(aes(xintercept=x.ref[1]),color="black")+
+  geom_vline(aes(xintercept=means[1]),color="blue")+
+  geom_vline(aes(xintercept=lower[1]),color="red")+
+  geom_vline(aes(xintercept=upper[1]),color="red")
+gg2 = ggplot(data=adolphipdfs$tie2,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+  ggtitle("Tie-point 2") + ylab("Density") + 
+  geom_vline(aes(xintercept=x.ref[2]),color="black")+
+  geom_vline(aes(xintercept=means[2]),color="blue")+
+  geom_vline(aes(xintercept=lower[2]),color="red")+
+  geom_vline(aes(xintercept=upper[2]),color="red")
+gg3 = ggplot(data=adolphipdfs$tie3,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+  ggtitle("Tie-point 3") + ylab("Density") + 
+  geom_vline(aes(xintercept=x.ref[3]),color="black")+
+  geom_vline(aes(xintercept=means[3]),color="blue")+
+  geom_vline(aes(xintercept=lower[3]),color="red")+
+  geom_vline(aes(xintercept=upper[3]),color="red")
+gg4 = ggplot(data=adolphipdfs$tie4,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+  ggtitle("Tie-point 4") + ylab("Density") + 
+  geom_vline(aes(xintercept=x.ref[4]),color="black")+
+  geom_vline(aes(xintercept=means[4]),color="blue")+
+  geom_vline(aes(xintercept=lower[4]),color="red")+
+  geom_vline(aes(xintercept=upper[4]),color="red")
+gg5 = ggplot(data=adolphipdfs$tie5,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+  ggtitle("Tie-point 5") + ylab("Density") + 
+  geom_vline(aes(xintercept=x.ref[5]),color="black")+
+  geom_vline(aes(xintercept=means[5]),color="blue")+
+  geom_vline(aes(xintercept=lower[5]),color="red")+
+  geom_vline(aes(xintercept=upper[5]),color="red")
+
+library(ggpubr)
+ggarrange(ggarrange(gg1,gg2,gg3,nrow=1),
+          ggarrange(gg4,gg5,nrow=1),
+          nrow=2)
+# ggsave("adolphipdfs-4800x2000.eps", device=cairo_ps,width=4800,height=2400,units="px",dpi=500,limitsize=FALSE)
+
 ```
@@
 }
+
+ggp = ggplot() + theme_bw() + xlim(rev(range(results$data$age)))+
+  xlab("GICC05 (yb2k)")+ylab(expression(paste(delta^18,"O (permil)")))
+nevents=results$.args$events$nevents
+agestarts = numeric(nevents)
+for(i in 1:nevents){
+  indexes = which(results$data[[paste0("psi0_",i)]]>0)
+  age_i = results$data$age[indexes]
+  agestarts[i]=age_i[1]
+  proxy_i = results$data$proxy[indexes]
+  if(i %% 2 ==0 ){
+    color="red"
+  }else{
+    color="blue"
+  }
+  ggp = ggp + geom_line(data=data.frame(age_i=age_i,proxy_i=proxy_i),
+                        mapping=aes(x=age_i,y=proxy_i),col=color)
+}
+agestarts[nevents]=last(results$data$age)
+ggp = ggp + geom_vline(xintercept = agestarts,col="black",size=0.15)
+print(ggp)
+# ggsave("proxyplot-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
+
 ```
@@
 library(ggplot2)
+library(INLA)
 synchronization = list(method="adolphi",nsims=nsims,
@@
                 control.fit=list(noise="ar1"),
-                control.sim=list(synchronized=TRUE,summary=list(CI.type="hpd")) ,
+                control.sim=list(synchronized=2,summary=list(CI.type="hpd")) ,
                 print.progress=TRUE
                 )
-tiepointranges = as.data.frame(matrix(NA,nrow=5,ncol=3))
+tiepointranges = as.data.frame(matrix(NA,nrow=4,ncol=3))
 gicc05=results_sync$original.chron$age[1+1:n]
@@
 colnames(tiepointranges)=c("lower","upper","depth")
-for(i in 2:5){
-  tiepointranges[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",i)]])-gicc05tie[i]
+for(i in 1:4){
+  tiepointranges[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",i+1)]])-gicc05tie[i]
 }
@@
 {
-  layout(mat=matrix(c(1,2,1,3),nrow=2))
-  par(mar=c(5,4,4,2)+0.1)
+  ggd1 = data.frame(depth=depth[interval],proxy=proxy[interval],
+                    x=results_event$linramp$data$x,
+                    mean=results_event$linramp$linrampfit$mean,
+                    lower=results_event$linramp$linrampfit$q0.025,
+                    upper=results_event$linramp$linrampfit$q0.975
+                    )
+  gg1 = ggplot(data=ggd1,aes(x=depth))+geom_line(aes(y=proxy),col="gray")+
+    theme_bw()+ xlim(rev(range(depth[interval])))+xlab("Depth (m)")+
+    ylab(expression(paste(delta^18,"O (permil)")))+ggtitle("(a) GI-11 Linear ramp fit")+
+    geom_ribbon(aes(x=x,ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(x=x,y=mean),col="blue",size=0.7)
 
-  plot(depth[interval],proxy[interval],type="l",col="gray",
-     main=paste0("(a) ",transitionlabel," linear ramp fit"),
-     xlab="Depth (m)", ylab=expression(paste(delta^18,"O (permil)")),
-     xlim = rev(range(depth[interval]))
-     )
-lines(results_event$linramp$data$x,results_event$linramp$linrampfit$mean,lwd=1.5)
-lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.025,lwd=1.5,col="red")
-lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.975,lwd=1.5,col="red")
-abline(v=depth.reference,lty=3,lwd=0.7,col="black")
   ybottom = min(results_event$linramp$data$y)
@@
   normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
-lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
-ybottom = min(results_event$linramp$data$y)
-ytop = min(results_event$linramp$linrampfit$q0.025)
+  gg1 = gg1 + geom_line(data=data.frame(x=margt0[,1],y=normt0.y),aes(x=x,y=y),col="blue")
+
   margt1 = results_event$linramp$param$t1$marginal
   normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
-lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
-plot(results_event$linramp$param$t0$marg.t0,type="l",xlab=expression(paste("Onset depth (m)")),ylab="Density",
-     main="(b) Marginal posterior of onset depth")#, 
-     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
-#abline(v=depth.reference,lty=3)
-abline(v=results_event$linramp$param$t0$mean,lwd=1.5)
-abline(v=results_event$linramp$param$t0$q0.025,col="red")
-abline(v=results_event$linramp$param$t0$q0.975,col="red")
-plot(results_event$linramp$param$t1$marginal,type="l",xlab=expression(paste("End point depth (m)")),ylab="Density",
-     main="(c) Marginal posterior of end point depth")#, 
-     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
-#abline(v=depth.reference,lty=3)
-abline(v=results_event$linramp$param$t1$mean,lwd=1.5)
-abline(v=results_event$linramp$param$t1$q0.025,col="red")
-abline(v=results_event$linramp$param$t1$q0.975,col="red")
-par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
-hist(results_event$event_dating$samples,main=paste0(transitionlabel," Onset age"),
-     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40, freq=0)
-abline(v=age.reference,lty=3)
-abline(v=results_event$event_dating$mean,lwd=2)
-abline(v=results_event$event_dating$q0.025,lwd=2,col="red")
-abline(v=results_event$event_dating$q0.975,lwd=2,col="red")
+  gg1 = gg1 + geom_line(data=data.frame(x=margt1[,1],y=normt1.y),aes(x=x,y=y),col="blue",
+                        linetype="dashed")
+
+  gg1 = gg1 + geom_vline(aes(xintercept=depth.reference),linetype="dotted",size=0.8)
+
+  gd2 = as.data.frame(results_event$linramp$param$t0$marg.t0)
+  gg2 = ggplot(data=gd2,mapping=aes(x=x,y=y)) + geom_line()+theme_bw()+
+    xlab("Onset depth (m)") + ylab("Density") + ggtitle("(b) Marginal posterior of onset depth")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t0$mean),col="blue")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t0$q0.025),col="red")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t0$q0.975),col="red")
+
+  gd3 = as.data.frame(results_event$linramp$param$t1$marginal)
+  gg3 = ggplot(data=gd3,mapping=aes(x=x,y=y)) + geom_line()+theme_bw()+
+    xlab("Onset depth (m)") + ylab("Density") + ggtitle("(c) Marginal posterior of end-point depth")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t1$mean),col="blue")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t1$q0.025),col="red")+
+    geom_vline(aes(xintercept = results_event$linramp$param$t1$q0.975),col="red")
+
+  ggarrange(gg1,ggarrange(gg2,gg3,nrow=1),nrow=2)
+  #ggsave("GI-11-rampfit-4800x3000.eps", device=cairo_ps,width=4800,height=3000,units="px",dpi=500,limitsize=FALSE)
+
+  gghist = ggplot(data = data.frame(ages=results_event$event_dating$samples)) + theme_bw()+
+    geom_histogram(aes(x=ages,y=..density..),fill="orange",col="black")+xlab("Onset age (yb2k)")+
+    ylab("Density") + ggtitle("GI-11 Onset age") +
+    geom_vline(aes(xintercept = age.reference),lty=3,size=0.8)+
+    geom_vline(aes(xintercept=results_event$event_dating$mean),col="blue",size=0.7)+
+    geom_vline(aes(xintercept=results_event$event_dating$q0.025),col="red",size=0.7)+
+    geom_vline(aes(xintercept=results_event$event_dating$q0.975),col="red",size=0.7)+
+    xlim(rev(range(results_event$event_dating$samples)))
+
+  #ggsave("GI-11-histogram-4800x3000.eps", device=cairo_ps,width=4800,height=3000,units="px",dpi=500,limitsize=FALSE)
+
 }
@@
 ```{r events_import, eval=FALSE}
+##### IMPORT MYRVOLL-NILSEN 2022 DATA AND COMPARE POSTERIORS ####
 library(readxl)
@@
 rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`
-buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
+buizert_onsets = read_excel("data/Buizert_onsets.xlsx",
                             col_types=c("text","numeric","numeric","numeric","numeric"))
@@
 buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]
-capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
+capron_onsets_NGRIP_d18O = read_excel("data/Capron_onsets.xls",
                            sheet="d18O",n_max=25)
 capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`
-capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
-                                sheet="Ca2+",n_max=24)
-capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`
 capind = numeric(29); buiind = numeric(29)
@@
 agesimmatrix_d18O = matrix(NA,29,nsims)
+agesimmatrix_d18O_unsync = matrix(NA,29,nsims)
 depthlist_d18O = c()
@@
   agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)
+  agetemp_d18O_unsync = events_depth_to_age(temp_d18O,
+                                            control.transition_dating=list(label="d18O",
+                                                                           sync=FALSE))
 
   agesimmatrix_d18O[eventnumber,] = agetemp_d18O$event_dating$samples 
+  agesimmatrix_d18O_unsync[eventnumber,] = agetemp_d18O_unsync$event_dating$samples 
   if(do.plot.agehist){
@@
 ```{r compare_transitions_2}
-{
-par(mfrow=c(5,6),mar=c(2,2,1.5,1))
+
+  # library(cowplot)
+
+  ggplots = c()
+
   for(i in 1:29){
@@
     dens = data.frame(x=dens$x,y=dens$y)#/max(dens2$y))
-  xlim = range(dens$x,event_intervals$GICC_age.yb2k[i]);ylim = range(dens$y)
+    densunsync = density(agesimmatrix_d18O_unsync[i,])
+    densunsync = data.frame(x=densunsync$x,y=densunsync$y)#/max(dens2$y))
+    xlim = rev(range(dens$x,densunsync$x))
 
-  if(!is.na(buiind[i])) xlim=range(xlim,buizert_onsets$Age[buiind[i]])
-  if(!is.na(capind[i])) xlim=range(xlim,capron_ages_NGRIP_d18O[capind[i]])
-  #par(mar=c(2,2,1.5,1))
+    ggd = data.frame(x=dens$x,y=dens$y,x0=densunsync$x,y0=densunsync$y)
+    ggdrasmus = data.frame(rasmus=event_intervals$GICC_age.yb2k[i])
+    title = str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1])
 
-  plot(dens,type="l",col=1,xlab="Onset year (b2k)",ylab="Density",
-       main=str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]),
-       xlim=xlim,ylim=ylim)
-  #Axis(side=1,labels=FALSE,at=dens1$x,tck=0.1)
+    gg = ggplot(data=ggd) + theme_bw()+xlim(xlim)+
+      geom_line(aes(x=x0,y=y0),col="gray",size=0.7)+
+      geom_line(aes(x=x,y=y),col="black",size=0.7)+
+      xlab("Onset year (yb2k)")+ylab("Density") +
+      ggtitle(title)+
+      theme(axis.text.y=element_blank(),
+            axis.ticks.y=element_blank())+
+      # geom_vline(aes(xintercept = event_intervals$GICC_age.yb2k[i]),col="blue",size=0.7)
+      geom_vline(data=ggdrasmus, aes(xintercept = rasmus),col="blue",size=0.7)
 
-  #lines(dens,type="l",col="gray")
-  if(FALSE){#95% credible intervals
-    abline(v=mean(agesimmatrix_dust[i,]),col=1)
-    zdens1 = inla.zmarginal(dens1,silent = TRUE)
-    abline(v=c(zdens1$quant0.025,zdens1$quant0.975),col=1,lty=3,lwd=0.8)
-    abline(v=mean(agesimmatrix_d18O[i,]),col=2)
-    zdens2 = inla.zmarginal(dens2,silent = TRUE)
-    abline(v=c(zdens2$quant0.025,zdens2$quant0.975),col=2,lty=3,lwd=0.8)
-  }
-
-  abline(v=event_intervals$GICC_age.yb2k[i],col="blue")  
   if(!is.na(buiind[i])){
-    abline(v=buizert_onsets$Age[buiind[i]],col="green")
+    ggdbuizert = data.frame(buizert=buizert_onsets$Age[buiind[i]])
+    gg = gg + geom_vline(data=ggdbuizert,aes(xintercept = buizert),col="green",size=0.7)
+    #abline(v=buizert_onsets$Age[buiind[i]],col="green")
     #abline(v=buizert_ages[buiind[i]],col="green")
@@
 
-  # abline(v=capron_ages_NGRIP_d18O,col="orange")
-  # abline(v=capron_ages_NGRIP_dust,col="orange",lty=3)
   if(!is.na(capind[i])){
-    abline(v=capron_ages_NGRIP_d18O[capind[i]],col="red")
+    ggdcapron = data.frame(capron=capron_ages_NGRIP_d18O[capind[i]])
+    gg = gg + geom_vline(data=ggdcapron,aes(xintercept = capron),col="red",size=0.7)
+    # abline(v=capron_ages_NGRIP_d18O[capind[i]],col="red")
+  }
+    # ggplots = c(ggplots,list(gg))
+    ggplots[[i]] = gg
   }
 
-}
-par(mar=c(0,0,0,0));plot(-1,axes=FALSE,xlab="",ylab="",xlim=c(0,1),ylim=c(0,1))
-legend(0,1,legend=c("Onset posterior distribution",#"Ca2+ onset posterior",
-                    "Rasmussen onset","Buizert onset",
-                    "Capron onset"),#"Capron Ca2+ onset"),
-       col=c("black", "blue", "green", "red"),
-       lty=c(1,1,1,1,1,1), cex=1,bty="n")
-par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)
-}
+  ggl=ggplot(data.frame(x = 1,x0=1, y = 1,y0=1, rasmus=1,buizert=1,capron=1))+
+    geom_line(aes(x=x,y=y,col="Synchronized onset posterior"),size=1)+
+    geom_line(aes(x=x0,y=y0,col="Unsynchronized onset posterior"))+
+    geom_line(aes(x=x,y=rasmus,col="Rasmussen onset"))+
+    geom_line(aes(x=x,y=buizert,col="Buizert onset"))+
+    geom_line(aes(x=x,y=capron,col="Capron onset"))+
+                        # colour = 'Something2'), aes(x, y, fill = colour))+
+  # geom_point(alpha=0, shape = 0)+ # completely transparent rectangular point 
+  # scale_fill_manual(values='black', drop=FALSE) +
+   # guides(color = guide_legend(override.aes = list(alpha=1, size = 1.3)))+ # showing the point in the legend
+    scale_color_manual(values=c("Synchronized onset posterior"="black",
+                                "Unsynchronized onset posterior"="gray",
+                                "Rasmussen onset"="blue",
+                                "Buizert onset"="green",
+                                "Capron onset"="red"),
+                       labels=c("Synchronized onset posterior",
+                                "Unsynchronized onset posterior",
+                                "Rasmussen onset",
+                                "Buizert onset",
+                                "Capron onset"))+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        # legend.text = element_text(size = 17),
+        legend.key = element_rect(fill='NA'),
+        # legend.key.size = unit(0.8,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(0.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+  ggplots[[30]] = ggl
+  ggarrange(plotlist=ggplots,widths=c(6,5))
+
+  # ggsave("allevents-4800x3000.eps",device=cairo_ps,width=4800,height=3000,units="px",limitsize=FALSE)
+
 rownames=c()
@@
 tiedepths = c(100,300,600)
-tieoffset = c(-50,0,100)
-#tieage = c(450,3000,8500)
-tieage = y[tiedepths]+tieoffset
+tieage = c(450,3000,8500)
 
@@
   geom_line(aes(y=mean),col="blue",size=0.7)+
-  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="magenta",size=1.2)
+  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="magenta",size=1)
 #plot(res)
@@
 
-tieoffset = c(-50,0,100)
-tiemeans = y[tiedepths]+tieoffset
-#tiemeans = c(450,3000,8500)
+tiemeans = c(450,3000,8500)
 tiesds = c(10,20,30)

### 2022-10-31 R/helpfulfunctions.R

@@ -522,7 +522,7 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
   #if(!is.null(synchronization))
   synchronization = set.options(synchronization,synchronization.default())
 
-  object$.args$synchronization = synchronization
+
 
 
   if(!is.null(synchronization$samples)){
@@ -539,16 +539,18 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     }
   }else{
     if(synchronization$method=="adolphi"){
-      synchronization$locations = c(11050,12050,13050,22050,42050)
-      synchronization$locations_unit="age"
-      locations_indexes = which.index(synchronization$locations,object$data$age)
-      locations_indexes = locations_indexes[!is.na(locations_indexes)]
-      synchronization$locations = synchronization$locations[!is.na(locations_indexes)]
+      adolphilocs = c(11050,12050,13050,22050,42050)
       tieshifts = synchronization$locations
+      samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
+                                        tieshifts=adolphilocs,...)
+      synchronization$locations_unit="age"
+      locations_indexes = which.index(adolphilocs,object$data$age)
+      samples = samples[,!is.na(locations_indexes)]
+      locations_indexes = locations_indexes[!is.na(locations_indexes)]
+      synchronization$locations = object$data$age[locations_indexes]
+      #synchronization$locations = synchronization$locations[!is.na(locations_indexes)]
       synchronization$x.ref=tieshifts
 
-      samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
-                                        tieshifts=tieshifts,...)
     }else if(synchronization$method %in% c("normal","gaussian","gauss")){
 
       if(tolower(synchronization$locations_unit) %in% c("depth","z")){
@@ -575,9 +577,8 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
       }
     }
 
-
-
   }
+  object$.args$synchronization = synchronization
 
   n = nrow(object$data)
   object$tie_points = list(samples=samples,

### 2022-11-04 inst/Examples/Paper_results/Paper_results_2.Rmd
Added support for general Gaussian noise models
Added support for custom Gaussian models for the noise component to be added using rgeneric. Added functions to make this specification easier, and made changes to plot, summary, etc such that the results can interpret the results from the new model property

@@
 # ggsave("proxyplot-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
+
+ggd = data.frame(age=data$age,depth=data$depth,proxy=data$proxy)
+ggdr = data.frame(depths=events_rasmussen$depth[events_rasmussen$depth>1492.45 & events_rasmussen$depth<2426],ages=events_rasmussen$age[events_rasmussen$age>11703.07 & events_rasmussen$age < 59944.50])
+library(ggplot2)
+ggp2 = ggplot(data=ggd) + theme_bw()+
+  geom_line(aes(x=age,y=proxy),col="black",size=0.5)+
+  geom_vline(data=ggdr,mapping=aes(xintercept=ages),col="gray",size=0.5)+
+  xlab("GICC05 age (yb2k)") + ylab(expression(paste(delta^18,"O (permil)")))
+
+# ggsave("proxyplot-bw-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
+
+# ggplot(data=ggd)+geom_line(aes(x=depth,y=age))
 

### 2022-11-04 R/bremla_chronology_simulation.R

@@ -90,8 +90,21 @@ bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE
 
     if(print.progress) cat("Simulating ",nsims, " hyperparameters from INLA posterior...",sep="")
 
+    if(tolower(noise) %in% c("rgeneric","custom")){
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      param.names = object$.args$control.fit$rgeneric$param.names
+      for(i in 1:length(object$.args$control.fit$rgeneric$from.theta)){
+        paramsamp = object$.args$control.fit$rgeneric$from.theta[[i]](hypersamples[,i])
+        if(is.null(param.names[i]) || is.na(param.names[i])){
+          tempname = paste0("hyperparameter",i)
+          object$simulation[[tempname]] = paramsamp
+        }else{
+          object$simulation[[param.names[i] ]] = paramsamp
 
-    if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
+        }
+      }
+
+    }else if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
       hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]))
     }else if (tolower(noise) %in% c(1,"ar1","ar(1)")){
@@ -162,7 +175,14 @@ bremla_chronology_simulation = function(object, control.sim,print.progress=FALSE
       if(control.sim$store.everything) object$simulation$dmean[,i] = dmeansim #store mean if we want
 
       ##sample noise component
-      if(tolower(noise) %in% c(0,"iid","independent","ar0","ar(0)")){
+      if(tolower(noise) %in% c("rgeneric","custom")){
+        theta = hypersamples[i,]
+
+        Q = object$.args$control.fit$rgeneric$Q(theta,n,ntheta=length(theta))
+        muvek = numeric(n)
+        noisesim = Qsimmer(1, Q, muvek)
+
+      }else if(tolower(noise) %in% c(0,"iid","independent","ar0","ar(0)")){
         noisesim = rnorm(n,mean=0,sd=object$simulation$sigma[i])
 
       }else if(tolower(noise) %in% c(1,"ar1","ar(1)")){

### 2022-11-04 R/bremla_modelfitter.R

@@ -157,7 +157,10 @@ bremla_modelfitter = function(object, control.fit,
     object$fitting$LS$params[["phi2"]] = phi2
     object$fitting$LS$noisefit=noisefit
   }else{
-    stop(paste0("Noise model is set to ",noise,". Only iid, ar1 and ar2 are currently supported!"))
+    if(!(tolower(object$.args$control.fit$noise) %in% c("rgeneric","custom"))){
+      stop(paste0("Noise model is set to ",noise,". Only iid, ar1, ar2 and 'rgeneric' are currently supported!"))
+    }
+
   }
   time.ls=Sys.time()
   if(print.progress) cat(" completed!\n",sep="")
@@ -176,18 +179,27 @@ bremla_modelfitter = function(object, control.fit,
     object$.internal$initial_fixed = my.control.fixed
     resi = object$fitting$LS$fit$residuals
 
-    initialmodes = log(1/object$fitting$LS$params$sigma^2)
+    if(tolower(object$.args$control.fit$noise) %in% c("rgeneric","custom")){
+      warning("rgeneric is experimental. Some irraneous behaviour can be fixed by changing the number of cores in 'control.fit$ncores'")
+      initialmodes = numeric(length(object$.args$control.fit$rgeneric$from.theta))
+      model.rgeneric = object$.args$model.rgeneric
+    }else{
+      initialmodes = log(1/object$fitting$LS$params$sigma^2)
 
-    if(tolower(object$.args$control.fit$noise) %in% c(1,"ar1","ar(1)")){
-      phi.ls = object$fitting$LS$params$phi
-      initialmodes = c(initialmodes, log( (1+phi)/(1-phi) ))
+      if(tolower(object$.args$control.fit$noise) %in% c(1,"ar1","ar(1)")){
+        phi.ls = object$fitting$LS$params$phi
+        initialmodes = c(initialmodes, log( (1+phi)/(1-phi) ))
 
-    }else if(tolower(object$.args$control.fit$noise) %in% c(2,"ar2","ar(2)")){
-      phi1.ls = object$fitting$LS$params$phi1
-      phi2.ls = object$fitting$LS$params$phi2
+      }else if(tolower(object$.args$control.fit$noise) %in% c(2,"ar2","ar(2)")){
+        phi1.ls = object$fitting$LS$params$phi1
+        phi2.ls = object$fitting$LS$params$phi2
 
[…]

### 2022-11-04 R/bremla_prepare.R

@@ -98,7 +98,12 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
   if(!is.null(events)) events = set.options(events,events.default())
   if(!is.null(synchronization)) synchronization = set.options(synchronization,
                                                               synchronization.default())
-  if(!is.null(control.fit)) control.fit = set.options(control.fit,control.fit.default())
+  if(is.null(control.fit)) {
+    control.fit=list(noise="ar1")
+
+  }
+  control.fit = set.options(control.fit,control.fit.default())
+
   if(!is.null(control.sim)) control.sim = set.options(control.sim,control.sim.default())
   if(!is.null(control.linramp)) control.linramp = set.options(
                   control.linramp,control.linramp.default())
@@ -216,20 +221,38 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
 
   if(!is.null(control.fit)){ #add random effect to formula string for use in INLA
     if(tolower(control.fit$method) %in% "inla" ){
-      if(tolower(control.fit$noise) %in% c("iid","independent","ar(0)")){
-        formulastring_inla = paste0(formulastring, " + f(idy,model=\"iid\")")
-      }else if(tolower(control.fit$noise) %in% c("ar1","ar(1)")){
-        formulastring_inla = paste0(formulastring, " + f(idy,model=\"ar1\")")
-      }else if(tolower(control.fit$noise) %in% c("ar2","ar(2)")){
-        formulastring_inla = paste0(formulastring, " + f(idy,model=\"ar\",order=2)")
+      if(!is.null(control.fit$rgeneric)){ #rgeneric model is specified
+        control.fit$noise = "rgeneric"
+
+        ntheta = length(control.fit$rgeneric$from.theta)
+
+        model.rgeneric = INLA::inla.rgeneric.define(rgeneric.fitting, n=n-1,
+                                              ntheta=ntheta,
+                                              Q = control.fit$rgeneric$Q,
+                                              log.prior=control.fit$rgeneric$log.prior)
+
+        formulastring_inla = paste0(formulastring, " + f(idy,model=model.rgeneric)")
+
+        # formula=dy ~ -1 + depth + f(idy,model=model.rgeneric)
+
+
+      }else{
+        if(tolower(control.fit$noise) %in% c("iid","independent","ar(0)")){
+          formulastring_inla = paste0(formulastring, " + f(idy,model=\"iid\")")
[…]

### 2022-11-04 R/bremla_synchronized_simulation.R

@@ -92,7 +92,11 @@ time.start = Sys.time()
     tie_locations=object$tie_points$locations
     tie_locations_unit=object$tie_points$locations_unit
     tiepointsims = object$tie_points$samples
-    if(object$tie_points$nsims != nsims) stop("Number of samples given does not correspond to 'nsims'...")
+    if(object$tie_points$nsims < nsims){
+      stop("Number of samples to be produced (nsims) exceeds the number of tie-point samples available. Run 'tiepointsimmer' again...")
+    }else{
+      tiepointsims = tiepointsims[1:nsims,]
+    }
   }else{
     object$tie_points = list(samples=tiepointsims,locations=tie_locations,
                              locations.type=tie_locations_unit,method="precomputed",nsims=nsims)
@@ -126,8 +130,20 @@ time.start = Sys.time()
 
     if(print.progress) cat("Simulating ",nsims, " hyperparameters from INLA posterior...",sep="")
 
+    if(tolower(noise) %in% c("rgeneric","custom")){
+      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
+      param.names = object$.args$control.fit$rgeneric$param.names
+      for(i in 1:length(object$.args$control.fit$rgeneric$from.theta)){
+        paramsamp = object$.args$control.fit$rgeneric$from.theta[[i]](hypersamples[,i])
+        if(is.null(param.names[i]) || is.na(param.names[i])){
+          tempname = paste0("hyperparameter",i)
+          object$simulation[[tempname]] = paramsamp
+        }else{
+          object$simulation[[param.names[i] ]] = paramsamp
 
-    if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
+        }
+      }
+    }else if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
       hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
       object$simulation = list(sigma = 1/sqrt(hypersamples[,1]))
     }else if (tolower(noise) %in% c(1,"ar1","ar(1)")){
@@ -174,22 +190,35 @@ time.start = Sys.time()
   if(print.progress) cat("Simulating synchronized chronologies...\n",sep="")
   timeage.start = Sys.time()
   for(r in 1:nsims){
[…]

### 2022-11-04 R/default_arguments.R

@@ -52,6 +52,24 @@ events.default <- function(){
 #'   the observations. Can be \code{"log"} or \code{"identity"}. If synchronization is
 #'   intended then only \code{"identity"} is currently supported, which is also the
 #'   default value here.}
+#'   \item{\code{rgeneric}} List containing necessary arguments for specifying a custom
+#'   model for the noise component of the layer increments using the \code{rgeneric}
+#'   framework of \code{INLA}. Note that this alternative generally performs less well
+#'   than using already implemented model components in \code{INLA} used when \code{noise}
+#'   is specified. Must include
+#'   \itemize{
+#'       \item{\code{from.theta} List of function specifying transformation from internal to original
+#'       parameter scaling. Function of \code{theta} (internal parameters). \code{length(from.theta)}
+#'       must correspond to \code{length(theta)}. This is only
+#'       used for transforming the parameters and can not be used in the functions below.}
+#'       \item{\code{param.names} Vector of names for parameters (optional).}
+#'       \item{\code{Q} Function that computes the precision matrix of the noise model. Must be
+#'       a function of \code{theta}, \code{n} and \code{ntheta}, and must return a
+#'       sparse matrix.}
+#'       \item{\code{log.prior} Function that computes the logarithm of the prior of the internal
+#'       parameter \code{theta}. Must be a function of \code{theta}, \code{n} and \code{ntheta}
+#'       (even if they are not used).}
+#'   }
 #' }
 #' @return Returns a list including default values for all variables in \code{control.fit}.
 #'
@@ -66,7 +84,8 @@ control.fit.default <- function(){
     verbose=FALSE,
     log.theta.prior=NULL,
     ncores=1,
-    transform="identity"
+    transform="identity",
+    rgeneric = NULL
     )
   )
 }

### 2022-11-04 R/plot.bremla.R

@@ -193,38 +193,74 @@ plot.bremla = function(x,
   if(!is.null(plot.inla.posterior) && !is.null(x$fitting$fit)){
     figure.count <- new.plot(postscript,pdf,prefix,figure.count,...) +1
     if(plot.inla.posterior$posteriors){
-      if(length(plot.inla.posterior$label)<=1){
-        plot.label1 = plot.inla.posterior$label
-        plot.label2 = plot.inla.posterior$label
-        plot.label3 = plot.inla.posterior$label
+
+      if(tolower(x$.args$control.fit$noise) %in% c("rgeneric","custom")){
+        ntheta = length(x$fitting$inla$hyperparameters$results)
+        if(length(plot.inla.posterior$label) == 1){
+          defaultlabel= plot.inla.posterior$label
+        }else{
+          defaultlabel = NULL
+        }
+        for(i in 1:ntheta){
+          if(!is.null(plot.inla.posterior$label[i]) && !is.na(plot.inla.posterior$label[i])){
+            mainlab = plot.inla.posterior$label[i]
+          }else{
+            if(!is.null(defaultlabel)){
+              mainlab=defaultlabel
+            }else{
+              mainlab=""
+            }
+          }
+          xlab = names(x$fitting$inla$hyperparameters$posteriors)[i]
+          plot(x$fitting$inla$hyperparameters$posteriors[[i]],type="l",xlab=xlab,
+               main = mainlab, ylab="Density",lwd=2,)
+          abline(v=x$fitting$inla$hyperparameters$results[[i]]$mean)
+          abline(v=c(x$fitting$inla$hyperparameters$results[[i]]$quant0.025,x$fitting$inla$hyperparameters$results[[i]]$quant0.975),col="gray")
+        }
       }else{
-        plot.label1 = plot.inla.posterior$label[1]
-      }
-      plot(x$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(paste(sigma[epsilon])),ylab="Density",lwd=2,main=plot.label1)
-      abline(v=x$fitting$inla$hyperparameters$results$sigma_epsilon$mean)
-      abline(v=c(x$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,x$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),col="gray")
 
[…]

### 2022-11-04 R/print.bremla.R

@@ -96,12 +96,14 @@ print.bremla = function(x,
 
 
 
-  if(!is.null(x$.args$noise)){
-    if(tolower(x$.args$noise) %in% c(0,"iid","independent")){
+  if(!is.null(x$.args$control.fit$noise)){
+    if(tolower(x$.args$control.fit$noise)){
+      noise = "rgeneric"
+    }else if(tolower(x$.args$control.fit$noise) %in% c(0,"iid","independent")){
       noise = "iid"
-    }else if(tolower(x$.args$noise) %in% c(1,"ar1","ar(1)")){
+    }else if(tolower(x$.args$control.fit$noise) %in% c(1,"ar1","ar(1)")){
       noise = "ar1"
-    }else if(tolower(x$.args$noise) %in% c(2,"ar2","ar(2)")){
+    }else if(tolower(x$.args$control.fit$noise) %in% c(2,"ar2","ar(2)")){
       noise = "ar2"
     }
   }
@@ -109,7 +111,23 @@ print.bremla = function(x,
 
 
     if(!is.null(x$fitting)){
-      if(noise == "iid"){
+      if(noise == "rgeneric"){
+
+        ntheta = length(x$fitting$inla$hyperparameters$posteriors)
+        hypers = matrix(NA,nrow=ntheta,ncol=7)
+        colnames = c("mean","sd","quant0.025","quant0.25","quant0.5","quant0.75","quant0.975")
+        rownames = names(x$fitting$inla$hyperparameters$results)
+        for(i in 1:ntheta){
+          for(j in 1:7){
+            hypers[i,j] = round(x$fitting$inla$hyperparameters$results[[i]][[j]]
+                                ,digits=digits)
+          }
+
+        }
+        hypers = as.data.frame(hypers)
+        colnames(hypers) = colnames
+        rownames(hypers) = rownames
+      }else if(noise == "iid"){
 
         hypers = matrix(round(as.numeric(x$fitting$inla$hyperparameters$results$sigma_epsilon),digits=digits),nrow=1)
         #hypers = matrix( round(as.numeric(x$fitting$hyperparameters$results$sigma_epsilon),digits=digits),ncol=mm )

### 2022-11-04 R/rgeneric.fitting.R

@@ -1,0 +1,171 @@
+#' rgeneric model specification for fitting custom models
+#'
+#' Non-standard models in INLA requires specification using the rgeneric modeling framework.
+#' This includes key functions that provide information on precision matrix, mean vector, priors, graph etc.
+#' The example belows shows how this can be used, and how it is supported within the 'bremla' framework.
+#' The necessary arguments for use in the bremla framework is specified under \code{rgeneric}
+#' in \code{\link{control.fit.default}}.
+#' See example below on how rgeneric  can be used.
+#'
+#' @param cmd Vector containing list of function names necessary for the rgeneric model.
+#' @param theta Vector describing the hyperparameters in internal scaling.
+#'
+#' @return When used an an input argument for \code{inla.rgeneric.define} this will return a model eligible to be used within the R-INLA framework (see example).
+#' @author Eirik Myrvoll-Nilsen, \email{eirikmn91@gmail.com}
+#' @seealso \code{\link{linrampfitter}}
+#' @keywords bremla rgeneric inla
+#'
+#' @examples
+#' \donttest{
+#' if(inlaloader()){
+#' require(INLA)
+#' set.seed(1)
+#'
+#' # Gaussian white noise model using rgeneric and bremla
+#'
+#' # The from.theta function is used to transform hyperparameters from its internal scaling 'theta'
+#' #   to the more natural scaling. This is used to compute posteriors and is not used in
+#' #   the INLA fitting procedure. This must be specified as a list containing functions for each
+#' #   hyperparameter, i.e. length(from.theta) must correspond to length(theta). This function
+#' #   cannot be used by the rgeneric functions 'Q' and 'log.prior'.
+#' from.theta = list(function(x)1/sqrt( exp(x) )) # convert from log(precision) -> standard deviation
+#'
+#' # The Q function computes the precision matrix. This is used in fitting the rgeneric model
+#' #   with INLA as well as for computing simulations via precision matrix. Must be a function
+#' #   of n, theta and ntheta event if they are not used in the function.
[…]

### 2022-11-04 R/summary.bremla.R

@@ -94,7 +94,9 @@ summary.bremla = function(object,
   }
 
   if(!is.null(object$.args$control.fit$noise)){
-    if(tolower(object$.args$control.fit$noise) %in% c(0,"iid","independent")){
+    if(tolower(object$.args$control.fit$noise) %in% c("rgeneric","custom")){
+      noise = "rgeneric"
+    }else if(tolower(object$.args$control.fit$noise) %in% c(0,"iid","independent")){
       noise = "iid"
     }else if(tolower(object$.args$control.fit$noise) %in% c(1,"ar1","ar(1)")){
       noise = "ar1"
@@ -105,7 +107,23 @@ summary.bremla = function(object,
 
 
   if(!is.null(object$fitting)){
-    if(noise == "iid"){
+    if(noise == "rgeneric"){
+      ntheta = length(object$fitting$inla$hyperparameters$posteriors)
+      hypers = matrix(NA,nrow=ntheta,ncol=7)
+      colnames = c("mean","sd","quant0.025","quant0.25","quant0.5","quant0.75","quant0.975")
+      rownames = names(object$fitting$inla$hyperparameters$results)
+      for(i in 1:ntheta){
+        for(j in 1:7){
+          hypers[i,j] = round(object$fitting$inla$hyperparameters$results[[i]][[j]]
+                             ,digits=digits)
+        }
+
+      }
+      hypers = as.data.frame(hypers)
+      colnames(hypers) = colnames
+      rownames(hypers) = rownames
+
+    }else if(noise == "iid"){
 
       hypers = matrix(round(as.numeric(object$fitting$inla$hyperparameters$results$sigma_epsilon),digits=digits),nrow=1)
       #hypers = matrix( round(as.numeric(object$fitting$hyperparameters$results$sigma_epsilon),digits=digits),ncol=mm )

### 2022-11-28 inst/Examples/Paper_results/Paper_results_2.Rmd
Minor changes
Added suppressWarnings in bremla_simulationsummarizer and small changes in the legend of a plot in Paper_results_2.Rmd

@@
                                 "Unsynchronized onset posterior"="gray",
-                                "Rasmussen onset"="blue",
-                                "Buizert onset"="green",
-                                "Capron onset"="red"),
+                                "Rasmussen et al. (2014)"="blue",
+                                "Buizert et al. (2015)"="green",
+                                "Capron et al. (2021)"="red"),
                        labels=c("Synchronized onset posterior",
                                 "Unsynchronized onset posterior",
-                                "Rasmussen onset",
-                                "Buizert onset",
-                                "Capron onset"))+
+                                "Rasmussen et al. (2014)",
+                                "Buizert et al. (2015)",
+                                "Capron et al. (2021)"))+
   theme(axis.title = element_blank(),

### 2022-11-28 R/helpfulfunctions.R

@@ -204,8 +204,8 @@ bremla_simulationsummarizer = function(object,sync=TRUE,print.progress=FALSE){
         dens = density(samples[i,])
 
 
-        lower[i] = INLA::inla.hpdmarginal(0.95,dens)[1]
-        upper[i] = INLA::inla.hpdmarginal(0.95,dens)[2]
+        lower[i] = suppressWarnings(INLA::inla.hpdmarginal(0.95,dens)[1])
+        upper[i] = suppressWarnings(INLA::inla.hpdmarginal(0.95,dens)[2])
       }
     }
   }else{

### 2022-11-28 README.Rmd
Changed version of INLA dependency (testing->stable)
Had some issues fitting the model with the testing version

@@
   fig.path = "man/figures/README-",
+  fig.width = 8,
+  fig.height = 4,
   dev = "png",
-  out.width = "80%"
+  out.width = "90%"
 )

### 2022-11-28 R/bremla_modelfitter.R

@@ -216,10 +216,15 @@ bremla_modelfitter = function(object, control.fit,
                    control.compute=list(config=TRUE),
                    #verbose=TRUE,
                    verbose=object$.args$control.fit$verbose,
-                   control.inla=list(restart=TRUE,h=0.005,dz=0.75,
+                   control.inla=list(restart=TRUE,h=0.005,dz=0.75,#cmin=0
                                      int.strategy="auto"),
                    control.mode=list(theta=initialmodes,restart=TRUE)
                    )
+
+    # inla.fit = INLA::inla(formula=dy~-1+depth2+proxy+f(idy,model="ar1"),
+    #                       data=object$data)
+    # inla.fit = INLA::inla(formula=dy~-1+depth2+proxy+f(idy,model="ar1"),
+    #                       data=object$data)
 
     if(print.progress){
       cat(" completed.\n",sep="")

### 2022-11-28 README.Rmd
Changes to README.Rmd
Adde d18O proxy to predictor in example. Also changed a line in rgeneric.uneven.AR1.R, so as to prevent 'theta' from being undefined

@@
 depth = NGRIP_5cm$depth
-proxy = NGRIP_5cm$d18O
+d18O = NGRIP_5cm$d18O
 
-data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
-formula = dy~-1+depth2
+data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,d18O=d18O)
+formula = dy~-1+depth2+d18O
 

### 2022-12-06 README.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

--- 2022-11-28 README.Rmd
@@
 depth = NGRIP_5cm$depth
-proxy = NGRIP_5cm$d18O
+depth2 = depth^2/depth[1]^2 #normalize for stability
+d18O = NGRIP_5cm$d18O
 
-data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
-formula = dy~-1+depth2
+data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,d18O=d18O)
+formula = dy~-1+depth2+d18O
 

### 2022-12-06 inst/Examples/Paper_results/Paper_results_1.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

@@
 depth = NGRIP_5cm$depth
+depth2 = depth^2/depth[1]^2 #normalize for stability
 proxy = NGRIP_5cm$d18O
 n=length(diff(age))
-data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
+data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
 events=list(locations = events_rasmussen$depth,

### 2022-12-06 inst/Examples/Paper_results/Paper_results_2.Rmd
Rescaled depth2 covariate in examples
Somehow the recent version of INLA struggled with large covariates common for depth^2. Rescaled by dividing by the first value

@@
 depth = z0+seq(from=0.5,500,length.out=1000)
+depth2 = depth^2/depth[1]^2 #normalize for stability
+
 dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
@@
 age = y0+cumsum(dy)
-df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth^2); data=rbind(c(y0,z0,0,0,0),df)
+df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth2); data=rbind(c(y0,z0,0,0,0),df)
 formula=dy~-1+proxy#+depth2
@@
 depth = NGRIP_5cm$depth
+depth2 = depth^2/depth[1]^2 #normalize for stability
 proxy = NGRIP_5cm$d18O
-data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
+data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
 formula = dy~-1+depth2

### 2022-12-06 R/bremla.R

@@ -51,10 +51,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))
@@ -99,9 +100,10 @@
 #' depth = NGRIP_5cm$depth
 #' d18O = NGRIP_5cm$d18O
 #' proxy=d18O
-#' data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
 #' formula = dy~-1+depth2
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
+#' data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)
 #'
 #' lowerints = which.index(event_intervals$depth_int_lower.m, depth[2:length(depth)])
 #' upperints = which.index(event_intervals$depth_int_upper.m, depth[2:length(depth)])

### 2022-12-06 R/bremla_biased_chronologies.R

@@ -31,10 +31,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/bremla_modelfitter.R

@@ -31,10 +31,10 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
-#'
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))
@@ -221,10 +221,6 @@ bremla_modelfitter = function(object, control.fit,
                    control.mode=list(theta=initialmodes,restart=TRUE)
                    )
 
-    # inla.fit = INLA::inla(formula=dy~-1+depth2+proxy+f(idy,model="ar1"),
-    #                       data=object$data)
-    # inla.fit = INLA::inla(formula=dy~-1+depth2+proxy+f(idy,model="ar1"),
-    #                       data=object$data)
 
     if(print.progress){
       cat(" completed.\n",sep="")

### 2022-12-06 R/bremla_prepare.R

@@ -57,7 +57,9 @@
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
+#'
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/bremla_synchronized_simulation.R

@@ -30,10 +30,10 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
-#'
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/helpfulfunctions.R

@@ -143,10 +143,10 @@ which.index = function(events, record,suppress=TRUE){ ## Finds which indices of 
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
-#'
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))
@@ -482,10 +482,12 @@ adolphi_tiepoint_simmer = function(nsims=10000,tieshifts = numeric(5), plotdens=
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#'
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/linrampfitter.R

@@ -32,10 +32,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/plot.bremla.R

@@ -41,10 +41,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/print.bremla.R

@@ -28,10 +28,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 R/summary.bremla.R

@@ -28,10 +28,11 @@
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))
@@ -303,10 +304,10 @@ return(ut)
 #' y0 = 11700;z0=1200
 #' age = y0+cumsum(dy)
 #' depth = 1200 + 1:n*0.05
-#'
+#' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #' formula = dy~-1+depth2 + proxy
-#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth^2)
+#' data = data.frame(age=age,dy=dy,proxy=proxy,depth=depth,depth2=depth2)
 #' data = rbind(c(y0,NA,NA,z0,NA),data) #First row is only used to extract y0 and z0.
 #'
 #' events=list(locations=c(1210,1220,1240))

### 2022-12-06 README.Rmd
Reimplemented rescale proxy feature in linrampfitter.R
Also restructured files related to reproducing results. Other minor changes. Cleaned up old plots and unused files

@@
 ---
-output: github_document
+output: github_document:
+  html_preview: false
 ---
@@
   dev = "png",
-  out.width = "90%"
+  out.width = "100%"
 )

### 2022-12-06 inst/reproduce_results/results.Rmd
Reimplemented rescale proxy feature in linrampfitter.R
Also restructured files related to reproducing results. Other minor changes. Cleaned up old plots and unused files

--- 2022-12-06 inst/Examples/Paper_results/Paper_results_2.Rmd
@@
 
-<!-- Paper_results_2.md is generated from paper_results_2.Rmd. Please edit that file -->
+<!-- results.md is generated from results.Rmd. Please edit that file -->
 
@@
   comment = "#>",
-  fig.path = "man/figures/results2-",
+  fig.path = "man/figures/reproduce/results/results-",
   dev = "png",
-  out.width = "80%"
+  out.width = "100%"
 )
@@
 depth = z0+seq(from=0.5,500,length.out=1000)
-depth2 = depth^2/depth[1]^2 #normalize for stability
+depth2 = depth^2/z0 #normalize for stability
 

### 2022-12-06 R/linrampfitter.R

@@ -123,7 +123,13 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   if(is.null(control.linramp$opt.params)){
     optparams = c(round(length(interval)/2),round(length(interval)/10),df_event$yy[1],df_event$yy[length(interval)]-df_event$yy[1])
   }else{
+    defoptparams = c(round(length(interval)/2),round(length(interval)/10),df_event$yy[1],df_event$yy[length(interval)]-df_event$yy[1])
     optparams = control.linramp$opt.params
+    for(i in 1:length(defoptparams)){
+      if(is.na(optparams[i])){
+        optparams[i] = defoptparams[i]
+      }
+    }
   }
 
   n=length(interval)
@@ -131,7 +137,8 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
   timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)
 
 
-  df0=data.frame(y=df_event$yy,x=df_event$xx)
+  df0=data.frame(y=df_event$yy*control.linramp$rescale.y.factor,
+                 x=df_event$xx)
 
   t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
   #library(numDeriv)
@@ -144,35 +151,40 @@ linrampfitter = function(object,control.linramp,print.progress=FALSE){
 
   ## Finding initial values in INLA optimization procedure by first using a simple optimization
   ## gradient and cost function given here
-  minfun.grad = function(param, args = NULL){
-    return (grad(minfun, param, args=args, method.args = list(r=6)))
+
+  if(control.linramp$imp.fit){
+    minfun.grad = function(param, args = NULL){
+      return (grad(minfun, param, args=args, method.args = list(r=6)))
+    }
+    minfun = function(param, args = NULL){
+      yhat = linramp(timepoints,t0=param[1],dt=param[2],y0=param[3],dy=param[4])
+      mse = sum((args$y-yhat)^2)
+      return(sqrt(mse))
+    }
+
+    ## perform optimization to find good starting values for INLA
+    param=optparams
+    args=list(y=y)
+    fit = optim(param,
+                fn = minfun,
+                gr = minfun.grad,
+                method = "BFGS",
+                control = list(
[…]

### 2022-12-06 README.Rmd
Minor cleanup
README.Rmd and more

@@
 ---
-output: github_document:
-  html_preview: false
+output: github_document
 ---
@@
 
-## installation
+## Installation
 
@@
 ```{r install, eval=FALSE}
-#install.packages("devtools")
-devtools::install_github("eirikmn/bremla")
+#install.packages("remotes")
+remotes::install_github("eirikmn/bremla")
 ```

### 2022-12-06 inst/reproduce_results/results.Rmd
Minor cleanup
README.Rmd and more

@@
 library(gridExtra)
-ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+# ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
 str = ""

### 2022-12-06 R/rgeneric.fitting.R

@@ -99,7 +99,6 @@
 #'                      control.fit=control.fit,
 #'                      control.sim=control.sim,
 #'                      synchronization=synchronization)
-#' #,print.progress=TRUE)
 #' summary(object)
 #' plot(object)
 #'

### 2023-04-14 README.Rmd
Minor bugfixes and features

@@
 
-The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](https://www.iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al.
+The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm.xls' downloaded from [Centre for Ice and Climate](https://www.iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al.
 (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485).

### 2023-04-14 inst/reproduce_results/results.Rmd
Minor bugfixes and features

--- 2022-12-06 inst/reproduce_results/results.Rmd
@@
 x.ref=c(11050,12050,13050,22050,42050)
-adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = FALSE,
+adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),
+                            plotdens = FALSE,
                             x.ref=x.ref)
-lower = c(-77.53, -30.92, -12.37, 336.19, 101.51)+x.ref
-upper = c(-41.09, 11.94, 21.99, 714.03, 509.08)+x.ref
-means = c(-59.95, -7.04, 3.34, 521.44, 275.24)+x.ref
-medians = c(-60.10, -6.42, 2.64, 532.36, 243.57)+x.ref
-gg1 = ggplot(data=adolphipdfs$tie1,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+lower = c(-77.53, -30.92, -12.37, 336.19, 101.51)#+x.ref
+upper = c(-41.09, 11.94, 21.99, 714.03, 509.08)#+x.ref
+means = c(-59.95, -7.04, 3.34, 521.44, 275.24)#+x.ref
+medians = c(-60.10, -6.42, 2.64, 532.36, 243.57)#+x.ref
+adolphipdfs_offsets = c()
+for(i in 1:5){
+  adolphipdfs_offsets[[i]] = cbind(as.numeric(adolphipdfs[[i]]$x)-x.ref[i],
+                                   as.numeric(adolphipdfs[[i]]$y))
+  colnames(adolphipdfs_offsets[[i]]) = c("x","y")
+  #adolphipdfs_offsets[[i]]$x = adolphipdfs[[i]]$x-x.ref
+}
+
+gg1 = ggplot(data=as.data.frame(adolphipdfs_offsets[[1]]),aes(x=x,y=y))+geom_line()+theme_bw()+xlab("Age offset (years)")+
   ggtitle("Tie-point 1") + ylab("Density") + 
-  geom_vline(aes(xintercept=x.ref[1]),color="black")+
+  xlim(c(-100,100))+
+  #geom_vline(aes(xintercept=x.ref[1]),color="black")+
   geom_vline(aes(xintercept=means[1]),color="blue")+
@@
   geom_vline(aes(xintercept=upper[1]),color="red")
-gg2 = ggplot(data=adolphipdfs$tie2,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+gg2 = ggplot(data=as.data.frame(adolphipdfs_offsets[[2]]),aes(x=x,y=y))+geom_line()+theme_bw()+xlab("Age offset (years)")+
   ggtitle("Tie-point 2") + ylab("Density") + 
-  geom_vline(aes(xintercept=x.ref[2]),color="black")+
+    xlim(c(-60,60))+
+  #geom_vline(aes(xintercept=x.ref[2]),color="black")+
   geom_vline(aes(xintercept=means[2]),color="blue")+
@@
   geom_vline(aes(xintercept=upper[2]),color="red")
-gg3 = ggplot(data=adolphipdfs$tie3,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+gg3 = ggplot(data=as.data.frame(adolphipdfs_offsets[[3]]),aes(x=x,y=y))+geom_line()+theme_bw()+xlab("Age offset (years)")+
   ggtitle("Tie-point 3") + ylab("Density") + 
-  geom_vline(aes(xintercept=x.ref[3]),color="black")+
+    xlim(c(-60,60))+
+  #geom_vline(aes(xintercept=x.ref[3]),color="black")+
   geom_vline(aes(xintercept=means[3]),color="blue")+
@@
   geom_vline(aes(xintercept=upper[3]),color="red")
-gg4 = ggplot(data=adolphipdfs$tie4,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+gg4 = ggplot(data=as.data.frame(adolphipdfs_offsets[[4]]),aes(x=x,y=y))+geom_line()+theme_bw()+xlab("Age offset (years)")+
   ggtitle("Tie-point 4") + ylab("Density") + 
-  geom_vline(aes(xintercept=x.ref[4]),color="black")+
+    xlim(c(-0,1000))+
+  #geom_vline(aes(xintercept=x.ref[4]),color="black")+
   geom_vline(aes(xintercept=means[4]),color="blue")+
@@
   geom_vline(aes(xintercept=upper[4]),color="red")
-gg5 = ggplot(data=adolphipdfs$tie5,aes(x=x,y=y))+geom_line()+theme_bw()+xlab("GICC05 age (yb2k)")+
+gg5 = ggplot(data=as.data.frame(adolphipdfs_offsets[[5]]),aes(x=x,y=y))+geom_line()+theme_bw()+xlab("Age offset (years)")+
   ggtitle("Tie-point 5") + ylab("Density") + 
-  geom_vline(aes(xintercept=x.ref[5]),color="black")+
+    xlim(c(-250,750))+
+  #geom_vline(aes(xintercept=x.ref[5]),color="black")+
   geom_vline(aes(xintercept=means[5]),color="blue")+
@@
 library(ggpubr)
-ggarrange(ggarrange(gg1,gg2,gg3,nrow=1),
+ggarrange(ggarrange(gg2,gg3,nrow=1),
           ggarrange(gg4,gg5,nrow=1),
           nrow=2)
-# ggsave("adolphipdfs-4800x2000.eps", device=cairo_ps,width=4800,height=2400,units="px",dpi=500,limitsize=FALSE)
+ # ggsave("adolphipdfs-no1-4800x2000.eps", device=cairo_ps,width=4800,height=2400,units="px",dpi=500,limitsize=FALSE)
 
@@
 }
+library(dplyr)
 agestarts[nevents]=last(results$data$age)
@@
 {
+  library(matrixStats)
   library(ggplot2)
@@
 ```{r fixed, eval=TRUE}
+library(INLA)
 fixedsims = cbind(#rep(inla.mmarginal(adolphipdfs$tie1),nsims),
@@
 
-Synchronizing our chronologies with the Adolphi tie-points where the uncertainties is incorporated cam be done using Monte Carlo simulations. All of this can be done using the bremla function.
+## Uncertain tie-points
+
+Synchronizing our chronologies with the Adolphi tie-points where the uncertainties is incorporated can be done using Monte Carlo simulations. All of this can be done using the bremla function.
 
@@
 
+Excluding the contested fourth tie-point can be done as follows:
+```{r sync2, eval = TRUE}
+library(ggplot2)
+library(INLA)
+synchronization = list(method="adolphi",nsims=nsims,
+                       locations=c(NA,1502.80,1532.70,NA,2132.40),
+                       locations_unit="depth")
+results_sync2 = bremla(formula,data,reference.label="GICC05",
+                nsims=nsims,
+                events=events,
+                synchronization=synchronization,
+                control.fit=list(noise="ar1"),
+                control.sim=list(synchronized=TRUE,summary=list(CI.type="hpd")) ,
+                print.progress=TRUE
+                )
+tiepointranges = as.data.frame(matrix(NA,nrow=3,ncol=3))
+gicc05=results_sync2$original.chron$age[1+1:n]
+gicc05tie = gicc05[results_sync2$tie_points$locations_indexes]
+colnames(tiepointranges)=c("lower","upper","depth")
+for(i in 1:3){
+  tiepointranges[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",c(2,3,5)[i])]])-gicc05tie[i]
+}
+tiepointranges[,3] = depth[results_sync2$tie_points$locations_indexes]
+{
+
+  ggd = data.frame(depth=results_sync2$data$depth,
+                   mean=results_sync2$simulation$summary_sync$mean-gicc05,
+                   lower=results_sync2$simulation$summary_sync$lower-gicc05,
+                   upper=results_sync2$simulation$summary_sync$upper-gicc05,
+                   zeros = numeric(n))
+  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
+    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
+    geom_line(aes(y=mean),col="blue")+
+    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
+    geom_linerange(data=tiepointranges,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
+  print(ggp)
+
+  # ggsave("sync-no4-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
+}
+```
+
 # Applications
@@
 ```{r events_individual, eval=FALSE}
+library(stringr)
 eventnumber=13 #number between 1 and 29. specifies which transition to consider
@@
                                     control.transition_dating=control.transition_dating)
+results_event0 = linrampfitter(results,control.linramp)
+results_event0 = events_depth_to_age(results_event0,
+                                    control.transition_dating=control.transition_dating)
+results_event2 = linrampfitter(results_sync2,control.linramp)
+results_event2 = events_depth_to_age(results_event2,
+                                    control.transition_dating=control.transition_dating)
 {
@@
 }
+
+{
+
+  tiepointranges_full = as.data.frame(matrix(NA,nrow=4,ncol=3))
+gicc05=results_sync$original.chron$age[1+1:n]
+gicc05tie = gicc05[results_sync$tie_points$locations_indexes]
+colnames(tiepointranges)=c("lower","upper","depth")
+for(i in 1:4){
+  tiepointranges_full[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",c(2,3,4,5)[i])]])-gicc05tie[i]
+}
+tiepointranges_full[,3] = depth[results_sync$tie_points$locations_indexes]
+colnames(tiepointranges_full) = c("lower","upper","depth")
+
+  ggd = data.frame(depth=results_sync2$data$depth,
+                   mean0=results$simulation$summary$mean-gicc05,
+                   lower0=results$simulation$summary$lower-gicc05,
+                   upper0=results$simulation$summary$upper-gicc05,
+                   mean=results_sync$simulation$summary_sync$mean-gicc05,
+                   lower=results_sync$simulation$summary_sync$lower-gicc05,
+                   upper=results_sync$simulation$summary_sync$upper-gicc05,
+                   mean2=results_sync2$simulation$summary_sync$mean-gicc05,
+                   lower2=results_sync2$simulation$summary_sync$lower-gicc05,
+                   upper2=results_sync2$simulation$summary_sync$upper-gicc05,
+                   zeros = numeric(n))
+  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
+    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
+    geom_ribbon(aes(ymin=lower0,ymax=upper0),fill="cyan2",col="cyan2",alpha=0.3)+
+    geom_line(aes(y=mean0),col="brown")+
+    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(y=mean),col="blue")+
+    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
+    ggtitle("(a) Simulated chronologies")+
+    geom_linerange(data=tiepointranges_full,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
+  print(ggp)
+
+  ggp1 = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
+    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
+    geom_ribbon(aes(ymin=lower2,ymax=upper2),fill="cyan2",col="cyan2",alpha=0.3)+
+    geom_line(aes(y=mean2),col="brown")+
+    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(y=mean),col="blue")+
+    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
+    ggtitle("(a) Synchronized chronologies")+
+    geom_linerange(data=tiepointranges_full,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
+  print(ggp1)
+
+   ggd1 = data.frame(depth=depth[interval],proxy=proxy[interval],
+                    x=results_event$linramp$data$x,
+                    mean=results_event$linramp$linrampfit$mean,
+                    lower=results_event$linramp$linrampfit$q0.025,
+                    upper=results_event$linramp$linrampfit$q0.975
+                    )
+  gg1 = ggplot(data=ggd1,aes(x=depth))+geom_line(aes(y=proxy),col="gray")+
+    theme_bw()+ 
+    #xlim(rev(range(depth[interval])))+
+    xlim(c(2161,2155.3))+
+    xlab("Depth (m)")+
+    ylab(expression(paste(delta^18,"O (permil)")))+ggtitle("(b) GI-11 Linear ramp fit")+
+    geom_ribbon(aes(x=x,ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(x=x,y=mean),col="blue",size=0.7)
+
+  ybottom = min(results_event$linramp$data$y)
+  ytop = min(results_event$linramp$linrampfit$q0.025)
+  margt0=results_event$linramp$param$t0$marg.t0
+  normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
+  gg1 = gg1 + geom_line(data=data.frame(x=margt0[,1],y=normt0.y),aes(x=x,y=y),col="brown",size=1.0)
+
+  gghist = ggplot(data = data.frame(ages=results_event$event_dating$samples)) + theme_bw()+
+    geom_histogram(aes(x=ages,y=..density..),fill="orange",col="black")+xlab("Onset age (yb2k)")+
+    ylab("Density") + ggtitle("(c) GI-11 Onset age") +
+    geom_vline(aes(xintercept = age.reference),lty=3,size=0.8)+
+    # geom_vline(aes(xintercept=results_event$event_dating$mean),col="blue",size=0.7)+
+    # geom_vline(aes(xintercept=results_event$event_dating$q0.025),col="red",size=0.7)+
+    # geom_vline(aes(xintercept=results_event$event_dating$q0.975),col="red",size=0.7)+
+    xlim(rev(range(results_event$event_dating$samples)))
+
+  library(stringr)
+
+ggl = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="Synchronous chronology mean"))+
+  geom_line(aes(x=x1,y=y1,col="Synchronous chronology 95% CI"))+
+  geom_line(aes(x=x2,y=y2,col="Unsynchronous chronology mean"))+
+  geom_line(aes(x=x,y=y,col="Unsynchronous chronology 95% CI"))+
+  geom_line(aes(x=x,y=y,col="Tie point 95% CI"))+
+  scale_color_manual("",values=c("Synchronous chronology mean"="blue",
+                                "Synchronous chronology 95% CI"="red",
+                                "Unsynchronous chronology mean"="brown",
+                                "Unsynchronous chronology 95% CI"="cyan2",
+                                "Tie point 95% CI"="magenta"),
+                       breaks=c("Synchronous chronology mean",
+                                "Synchronous chronology 95% CI",
+                                "Unsynchronous chronology mean",
+                                "Unsynchronous chronology 95% CI",
+                                "Tie point 95% CI")
+                                 )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        # legend.text = element_text(size = 17),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.size = unit(0.8,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(0.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+
+ggl1 = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="\n(a) Chronology mean\n(b) Linear ramp mean\n"))+
+  geom_line(aes(x=x1,y=y1,col="(a) Chronology 95% CI\n(b) Linear ramp 95% CI"))+
+  geom_line(aes(x=x2,y=y2,col="\n(a) Chronology mean, tie point 4 excluded\n(b) Onset posterior distribution\n"))+
+  geom_line(aes(x=x,y=y,col="(a) Chronology 95% CI, tie point 4 excluded"))+
+  geom_line(aes(x=x,y=y,col="\n(a) Tie point 95% CI\n"))+
+  geom_line(aes(x=x,y=y,col="(b) d18O data"))+
+  scale_color_manual("",values=c("\n(a) Chronology mean\n(b) Linear ramp mean\n"="blue",
+                                "(a) Chronology 95% CI\n(b) Linear ramp 95% CI"="red",
+                                "\n(a) Chronology mean, tie point 4 excluded\n(b) Onset posterior distribution\n"="brown",
+                                "(a) Chronology 95% CI, tie point 4 excluded"="cyan2",
+                                "\n(a) Tie point 95% CI\n"="magenta",
+                                "(b) d18O data"="gray"),
+                       breaks=c("\n(a) Chronology mean\n(b) Linear ramp mean\n",
+                                "(a) Chronology 95% CI\n(b) Linear ramp 95% CI",
+                                "\n(a) Chronology mean, tie point 4 excluded\n(b) Onset posterior distribution\n",
+                                "(a) Chronology 95% CI, tie point 4 excluded",
+                                "\n(a) Tie point 95% CI\n",
+                                "(b) d18O data")
+                                 )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.height=unit(0.3,"cm"),
+        # legend.key.size = unit(17,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(1.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+
+  #(a) (b) and (c) legends in same legend (maybe such that blue means something in a and something in b etc)
+  plots = c()
+  plots[[1]] = ggp
+  plots[[2]] = ggl
+  plots[[3]] = gg1
+  plots[[4]] = gghist
+
+  ggarrange(plotlist=plots,nrow=2,ncol=2)
+  # ggsave("multiplot-unsync-3000x2000.eps",device=cairo_ps,width=3000,height=2000,units="px",limitsize=FALSE)
+
+  plots1 = c()
+  plots1[[1]] = ggp1
+  plots1[[2]] = ggl1
+  plots1[[3]] = gg1
+  plots1[[4]] = gghist
+
+  ggarrange(plotlist=plots1,nrow=2,ncol=2,widths = c(0.67,0.5,0.5,0.5))
+  plotrow1 = c()
+  plotrow1[[1]] = ggp1
+  plotrow1[[2]] = ggl1
+  plotrow2 = c()
+  plotrow2[[1]] = gg1
+  plotrow2[[2]] = gghist
+  ggarow1 = ggarrange(plotlist=plotrow1,widths=c(0.7,0.3))
+  ggarow2 = ggarrange(plotlist=plotrow2)
+
+  ggarrange(c(ggarow1,ggarow2),nrow=2)
+
+  ggarrange(plotlist= c(ggarrange(plotlist=plotrow1,widths=c(0.67,0.33)),
+            ggarrange(plotlist=plotrow1,widths=c(0.5,0.5))) )
+  # ggsave("multiplot-sync-no4-2200x4000.eps",device=cairo_ps,height=2200,width=4000,units="px",limitsize=FALSE)
+  #
+
+}
+
+```
+
+```{r multifig-all, eval=FALSE}
+
+eventdepth = results_event$linramp$param$t0$mean
+  tiepointranges_full = as.data.frame(matrix(NA,nrow=4,ncol=3))
+gicc05=results_sync$original.chron$age[1+1:n]
+gicc05tie = gicc05[results_sync$tie_points$locations_indexes]
+colnames(tiepointranges)=c("lower","upper","depth")
+for(i in 1:4){
+  tiepointranges_full[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",c(2,3,4,5)[i])]])-gicc05tie[i]
+}
+tiepointranges_full[,3] = depth[results_sync$tie_points$locations_indexes]
+colnames(tiepointranges_full) = c("lower","upper","depth")
+
+  ggd = data.frame(depth=results_sync2$data$depth,
+                   mean0=results$simulation$summary$mean-gicc05,
+                   lower0=results$simulation$summary$lower-gicc05,
+                   upper0=results$simulation$summary$upper-gicc05,
+                   mean=results_sync$simulation$summary_sync$mean-gicc05,
+                   lower=results_sync$simulation$summary_sync$lower-gicc05,
+                   upper=results_sync$simulation$summary_sync$upper-gicc05,
+                   mean2=results_sync2$simulation$summary_sync$mean-gicc05,
+                   lower2=results_sync2$simulation$summary_sync$lower-gicc05,
+                   upper2=results_sync2$simulation$summary_sync$upper-gicc05,
+                   zeros = numeric(n))
+
+  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
+    coord_cartesian(ylim=c(range(ggd$lower0,ggd$upper))) +
+    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
+    geom_ribbon(aes(ymin=lower0,ymax=upper0),fill="black",col="black",alpha=0.3)+
+    geom_ribbon(aes(ymin=lower2,ymax=upper2),fill="red",col="red",alpha=0.3)+
+    geom_ribbon(aes(ymin=lower,ymax=upper),fill="blue",col="blue",alpha=0.3)+
+    geom_line(aes(y=mean0),col="black",size=0.9)+
+    geom_line(aes(y=mean2),col="red",size=0.9)+
+    geom_line(aes(y=mean),col="blue",size=0.9)+
+    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
+    ggtitle("(a) Simulated chronologies")+
+    geom_linerange(data=data.frame(depth=eventdepth),
+                   aes(x=depth,ymin=-500,ymax=1000),
+                   col="magenta",size=0.8)+
+    geom_linerange(data=tiepointranges_full,aes(x=depth,ymin=lower,ymax=upper),color="green",
+                   size=0.8)
+  print(ggp)
+
+  ggp1 = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
+    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
+    geom_ribbon(aes(ymin=lower2,ymax=upper2),fill="cyan2",col="cyan2",alpha=0.3)+
+    geom_line(aes(y=mean2),col="black")+
+    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(y=mean),col="blue")+
+    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
+    ggtitle("(a) Synchronized chronologies")+
+    geom_linerange(data=tiepointranges_full,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
+  print(ggp1)
+
+   ggd1 = data.frame(depth=depth[interval],proxy=proxy[interval],
+                    x=results_event$linramp$data$x,
+                    mean=results_event$linramp$linrampfit$mean,
+                    lower=results_event$linramp$linrampfit$q0.025,
+                    upper=results_event$linramp$linrampfit$q0.975,
+                    mean0=results_event0$linramp$linrampfit$mean,
+                    lower0=results_event0$linramp$linrampfit$q0.025,
+                    upper0=results_event0$linramp$linrampfit$q0.975,
+                    mean2=results_event2$linramp$linrampfit$mean,
+                    lower2=results_event2$linramp$linrampfit$q0.025,
+                    upper2=results_event2$linramp$linrampfit$q0.975
+                    )
+  gg1 = ggplot(data=ggd1,aes(x=depth))+geom_line(aes(y=proxy),col="gray")+
+    theme_bw()+ 
+    #xlim(rev(range(depth[interval])))+
+    xlim(c(2161,2155.3))+
+    xlab("Depth (m)")+
+    ylab(expression(paste(delta^18,"O (permil)")))+ggtitle("(b) GI-11 Linear ramp fit")+
+    geom_ribbon(aes(x=x,ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
+    geom_line(aes(x=x,y=mean),col="blue",size=0.7)
+
+  ybottom = min(results_event$linramp$data$y)
+  ytop = min(results_event$linramp$linrampfit$q0.025)
+  margt0=results_event$linramp$param$t0$marg.t0
+  normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
+  gg1 = gg1 + geom_line(data=data.frame(x=margt0[,1],y=normt0.y),aes(x=x,y=y),col="black",size=1.0)
+
+  ggdhistfull = data.frame(ages=results_event$event_dating$samples,
+                  ages0 = results_event0$event_dating$samples,
+                  ages2 = results_event2$event_dating$samples)
+  gghistfull = ggplot(data = ggdhistfull) + 
+    theme_bw()+
+    geom_histogram(aes(x=ages,y=..density..),fill="red",col="red",alpha=0.3)+
+    xlab("Onset age (yb2k)")+
+    geom_histogram(aes(x=ages2,y=..density..),fill="blue",col="blue",alpha=0.3)+
+    geom_histogram(aes(x=ages0,y=..density..),fill="black",col="black",alpha=0.3)+
+    ylab("Density") + ggtitle("(c) GI-11 Onset age") +
+    # geom_vline(aes(xintercept = age.reference),lty=3,size=0.8)+
+    # geom_vline(aes(xintercept=results_event$event_dating$mean),col="blue",size=0.7)+
+    # geom_vline(aes(xintercept=results_event$event_dating$q0.025),col="red",size=0.7)+
+    # geom_vline(aes(xintercept=results_event$event_dating$q0.975),col="red",size=0.7)+
+    xlim(rev(range(results_event$event_dating$samples,results_event0$event_dating$samples)))
+
+  gghist1 = ggplot(data = ggdhistfull) + 
+    theme_bw()+
+    geom_histogram(aes(x=ages,y=..density..),fill="orange",col="black",alpha=1)+
+    xlab("Onset age (yb2k)")+
+    ylab("Density") +
+    ggtitle("(i) Synchronized") +
+    xlim(rev(range(results_event$event_dating$samples,results_event0$event_dating$samples)))
+  gghist2 = ggplot(data = ggdhistfull) + 
+    theme_bw()+
+    geom_histogram(aes(x=ages2,y=..density..),fill="orange",col="black",alpha=1)+
+    xlab("Onset age (yb2k)")+
+    ylab("Density") +
+    ggtitle("(ii) Synchronized without tie point 4") +
+    xlim(rev(range(results_event$event_dating$samples,results_event0$event_dating$samples)))
+  gghist0 = ggplot(data = ggdhistfull) + 
+    theme_bw()+
+    geom_histogram(aes(x=ages0,y=..density..),fill="orange",col="black",alpha=1)+
+    xlab("Onset age (yb2k)")+
+    ylab("Density") +
+    ggtitle("(iii) Unsynchronized") +
+    xlim(rev(range(results_event$event_dating$samples,results_event0$event_dating$samples)))
+
+  histlist = c()
+  histlist[[1]] = gghist1
+  histlist[[2]] = gghist2
+  histlist[[3]] = gghist0
+  gghistlist = ggarrange(plotlist=histlist,nrow=3,ncol=1)
+
+  library(stringr)
+
+gglg1 = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n"))+
+  geom_line(aes(x=x1,y=y1,col="(a) Synchronous chronology mean and 95% CI, tie point 4 excluded\n(b) Linear ramp 95% CI"))+
+  geom_line(aes(x=x2,y=y2,col="\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset depth posterior distribution\n"))+
+  geom_line(aes(x=x,y=y,col="(a) Tie point 95% CI"))+
+  geom_line(aes(x=x,y=y,col="\n(b) GI-11 d18O data\n"))+
+  scale_color_manual("",values=c("\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n"="blue",
+                                "(a) Synchronous chronology mean and 95% CI, tie point 4 excluded\n(b) Linear ramp 95% CI"="red",
+                                "\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset depth posterior distribution\n"="black",
+                                "(a) Tie point 95% CI"="green",
+                                "\n(b) GI-11 d18O data\n"="gray"),
+                       breaks=c("\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n",
+                                "(a) Synchronous chronology mean and 95% CI, tie point 4 excluded\n(b) Linear ramp 95% CI",
+                                "\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset depth posterior distribution\n",
+                                "(a) Tie point 95% CI",
+                                "\n(b) GI-11 d18O data\n")
+                                 )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.height=unit(0.3,"cm"),
+        # legend.key.size = unit(17,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(1.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+gglg2 = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n"))+
+  geom_line(aes(x=x1,y=y1,col="(a) Synchronous chronology mean and 95% CI, \n      tie point 4 excluded\n(b) Linear ramp 95% CI"))+
+  geom_line(aes(x=x2,y=y2,col="\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset posterior distribution\n"))+
+  geom_line(aes(x=x,y=y,col="(a) Tie point 95% CI"))+
+  geom_line(aes(x=x,y=y,col="\n(a) Depth of event GI-11\n"))+
+  geom_line(aes(x=x,y=y,col="(b) GI-11 d18O data"))+
+  scale_color_manual("",values=c("\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n"="blue",
+                                "(a) Synchronous chronology mean and 95% CI, \n      tie point 4 excluded\n(b) Linear ramp 95% CI"="red",
+                                "\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset posterior distribution\n"="black",
+                                "(a) Tie point 95% CI"="green",
+                                "\n(a) Depth of event GI-11\n"="magenta",
+                                "(b) GI-11 d18O data"="gray"),
+                       breaks=c("\n(a) Synchronous chronology mean and 95% CI\n(b) Linear ramp mean\n",
+                                "(a) Synchronous chronology mean and 95% CI, \n      tie point 4 excluded\n(b) Linear ramp 95% CI",
+                                "\n(a) Unsynchronous chronology mean and 95%CI\n(b) Onset posterior distribution\n",
+                                "(a) Tie point 95% CI",
+                                "\n(a) Depth of event GI-11\n",
+                                "(b) GI-11 d18O data")
+                                 )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.height=unit(0.3,"cm"),
+        # legend.key.size = unit(17,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(1.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+
+  #(a) (b) and (c) legends in same legend (maybe such that blue means something in a and something in b etc)
+
+  plots = c()
+  plots[[1]] = ggp
+  plots[[2]] = gglg1
+  plots[[3]] = gg1
+  plots[[4]] = gghistfull
+
+  aaa = grid.arrange(ggp,gglg1,gg1,gghistlist,
+               layout_matrix=rbind(c(1,1,1,1,2,2),
+                                   c(3,3,3,4,4,4)))
+  bbb = arrangeGrob(ggp,gglg2,gg1,gghistlist,
+               layout_matrix=rbind(c(1,1,1,1,2,2),
+                                   c(3,3,3,4,4,4)))
+
+  #ggarrange(plotlist=plots,nrow=2,ncol=2)
+  ggsave("multiplot-new2-4000x3000.eps",plot=bbb,device=cairo_ps,width=4000,height=3000,units="px",limitsize=FALSE)
+
+  cairo_ps(filename = "multiplot.eps",
+         width = 7, height = 7, pointsize = 12,
+         fallback_resolution = 300)
+plot(aaa)
+dev.off()
+
+  plots1 = c()
+  plots1[[1]] = ggp
+  plots1[[2]] = gglg1
+  plots1[[3]] = gg1
+  plots1[[4]] = gghistlist
+
+  ggarrange(plotlist=plots1,nrow=2,ncol=2)
+  # ggsave("multiplot-new22-3500x2800.eps",device=cairo_ps,width=3500,height=2800,units="px",limitsize=FALSE)
+  #
+
 ```
@@
 capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`
-capind = numeric(29); buiind = numeric(29)
+corrick_onsets = read_excel("data/Corrick_onsets.xlsx")
+corrick_ages = corrick_onsets$`SIOC19 age`
+
+capind = numeric(29); buiind = numeric(29); corind = numeric(29)
+
 capind = c(NA,2,3, 4,5,6,NA,NA,7,8,NA,9,10,11,NA,NA,NA,NA,NA,12,13,14,NA,NA,15,NA,NA,16,17)
 buiind = c(NA,2,NA,3,4,6,NA,NA,8,9,NA,11,12,13,NA,NA,NA,NA,NA,15,16,17,NA,NA,19,NA,NA,20,21)
+corind = c(NA,1,NA,2,3,5,6,NA,8,9,10,11,12,13,NA,15,NA,16,NA,NA,NA,17,NA,19,20,21,NA,22,23)
+
+```
+
+```{r adolphilinrampfitter}
+adolphi_transfer = read.table("data/Adolphi_CP2018_SI.txt",skip=6,header=FALSE)
+adolphi_transfer[,1] = adolphi_transfer[,1] + 50
+adolphi_transfer_mean = adolphi_transfer[,2]
+adolphi_transfer_lower = adolphi_transfer[,5]
+adolphi_transfer_upper = adolphi_transfer[,6]
+
+ngrip_to_gicc05 = splinefun(results_sync$data$depth,results_sync$data$age) #interpolate age as function of depth
+
+maxdepth = 2181.87
+adolphidepth = results_sync$data$depth[results_sync$data$depth<maxdepth]
+
+adolphi_offset_mean_spline = function(depth){
+  gicc05_age = ngrip_to_gicc05(depth)
+  spline = splinefun(adolphi_transfer[,1], adolphi_transfer_mean) #as a function of gicc05 age
+  return(spline(gicc05_age))
+}
+adolphi_offset_lower_spline = function(depth){
+  gicc05_age = ngrip_to_gicc05(depth)
+  spline = splinefun(adolphi_transfer[,1], adolphi_transfer_lower) #as a function of gicc05 age
+  return(spline(gicc05_age))
+}
+adolphi_offset_upper_spline = function(depth){
+  gicc05_age = ngrip_to_gicc05(depth)
+  spline = splinefun(adolphi_transfer[,1], adolphi_transfer_upper) #as a function of gicc05 age
+  return(spline(gicc05_age))
+}
+
+adolphi_offset_mean = adolphi_offset_mean_spline(adolphidepth)
+adolphi_offset_lower = adolphi_offset_lower_spline(adolphidepth)
+adolphi_offset_upper = adolphi_offset_upper_spline(adolphidepth)
+
+ggd_adolphi = data.frame(depth=adolphidepth,mean=adolphi_offset_mean,lower=adolphi_offset_lower, upper=adolphi_offset_upper)
+
+gicc05 = results_sync$data$age
+ggd_sync = data.frame(depth = results_sync$data$depth, age=results_sync$data$age, mean=results_sync$simulation$summary_sync$mean-gicc05, lower = results_sync$simulation$summary_sync$lower-gicc05, upper=results_sync$simulation$summary_sync$upper-gicc05)
+
+tie_indexes = results_sync$tie_points$locations_indexes
+adolphi_tie_mean = adolphi_offset_mean_spline(results_sync$data$depth[tie_indexes])
+adolphi_tie_lower = adolphi_offset_lower_spline(results_sync$data$depth[tie_indexes])
+adolphi_tie_upper = adolphi_offset_upper_spline(results_sync$data$depth[tie_indexes])
+
+ggd_tie = data.frame(depths = results_sync$data$depth[tie_indexes], mean = results_sync$simulation$summary_sync$mean[tie_indexes]-gicc05tie, lower = results_sync$simulation$summary_sync$lower[tie_indexes]-gicc05tie, upper = results_sync$simulation$summary_sync$upper[tie_indexes]-gicc05tie)
+
+ggd_tie_adolphi = data.frame(depths = results_sync$data$depth[tie_indexes],
+                             mean = adolphi_tie_mean,
+                             lower = adolphi_tie_lower,
+                             upper = adolphi_tie_upper)
+
+ggp_adolphi =ggplot() + theme_bw() +
+  geom_ribbon(data=ggd_adolphi,aes(x=depth,ymin=lower,ymax=upper),fill="black",col="black",alpha=0.2) +
+  geom_ribbon(data=ggd_sync,aes(x=depth,ymin=lower,ymax=upper),fill="red",col="red",alpha=0.2) +
+
+  geom_line(data=ggd_adolphi,aes(x=depth,y=mean),col="black",size=1)+
+  geom_line(data=ggd_sync,aes(x=depth,y=mean),col="red",size=1)+
+  geom_segment(data=ggd_tie_adolphi,aes(x=depths,xend=depths,y=lower,yend=upper), 
+               col="black",size=0.7)+
+  geom_segment(data=ggd_tie,mapping=aes(x=depths,xend=depths,y=lower,yend=upper), 
+               col="red",size=0.7)+
+  xlab("NGRIP depth (m)") + 
+  ylab("Synchronous time scale - GICC05 (years)")
+
+plot(ggp_adolphi)
+
+adolphiage = results$data$age[results$data$age<=45000]
+meanspline = splinefun(adolphi_transfer[,1], adolphi_transfer_mean) 
+lowerspline = splinefun(adolphi_transfer[,1], adolphi_transfer_lower) 
+upperspline = splinefun(adolphi_transfer[,1], adolphi_transfer_upper)
+
+ggd_adolphi_age = data.frame(depth=adolphidepth, age = adolphiage,mean=meanspline(adolphiage), lower=lowerspline(adolphiage), upper=upperspline(adolphiage))
+
+ggp_adolphi_age =ggplot() + theme_bw() +
+  geom_ribbon(data=ggd_adolphi_age,aes(x=age,ymin=lower,ymax=upper),fill="black",col="black",alpha=0.2) +
+  geom_ribbon(data=ggd_sync,aes(x=age,ymin=lower,ymax=upper),fill="red",col="red",alpha=0.2) +
+
+  geom_line(data=ggd_adolphi_age,aes(x=age,y=mean),col="black",size=1)+
+  geom_line(data=ggd_sync,aes(x=age,y=mean),col="red",size=1)+
+  xlab("NGRIP depth (m)") + 
+  ylab("GICC05 age (yr b2k)")+
+  xlim(results_sync$data$age[1],45000)
+
+plot(ggp_adolphi_age)
+ggsave("adolphicomparison-3500x1500.eps",device=cairo_ps,width=3500,height=1500,units="px",limitsize=FALSE)
+#####
+# 
+# ylim = range(adolphi_offset_mean,adolphi_offset_upper,adolphi_offset_lower)
+# plot(adolphidepth,adolphi_offset_mean,col="blue",type="l",ylim=ylim)
+# lines(adolphidepth,adolphi_offset_lower,col="red")
+# lines(adolphidepth,adolphi_offset_upper,col="red")
+# 
+#   
+# adolphi_offset_lower_spline = splinefun(results_sync$data$depth, adolphi_transfer_lower) #as a function of depth
+# adolphi_offset_upper_spline = splinefun(results_sync$data$depth, adolphi_transfer_upper) #as a function of depth
+# 
+adolphi_age_to_offset = splinefun(adolphi_transfer$V1,adolphi_transfer$V2) #take inn GICC05 age, gives offset
+# 
+adolphi_depth_to_corrected = function(depth){
+
+  gicc05_age = ngrip_to_gicc05(depth)
+  corrected = gicc05_age + adolphi_age_to_offset( gicc05_age  )
+  return(corrected)
+}
+
+#  test = adolphi_depth_to_corrected(depth[age<45000])
+#  plot(test-age[age<45000])
+maxadolphidepth = max(depth[age<45000])
+
 ```
@@
 agesimmatrix_d18O_unsync = matrix(NA,29,nsims)
+agesimmatrix_d18O_no4 = matrix(NA,29,nsims)
+adolphimeans = rep(NA,29)
 depthlist_d18O = c()
@@
 for(eventnumber in 1:29){
+
   transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                       str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])
+  cat(eventnumber,". Event: ",transitionlabel,"\n",sep="")
   lowerints = which.index(event_intervals$depth_int_lower.m, results_sync$data$depth)
@@
   temp_d18O = linrampfitter(results_sync,control.linramp)
+  if(sum(temp_d18O$linramp$param$t0$marg.t0[,1]>maxadolphidepth)==0){
+    adolphimeans[eventnumber] = adolphi_depth_to_corrected(temp_d18O$linramp$param$t0$mean)
+  }
+  temp_d18O_no4 = linrampfitter(results_sync2,control.linramp)
   if(do.plot.rampfit){
@@
   agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)
+  agetemp_d18O_no4 = events_depth_to_age(temp_d18O_no4,control.transition_dating)
   agetemp_d18O_unsync = events_depth_to_age(temp_d18O,
@@
   agesimmatrix_d18O[eventnumber,] = agetemp_d18O$event_dating$samples 
+  agesimmatrix_d18O_no4[eventnumber,] = agetemp_d18O_no4$event_dating$samples 
   agesimmatrix_d18O_unsync[eventnumber,] = agetemp_d18O_unsync$event_dating$samples 
@@
     dens = data.frame(x=dens$x,y=dens$y)#/max(dens2$y))
+    dens2 = density(agesimmatrix_d18O_no4[i,])
+    dens2 = data.frame(x=dens2$x,y=dens2$y)#/max(dens2$y))
     densunsync = density(agesimmatrix_d18O_unsync[i,])
     densunsync = data.frame(x=densunsync$x,y=densunsync$y)#/max(dens2$y))
-    xlim = rev(range(dens$x,densunsync$x))
-
-    ggd = data.frame(x=dens$x,y=dens$y,x0=densunsync$x,y0=densunsync$y)
+    xlim = rev(range(dens$x,densunsync$x,dens2$x))
+    if(!is.na(adolphimeans[i])){
+      xlim=rev(range(xlim,adolphimeans[i]))
+    }
+
+    ggd = data.frame(x=dens$x,y=dens$y,x0=densunsync$x,y0=densunsync$y,
+                     x2=dens2$x,y2=dens2$y)
     ggdrasmus = data.frame(rasmus=event_intervals$GICC_age.yb2k[i])
@@
       geom_line(aes(x=x0,y=y0),col="gray",size=0.7)+
+      geom_line(aes(x=x2,y=y2),col="black",size=0.7,linetype="dashed")+
       geom_line(aes(x=x,y=y),col="black",size=0.7)+
@@
   }
+   if(!is.na(corind[i])){
+    ggdcorrick = data.frame(corrick=corrick_onsets$`SIOC19 age`[corind[i]])
+    gg = gg + geom_vline(data=ggdcorrick,aes(xintercept = corrick),col="brown",size=0.7)
+    #abline(v=buizert_onsets$Age[buiind[i]],col="green")
+    #abline(v=buizert_ages[buiind[i]],col="green")
+  }
+  if(!is.na(adolphimeans[i])){
+    ggdadolphi = data.frame(adolphi=adolphimeans[i])
+    gg = gg + geom_vline(data=ggdadolphi,aes(xintercept = adolphi),col="purple",size=0.7)
+    #abline(v=buizert_onsets$Age[buiind[i]],col="green")
+    #abline(v=buizert_ages[buiind[i]],col="green")
+  }
     # ggplots = c(ggplots,list(gg))
     ggplots[[i]] = gg
-  }
-
-  ggl=ggplot(data.frame(x = 1,x0=1, y = 1,y0=1, rasmus=1,buizert=1,capron=1))+
-    geom_line(aes(x=x,y=y,col="Synchronized onset posterior"),size=1)+
-    geom_line(aes(x=x0,y=y0,col="Unsynchronized onset posterior"))+
-    geom_line(aes(x=x,y=rasmus,col="Rasmussen onset"))+
-    geom_line(aes(x=x,y=buizert,col="Buizert onset"))+
-    geom_line(aes(x=x,y=capron,col="Capron onset"))+
-                        # colour = 'Something2'), aes(x, y, fill = colour))+
-  # geom_point(alpha=0, shape = 0)+ # completely transparent rectangular point 
-  # scale_fill_manual(values='black', drop=FALSE) +
-   # guides(color = guide_legend(override.aes = list(alpha=1, size = 1.3)))+ # showing the point in the legend
-    scale_color_manual(values=c("Synchronized onset posterior"="black",
+
+  }
+
+ggl = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="Synchronized onset posterior",linetype="Synchronized onset posterior"))+
+  geom_line(aes(x=x1,y=y1,col="Synchronized onset posterior (tie-point 4 excluded)",linetype="Synchronized onset posterior (tie-point 4 excluded)"))+
+  geom_line(aes(x=x2,y=y2,col="Unsynchronized onset posterior",linetype="Unsynchronized onset posterior"))+
+  geom_line(aes(x=x,y=y,col="Rasmussen et al. (2014)",linetype="Rasmussen et al. (2014)"))+
+  geom_line(aes(x=x1,y=y1,col="Buizert et al. (2015)",linetype="Buizert et al. (2015)"))+
+  geom_line(aes(x=x2,y=y2,col="Capron et al. (2021)",linetype="Capron et al. (2021)"))+
+  geom_line(aes(x=x,y=y,col="Corrick et al. (2020)",linetype="Corrick et al. (2020)"))+
+  geom_line(aes(x=x,y=y,col="Adolphi et al. (2018)",linetype="Adolphi et al. (2018)"))+
+  scale_color_manual("",values=c("Synchronized onset posterior"="black",
+                                "Synchronized onset posterior (tie-point 4 excluded)"="grey1",
                                 "Unsynchronized onset posterior"="gray",
@@
                                 "Buizert et al. (2015)"="green",
-                                "Capron et al. (2021)"="red"),
-                       labels=c("Synchronized onset posterior",
+                                "Capron et al. (2021)"="red",
+                                "Corrick et al. (2020)"="brown",
+                                "Adolphi et al. (2018)"="purple"),
+                       breaks=c("Synchronized onset posterior",
+                                "Synchronized onset posterior (tie-point 4 excluded)",
                                 "Unsynchronized onset posterior",
+                                "Adolphi et al. (2018)",
                                 "Rasmussen et al. (2014)",
                                 "Buizert et al. (2015)",
-                                "Capron et al. (2021)"))+
+                                "Capron et al. (2021)",
+                                "Corrick et al. (2020)")
+                                 )+
+  scale_linetype_manual("",values=c("Synchronized onset posterior"="solid",
+                                "Synchronized onset posterior (tie-point 4 excluded)"="dashed",
+                                "Unsynchronized onset posterior"="solid",
+                                "Rasmussen et al. (2014)"="solid",
+                                "Buizert et al. (2015)"="solid",
+                                "Capron et al. (2021)"="solid",
+                                "Corrick et al. (2020)"="solid",
+                                "Adolphi et al. (2018)"="solid"),
+                       breaks=c("Synchronized onset posterior",
+                                "Synchronized onset posterior (tie-point 4 excluded)",
+                                "Unsynchronized onset posterior",
+                                "Adolphi et al. (2018)",
+                                "Rasmussen et al. (2014)",
+                                "Buizert et al. (2015)",
+                                "Capron et al. (2021)",
+                                "Corrick et al. (2020)")
+                                    )+
   theme(axis.title = element_blank(),
@@
         legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
         # legend.key.size = unit(0.8,'cm'),
@@
   )
+
   ggplots[[30]] = ggl
-  ggarrange(plotlist=ggplots,widths=c(6,5))
-
-  # ggsave("allevents-4800x3000.eps",device=cairo_ps,width=4800,height=3000,units="px",limitsize=FALSE)
+  ggarrange(plotlist=ggplots,nrow=6,ncol=5,widths=c(6,5))
+
+  # ggsave("allevents-5200x6500.eps",device=cairo_ps,width=5200,height=6500,units="px",limitsize=FALSE)
+
+  ggl1 = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="Synchronized onset posterior",linetype="Synchronized onset posterior"))+
+  geom_line(aes(x=x1,y=y1,col="Synchronized onset posterior (tie-point 4 excluded)",linetype="Synchronized onset posterior (tie-point 4 excluded)"))+
+  geom_line(aes(x=x2,y=y2,col="Unsynchronized onset posterior",linetype="Unsynchronized onset posterior"))+
+  geom_line(aes(x=x,y=y,col="Adolphi et al. (2018)",linetype="Adolphi et al. (2018)"))+
+  scale_color_manual("",values=c("Synchronized onset posterior"="black",
+                                "Synchronized onset posterior (tie-point 4 excluded)"="grey1",
+                                "Unsynchronized onset posterior"="gray",
+                                "Adolphi et al. (2018)"="purple"),
+                       breaks=c("Synchronized onset posterior",
+                                "Synchronized onset posterior (tie-point 4 excluded)",
+                                "Unsynchronized onset posterior",
+                                "Adolphi et al. (2018)")
+                                 )+
+  scale_linetype_manual("",values=c("Synchronized onset posterior"="solid",
+                                "Synchronized onset posterior (tie-point 4 excluded)"="dashed",
+                                "Unsynchronized onset posterior"="solid",
+                                "Adolphi et al. (2018)"="solid"),
+                       breaks=c("Synchronized onset posterior",
+                                "Synchronized onset posterior (tie-point 4 excluded)",
+                                "Unsynchronized onset posterior",
+                                "Adolphi et al. (2018)")
+                                    )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(0.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        # legend.text = element_text(size = 17),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.size = unit(0.8,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(0.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+
+  ggl2 = ggplot(data.frame(x=c(1,2),y=c(2,3),x1=c(1,2),y1=c(1,2),x2=c(1,2),y2=c(1,2)),
+             x3=c(1,2),y3=c(2,3),x4=c(1,2),y4=c(1,2),x5=c(1,2),y5=c(1,2),x6=c(1,2),x7=c(1,2),
+             x8 = c(1,2), y8=c(1,2)) +
+  geom_line(aes(x=x,y=y,col="Rasmussen et al. (2014)",linetype="Rasmussen et al. (2014)"))+
+  geom_line(aes(x=x1,y=y1,col="Buizert et al. (2015)",linetype="Buizert et al. (2015)"))+
+  geom_line(aes(x=x2,y=y2,col="Capron et al. (2021)",linetype="Capron et al. (2021)"))+
+  geom_line(aes(x=x,y=y,col="Corrick et al. (2020)",linetype="Corrick et al. (2020)"))+
+  scale_color_manual("",values=c("Rasmussen et al. (2014)"="blue",
+                                "Buizert et al. (2015)"="green",
+                                "Capron et al. (2021)"="red",
+                                "Corrick et al. (2020)"="brown"),
+                       breaks=c("Rasmussen et al. (2014)",
+                                "Buizert et al. (2015)",
+                                "Capron et al. (2021)",
+                                "Corrick et al. (2020)")
+                                 )+
+  scale_linetype_manual("",values=c("Rasmussen et al. (2014)"="solid",
+                                "Buizert et al. (2015)"="solid",
+                                "Capron et al. (2021)"="solid",
+                                "Corrick et al. (2020)"="solid"),
+                       breaks=c("Adolphi et al. (2018)",
+                                "Rasmussen et al. (2014)",
+                                "Buizert et al. (2015)",
+                                "Capron et al. (2021)",
+                                "Corrick et al. (2020)")
+                                    )+
+  theme(axis.title = element_blank(),
+        axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        legend.position = c(1.5, 0.5), # move the legend to the center
+        legend.title = element_blank(),
+        # legend.text = element_text(size = 17),
+        legend.key = element_rect(fill='NA'),
+        legend.key.width=unit(1,"cm"),
+        # legend.key.size = unit(0.8,'cm'),
+        panel.grid = element_blank(),
+        # legend.spacing = unit(3,'cm'),
+        # legend.spacing.y = unit(0.5,'cm'),
+        panel.border = element_rect(colour = "white", fill='white', size=1)
+  )
+  ggplots2 = ggplots
+  ggplots2[[30]] = ggl1
+  ggplots2[[31]] = ggl2
+  ggarrange(plotlist=ggplots2,nrow=8,ncol=4,widths=c(6,5))
+
+  # ggsave("allevents2-4500x5500.eps",device=cairo_ps,width=4500,height=5500,units="px",limitsize=FALSE)
 
@@
 ```{r illustrationexample}
-rm(list=ls())
+# rm(list=ls())
 require(stats)
@@
 z0 = 0
-phi=0.8; sigma=1.2
+phi=0.8; sigma=15
 depth = z0+seq(from=1,to=n,length.out=n)
@@
                  upper = res$simulation$summary$upper-orichron)
-ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+ggd1_2 = data.frame(depth=res$data$depth,
+                 mean = res$simulation$summary$mean,
+                 lower = res$simulation$summary$lower,
+                 upper = res$simulation$summary$upper)
+ggpunc = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)
+ggpunc2 = ggplot(data=ggd1_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
   geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
@@
                  upper = res2$simulation$summary_sync$upper-orichron)
+ggd_2 = data.frame(depth=res2$data$depth,
+                 mean = res2$simulation$summary_sync$mean,
+                 lower = res2$simulation$summary_sync$lower,
+                 upper = res2$simulation$summary_sync$upper)
 
@@
                   mean=res2$simulation$summary_sync$mean[tiedepths]-oritie)
-
-ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+ggd2_2 = data.frame(depth=res2$tie_points$locations,
+                  mean=res2$simulation$summary_sync$mean[tiedepths])
+
+ggpfix = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
   geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
@@
   geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="magenta",size=1)
+ggpfix2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)+
+  geom_point(data=ggd2_2,mapping=aes(x=depth,y=mean),col="magenta",size=1)
+
 #plot(res)
@@
 tiemeans = c(450,3000,8500)
-tiesds = c(10,20,30)
+tiesds = c(100,500,200)
 res3 = tiepointsimmer(res,synchronization = list(locations=tiedepths,
@@
                  upper = res3$simulation$summary_sync$upper-orichron)
+ggd_2 = data.frame(depth=res3$data$depth,
+                 mean = res3$simulation$summary_sync$mean,
+                 lower = res3$simulation$summary_sync$lower,
+                 upper = res3$simulation$summary_sync$upper)
 
@@
                   upper=res3$simulation$summary_sync$upper[tiedepths]-oritie)
-
-ggp = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+ggd2_2 = data.frame(depth=res3$tie_points$locations,
+                  lower=res3$simulation$summary_sync$lower[tiedepths],
+                  upper=res3$simulation$summary_sync$upper[tiedepths])
+
+ggprand = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
   geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
@@
                col="magenta",size=0.7)
+ggprand2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
+  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
+  geom_line(aes(y=mean),col="blue",size=0.7)+
+  geom_segment(data=ggd2_2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
+               col="magenta",size=0.7)
 
 #plot(res3)
+ylim = range(res$simulation$summary$lower-orichron,res$simulation$summary$upper-orichron,
+             res3$simulation$summary_sync$lower-orichron,res3$simulation$summary_sync$upper-orichron)
+ggpunc = ggpunc + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")
+ggpfix = ggpfix + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")
+ggprand = ggprand + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")
+ggplotlist0 = c(); ggplotlist0[[1]] = ggpunc; ggplotlist0[[2]] = ggpfix; ggplotlist0[[3]] = ggprand
+ggplotlist00 = ggarrange(plotlist=ggplotlist0,nrow=3,ncol=1)
+
+ylim = range(res$simulation$summary$lower,res$simulation$summary$upper,
+             res3$simulation$summary_sync$lower,res3$simulation$summary_sync$upper)
+ggpunc2 = ggpunc2 + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")+ylab("Ensemble age")
+ggpfix2 = ggpfix2 + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")+ylab("Ensemble age")
+ggprand2 = ggprand2 + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")+ylab("Ensemble age")
+ggplotlist1 = c(); ggplotlist1[[1]] = ggpunc2; ggplotlist1[[2]] = ggpfix2; ggplotlist1[[3]] = ggprand2
+ggplotarrange = ggarrange(plotlist=ggplotlist1,nrow=3,ncol=1)
+plot(ggplotarrange)
+ggsave("illustrative-all-3000x3000.eps",device=cairo_ps,width=3000,height=3000,units="px")
+
 #ggsave("illustrative-sync.eps",device=cairo_ps,width=4800,height=2000,units="px")
@@
 ```{r simex}
-rm(list=ls())
+#rm(list=ls())
 require(stats)
@@
 n=1000
-y0 = 1000
-z0 = 1200
-phi=0.8; sigma=1.2
-depth = z0+seq(from=0.5,500,length.out=1000)
-depth2 = depth^2/z0 #normalize for stability
+y0 = 0
+z0 = 0
+phi=0.8; sigma=3
+
+depth = z0+seq(from=1,1000,length.out=1000)
+depth2 = depth^2#/z0 #normalize for stability
 
@@
 proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))
+
+truevals = c(sigma,phi,       0.5, 10, 0.01, 20, 0.02, 0, 0.03)
+namevals = c("\\sigma","\\phi","b_w","a_1","b_1","a_2","b_2","a_3","b_3")
+
 events=list(locations=c(1,250,750,1000))
-a1 = c(-rep(1,249)*15,numeric(751))
-a2 = c(numeric(249),rep(1,500)*50,numeric(251))
-a3 = c(numeric(749),-rep(1,251)*110)
-c1 = c(depth[1:249]/60,numeric(751))
-c2 = c(numeric(249),-depth[1:500]/40,numeric(251))
-c3 = c(numeric(749),depth[1:251]/10)
-dy = dnoise + a1+a2+a3+c1+c2+c3 + proxy*0.5
+a1 = c(rep(1,249),numeric(751))
+a2 = c(numeric(249),rep(1,500),numeric(251))
+a3 = c(numeric(749),rep(1,251))
+c1 = c(depth[1:249],numeric(751))
+c2 = c(numeric(249),depth[250:749],numeric(251))
+c3 = c(numeric(749),depth[750:1000])
+dy = dnoise + a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
+  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
+plot(dy)
 age = y0+cumsum(dy)
-df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth,depth2=depth2); data=rbind(c(y0,z0,0,0,0),df)
+df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth); data=rbind(c(y0,z0,NA,NA),df)
 formula=dy~-1+proxy#+depth2
 nsims = 10000
-synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,5600,11500),sd=c(10,50,60)))
+synchronization = list(locations = c(200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(2200,6800,19000),sd=c(10,50,60)))
 res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
+             print.progress=TRUE,
              events=list(locations=c(1,250,750,1000),locations_unit="index"),
@@
              synchronization=synchronization,
-             control.sim = list(synchronized=TRUE,
+             control.sim = list(synchronized=2,
                                 nsims=nsims))
@@
 
+  ggp
+
   ggd2 = data.frame(mean=res$simulation$summary_sync$mean-res$data$age,
@@
   ggtitle("(b) Simulated chronologies")
+
+ggp2
+
+ggarrange(ggp,ggp2,nrow=2)
+
 library(gridExtra)
-ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+# ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+dig=3
 str = ""
+str = paste0(str, " $",namevals[1],"$ & ",truevals[1]," & ", format(res$fitting$inla$hyperparameters$results$sigma_epsilon$mean,digits=dig,scientific=FALSE)," & (",format(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,digits=dig,scientific=FALSE), ", ",format(res$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975,difits=dig,scientific=FALSE),") \\\\ \n")
+
+str = paste0(str, " $",namevals[2],"$ & ",truevals[2]," & ", format(res$fitting$inla$hyperparameters$results$phi$mean,digits=dig,scientific=FALSE)," & (",format(res$fitting$inla$hyperparameters$results$phi$quant0.025,digits=dig,scientific=FALSE), ", ",format(res$fitting$inla$hyperparameters$results$phi$quant0.975,difits=dig,scientific=FALSE),") \\\\ \n")
+
 for(i in 1:length(res$fitting$inla$fit$summary.fixed$mean)){
-  str=paste0(str," & & ",format(res$fitting$inla$fit$summary.fixed$mean[i],digits=7,scientific=FALSE), " & (",
-             format(res$fitting$inla$fit$summary.fixed$`0.025quant`[i],digits=7,scientific=FALSE),", ",
-             format(res$fitting$inla$fit$summary.fixed$`0.975quant`[i],digits=7,scientific=FALSE),") \\\\ \n")
+  str=paste0(str, " $",namevals[i+2],"$ & ",truevals[i+2], " & ",format(res$fitting$inla$fit$summary.fixed$mean[i],digits=dig,scientific=FALSE), " & (",
+             format(res$fitting$inla$fit$summary.fixed$`0.025quant`[i],digits=dig,scientific=FALSE),", ",
+             format(res$fitting$inla$fit$summary.fixed$`0.975quant`[i],digits=dig,scientific=FALSE),") \\\\ \n")
 }
 cat(str)
+
 ```

### 2023-04-14 R/bremla.R

@@ -88,6 +88,7 @@
 #' plot(object)
 #'
 #' }
+#'
 #' if(inlaloader()){
 #' ### Real data example ###
 #' require(stringr)
@@ -118,7 +119,7 @@
 #' nsims=5000
 #' events=list(locations = events_rasmussen$depth,
 #'             locations_unit="depth",degree=1)
-#' synchronization = list(method="adolphi")
+#' synchronization = list(method="adolphi", locations = c(FALSE,TRUE,TRUE,TRUE,TRUE))
 #' control.fit=list(method="inla")
 #' control.sim=list(synchronized = 2)
 #'
@@ -141,7 +142,8 @@
 #' }
 #' @export
 #' @import matrixStats
-bremla <- function(formula,data,reference.label=NULL,
+bremla <- function(formula,data,
+                            reference.label=NULL,
                             x.label=NULL,
                             y.label=NULL,
                             nsims=10000,

### 2023-04-14 R/bremla_modelfitter.R

@@ -208,6 +208,10 @@ bremla_modelfitter = function(object, control.fit,
 
     my.control.fixed$prec=1
 
+    if(!object$.args$control.fit$improve.fixed){
+      my.control.fixed = NULL
+    }
+
     ## fit using INLA
     inlafit = INLA::inla(object$formula, family="gaussian",data=object$data,
                    control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))),

### 2023-04-14 R/default_arguments.R

@@ -46,6 +46,9 @@ events.default <- function(){
 #'   Only \code{inla} is currently supported, but more will hopefully come in the future.}
 #'   \item{\code{verbose} }{Boolean telling whether or not the \code{inla}-program should run in
 #'   verbose mode. Default \code{FALSE}.}
+#'   \item{\code{improve.fixed} }{Boolean indicating whether or not the least squares estimate of
+#'   the fixed values should be used as initial values for the optimization procedure in INLA.
+#'   This feature has not been extensively tested so it might be unstable. Default \code{FALSE}}
 #'   \item{\code{ncores} }{Integer used to set the number of cores to be used in the \code{inla} fitting
 #'   procedure. Default value is \code{1}.}
 #'   \item{\code{transform} }{Character indicating which transformation should be applied to
@@ -83,6 +86,7 @@ control.fit.default <- function(){
     method="inla",
     verbose=FALSE,
     log.theta.prior=NULL,
+    improve.fixed=FALSE,
     ncores=1,
     transform="identity",
     rgeneric = NULL

### 2023-04-14 R/helpfulfunctions.R

@@ -541,17 +541,25 @@ tiepointsimmer = function(object, synchronization,print.progress=FALSE,...){
     }
   }else{
     if(synchronization$method=="adolphi"){
+      tieinclude = synchronization$locations
+      if(is.null(tieinclude)){
+        tieinclude = rep(TRUE,5)
+      }
+      tieinclude[is.na(tieinclude)] = FALSE
+      tieinclude[ tieinclude != FALSE ] = TRUE
+
       adolphilocs = c(11050,12050,13050,22050,42050)
-      tieshifts = synchronization$locations
       samples = adolphi_tiepoint_simmer(nsims=synchronization$nsims,
                                         tieshifts=adolphilocs,...)
+      samples = samples[,!is.na(tieinclude)] #only use tie-points where location is not NA
+      adolphilocs = adolphilocs[!is.na(tieinclude)]
       synchronization$locations_unit="age"
       locations_indexes = which.index(adolphilocs,object$data$age)
       samples = samples[,!is.na(locations_indexes)]
       locations_indexes = locations_indexes[!is.na(locations_indexes)]
       synchronization$locations = object$data$age[locations_indexes]
       #synchronization$locations = synchronization$locations[!is.na(locations_indexes)]
-      synchronization$x.ref=tieshifts
+      synchronization$x.ref=adolphilocs
 
     }else if(synchronization$method %in% c("normal","gaussian","gauss")){
 

### 2023-07-05 inst/reproduce_results/test.Rmd
Fixed some bugs and changed the hyperparameter posterior sampler
no samples hyperparameters from joint distribution

---
output: github_document
---

<!-- results.md is generated from results.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/reproduce/results/results-",
  dev = "png",
  out.width = "100%"
)
```

# Extract DO events from Corrick pdf

```{r initialize, eval=TRUE}
library(pdftools)
library(data.table)
library(dplyr)
library(stringr)

# ttt = pdf_text("inst/reproduce_results/corricktable.pdf")
# 
# data = toString(ttt)
# data <- read.table(text = data, sep = "\n", as.is = TRUE)
# data <- data[-c(1, 2, 3, 4, 5, 6, 7, 14, 20, 26, 34, 47, 52, 58, 65, 66, 67), ]
# data <- gsub("[,]", "", data)
# data <- gsub("[$]", "", data)
# data <- gsub("\\s+", ",", gsub("^\\s+|\\s+$", "",data))
# 
# 
# 
# df <- ttt %>% 
#   read_lines() %>%    # separate lines
#   grep('^\\s{2}\\w', ., value = TRUE) %>%    # select lines with states, which start with space, space, letter
#   paste(collapse = '\n') %>%    # recombine
#   read_fwf(fwf_empty(.)) %>%    # read as fixed-width file
#   mutate_at(-1, parse_number) %>%    # make numbers numbers
#   mutate(X1 = sub('*', '', X1, fixed = TRUE))    # get rid of asterisks in state names
# 
# df

colnames = c("age", "avg error", "negative error"," positive error",
             "precision", "country", "region")
eventnames = c("GI 1e", "GI 3", "GI 4", "GI 5.1", "GI 5.2", "GI 6", "GI 7a",
               "GI 7c", "GI 8c", "GI 9", "GI 10", "GI 11","GI12c", "GI 13a", 
               "GI 13c", "GI14c", "GI 15.1", "GI 15.2", "GI 16.1a", "GI 16.1c",
               "GI 16.2", "GI 17.1a", "GI 17.1c", "GI 17.2", "GI 18", "GI 19.2",
               "GI 20c", "GI 21.1a", "GI 21.1e", "GI 22g", "GI 23.2", "GI 24.1a",
               "GI 24.1c", "GI 24.2", "GI 25c")

fullfile = scan("corrictabletext.txt", what="character",sep="\t")
findDOlines = gregexpr("GI",fullfile)
DOlines = numeric(length(eventnames))
teller=0
for(i in 1:length(findDOlines)){
  if(findDOlines[[i]][1] > -1){
    teller = teller+1
    DOlines[teller] = i
  }
}
DOlines = c(DOlines, length(fullfile)+1)

dflist = list()

#for(i in 19:19){
for(i in 2:length(DOlines)){  
  startline = DOlines[i-1]+1
  endline = DOlines[i]-1
  eventname = eventnames[i-1]
  rownames=c()
  rows = c()
  for(iterline in startline:endline){
    currentline = fullfile[iterline]
    elements = str_split(currentline," ")[[1]]
    elemarray = c()
    for(elem in 1:length(elements)){
      if(elements[elem] != ""){
        elemarray=c(elemarray, elements[elem])
      }
    }
    rownames = c(rownames, elemarray[1])
    rows=rbind(rows, elemarray[-1])
  }
  colnames(rows) = colnames
  rownames(rows) = rownames
  dflist[[eventname]] = rows
}

dflist[[1]]

```
```{r functions}

cleanstring = function(str,do.tolower=TRUE){
  str = gsub("\\s+|\\.|,|_|-", "", str, perl = TRUE)

  if(do.tolower){
    return(tolower(str))
  }else{
    return(str)
  }
}

stringcut = function(str,expr="_",before=TRUE,margin=1){
  pos <- regexpr(paste0("\\",expr), str)
  if(pos==-1){
    return(str)
  }else{
    if(before){
      return(substr(str,1,pos-margin))
    }else{
      return(substr(str,pos+margin,nchar(str)))
    }
  }
}
which.col = function(mat, val){
  loc=c()
  for(i in 1:ncol(mat)){
    ran = range(mat[!is.na(mat[,i]),i])
    if(ran[1] <= val && ran[2] >= val){
      loc=c(loc, i)
    }
  }
  if(length(loc)==0){
    warning("Could not find element in matrix columns!\n")
  }else if(length(loc)>1){
    warning("Element found in multiple columns...\n")
  }
  return(loc)
}
```

```{r dflist_bygroup}

codenames = c()
for(i in 1:length(dflist)){
  codenames = c(codenames, rownames(dflist[[i]]) )
}
codenames = unique(codenames)

records = c()

for(i in 1:length(codenames)){
  locstr = stringcut(codenames[i],expr="_",before=FALSE)
  groupstr = stringcut(codenames[i],expr="_",before=TRUE)
  groupstr = stringcut(groupstr,expr="(",before=TRUE)

  records = c(records, paste0(groupstr,"_",locstr))
}
records = unique(records)
nrecords = length(records)

fullfile = scan("corrictabletext.txt", what="character",sep="\t")

findDOlines = gregexpr("GI",fullfile)
DOlines = numeric(length(eventnames))

teller=0
for(i in 1:length(findDOlines)){
  if(findDOlines[[i]][1] > -1){
    teller = teller+1
    DOlines[teller] = i
  }
}
DOlines = c(DOlines, length(fullfile)+1)

#for(i in 19:19){
#for(i in 1:nrecords){  
  #findeventlines = gregexpr(tolower(records[i]),tolower(fullfile))
endline = length(fullfile)
df2list = list()
for(iterline in 4:endline){
  currentline = fullfile[iterline]
  elements = str_split(currentline," ")[[1]]

  elemarray = c()
  for(elem in 1:length(elements)){
    if(elements[elem] != ""){
      elemarray=c(elemarray, elements[elem])
    }
  }
  if("GI" %in% elemarray[1]){
  }else{
    tempdf = data.frame(age=elemarray[2], avgerror=elemarray[3], negerror=elemarray[4],
                      poserror=elemarray[5],precision=elemarray[6])

    eventind = max(which(DOlines<iterline))
    eventstr = eventnames[eventind]
    rownames(tempdf) = eventstr

    namestr = elemarray[1]
    groupstr = stringcut(namestr,"_",before=TRUE)
    groupstr = stringcut(groupstr,"(",before=TRUE)
    locstr = stringcut(namestr,"_",before=FALSE)
    recstring = paste0(groupstr,"_",locstr)

    df2list[[recstring]] = rbind(df2list[[recstring]],tempdf)

  }
  #}

}

df2list[[1]]

```

`
```{r rgeneric-linramp}
rgeneric.uneven.AR1 = function( #specifies necessary functions for INLA to define the linear ramp model
  cmd = c("graph", "Q","mu", "initial", "log.norm.const", "log.prior", "quit"),
  theta = NULL)
{

  envir = environment(sys.call()[[1]])

  linramp = function(t,t0=0,dt=1,y0=0,dy=1){ #linear ramp function
    y = numeric(length(t))
    y = y0 + dy*(t-t0)/dt

    y[t<t0]=y0
    y[t>t0+dt]=y0+dy

    return(y)
  }

  interpret.theta = function() { #helpful function to transform back from internal parametrization
    y0 = theta[3]
    dy = (theta[4])
    y1 = y0+dy
    t0 = theta[1]
    Dt = exp(theta[2])
    t1=t0+Dt

    prec = exp(theta[6])
    tau = exp(theta[5])
    return(list(y0=y0,dy=dy,y1=y1,t0=t0,t1=t1,prec=prec,tau=tau,Dt=Dt))
  }

  mu = function() { #mean vector defined as a linear ramp
    if(!is.null(envir)){
      timepoints=get("timepoints",envir)

    }
    param = interpret.theta()
    y0=param$y0; dy=param$dy; y1=param$y1; t0=param$t0; t1=param$t1; Dt = param$Dt; dt=Dt

    mvek = linramp(timepoints,t0=t0,dt=dt,y0=y0,dy=dy)

    return(mvek)
  }

  graph = function(){ #graphs of conditional dependence structure. 1 where Q[i,j] != 0, 0 where Q[i,j] = 0
    G = Q()
    G[G != 0] = 1
    return (G)
  }

  Q = function(){ #inverse covariance matrix: AR(1) that allow for unequal spacing
    if(!is.null(envir)){
      n=get("n",envir)
      timepoints=get("timepoints",envir)
    }
    param=interpret.theta()
    ii = 1:n
    jj = 1:n

    rhos=rep(NA,n)
    rhos[2:n] = exp(-diff(timepoints)/param$tau)
    kappa0 = param$prec

    xx = rep(NA,2*n-1)
    xx[1] = 1+rhos[2]^2/(1-rhos[2]^2)
    xx[2] = 1/(1-rhos[n]^2)
    xx[3:n] = 1/(1-rhos[2:(n-1)]^2) + rhos[3:n]^2/(1-rhos[3:n]^2)
    xx[(n+1):(2*n-1)] = -rhos[2:n]/(1-rhos[2:n]^2)

    xx = kappa0*xx

    i = c(1L, n, 2L:(n - 1L), 1L:(n - 1L))
    j = c(1L, n, 2L:(n - 1L), 2L:n)

    Q = Matrix::sparseMatrix(
      i = i,
      j = j,
      x = xx,
      symmetric = TRUE
    )
    #diag(Q)=diag(Q)

    return (Q)
  }

  log.norm.const = function(){ #INLA computes this automatically
    return(numeric(0))
  }

  log.prior = function(){
    if(!is.null(envir)){
      tslutt=get("tslutt",envir)
      tstart=get("tstart",envir)
      ystart=get("ystart",envir)
      #log.theta.prior=get("log.theta.prior",envir)
      # rescale.y=get("rescale.y",envir)
    }

    #if(!is.null(log.theta.prior)){
    #  lprior = log.theta.prior(theta)
    #}else{
      params = interpret.theta()

      #log-priors are given for internal parametrisation using the change of variables theorem

      lprior = dnorm(theta[1],mean=round(0.5*(tslutt+tstart)),sd=50,log=TRUE) #t0
      lprior = lprior + dgamma(exp(theta[2]),shape=1.0,rate=0.02,log=TRUE) + theta[2] #dt
      # if(rescale.y){
      #   lprior = lprior + dnorm(theta[3],mean=ystart,sd=0.025,log=TRUE) #y0
      #   lprior = lprior + dnorm(theta[4],mean=1-ystart,sd=0.01,log=TRUE) #dy
      #   lprior = lprior + dgamma(exp(theta[5]),1,rate = 3,log=TRUE) + theta[6] #tau/rho
      #   lprior = lprior + dgamma(exp(theta[6]),0.1,rate = 1,log=TRUE) + theta[6] #sigma/kappa
      # }else{
      lprior = lprior + dnorm(theta[3],mean=ystart,sd=5,log=TRUE) #y0
      lprior = lprior + dnorm(theta[4],mean=0,sd=10.0,log=TRUE) #dy
      lprior = lprior + dgamma(exp(theta[5]),2.5,rate = 0.15,log=TRUE) + theta[6] #tau/rho
      lprior = lprior + dgamma(exp(theta[6]),2,rate = 0.15,log=TRUE) + theta[6] #sigma/kappa
      # }
    #}

    return (lprior)
  }

  initial = function(){
    ini = c(0,0,0,0,1,2)
    return (ini)
  }

  quit = function(){
    return ()
  }
  if (!length(theta)) theta = initial()

  cmd = match.arg(cmd)
  val = do.call(cmd, args = list())
  return (val)
}
```

```{r linrampfunction}

linrampfunc = function(x,y, steplength=0.005,nsims=0, opt.params=NULL,imp.fit=TRUE, rescale.y.factor=1, print.progress=TRUE){

  n = length(x)

  input_x=x
  input_y=y
  if(print.progress) cat("Initializing linear ramp fit using INLA.\n",sep="")
  df_event = data.frame(xx=x, yy=y)

    #timepoints = input_x
  timepoints = round((df_event$xx - df_event$xx[1])/(df_event$xx[n]-df_event$xx[1])*(n-1)+1,digits=4)

    ## default initial values for optim function
  if(is.null(opt.params)){
    optparams = c(round(length(timepoints)/2),round(length(timepoints)/10),df_event$yy[1],df_event$yy[length(timepoints)]-df_event$yy[1])
  }else{
    defoptparams = c(round(length(timepoints)/2),round(length(timepoints)/10),df_event$yy[1],df_event$yy[length(timepoints)]-df_event$yy[1])
    optparams = control.linramp$opt.params
    for(i in 1:length(defoptparams)){
      if(is.na(optparams[i])){
        optparams[i] = defoptparams[i]
      }
    }
  }

  df0=data.frame(y=df_event$yy*rescale.y.factor,
                 x=df_event$xx)

  t_start = df0$time[1];t_end = df0$time[n];y_start = df0$y[1]
  #library(numDeriv)
  #library(compiler)
  y = df0$y
  if(print.progress) cat("Using 'optim' to find initial positions for hyperparameters in INLA.\n",sep="")
  ## for stability the x-axis is transformed to values ranging from 1:n.

  df0$time = timepoints

  ## Finding initial values in INLA optimization procedure by first using a simple optimization
  ## gradient and cost function given here

  if(!is.null(imp.fit)){
    minfun.grad = function(param, args = NULL){
      return (grad(minfun, param, args=args, method.args = list(r=6)))
    }
    minfun = function(param, args = NULL){
      yhat = linramp(timepoints,t0=param[1],dt=param[2],y0=param[3],dy=param[4])
      mse = sum((args$y-yhat)^2)
      return(sqrt(mse))
    }

    ## perform optimization to find good starting values for INLA
    param=optparams
    args=list(y=y)
    fit = optim(param,
                fn = minfun,
                gr = minfun.grad,
                method = "BFGS",
                control = list(
                  abstol = 0,
                  maxit = 100000,
                  reltol = 1e-11),
                args = args)

    ### use least squares estimates for fixed effects as initial values in inla

    muvek = linramp(timepoints,t0=fit$par[1],dt=fit$par[2],y0=fit$par[3],dy=fit$par[4])
    #plot(timepoints,y); lines(timepoints,muvek)
    init = c(fit$par[1],log(fit$par[2]),fit$par[3],fit$par[4],0,0)
  }

  if(print.progress) cat("Fitting linear ramp model in INLA using rgeneric model specification...\n",sep="")
  ## creating linear ramp INLA model using rgeneric framework. Requires further specification, see "rgeneric.uneven" function
  time.startinla = Sys.time()
  model.rgeneric = INLA::inla.rgeneric.define(rgeneric.uneven.AR1,n=n,
                                              tstart=timepoints[1],
                                              tslutt=timepoints[n],
                                              ystart=y[1],
                                              timepoints = timepoints)#,
                                              #log.theta.prior=control.linramp$log.theta.prior)
  formula = y ~ -1+ f(idx, model=model.rgeneric)

  if(!is.null(imp.fit)){
    r = INLA::inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=1:n),#as.integer(df0$time)),
                   control.mode=list(theta=init,
                                     restart=TRUE),
                   num.threads = 1,
                   verbose=FALSE,
                   silent=FALSE,
                   control.inla=list(h=steplength),
                   control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))) )#, num.threads = 1)

  }else{
    r = INLA::inla(formula,family="gaussian", data=data.frame(y=df0$y,idx=1:n),#as.integer(df0$time)),
                   control.mode=list(restart=TRUE),
                   num.threads = 1,
                   verbose=FALSE,
                   silent=FALSE,
                   control.inla=list(h=steplength),
                   control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))) )#, num.threads = 1)

  }

  time.endinla = Sys.time()
  elapsedinla = difftime(time.endinla,time.startinla,units="secs")[[1]]
  if(print.progress) cat("Completed in ",elapsedinla," seconds.\n",sep="")
  if(print.progress) cat("Gathering results...\n",sep="")

  object = list(timepoints=timepoints,data=df0,inlafit=r)
  object$linramp$data$y = object$linramp$data$y/rescale.y.factor

  ## compute posterior marginals and posterior marginal means of z^* = t0 (transition onset), dt (transition duration), y0 (initial ramp level), dy (change in ramp level) sigma (amplitude of AR(1) noise) and tau (parameter for correlation of AR(1) noise)
  t0=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta1 for idx`); dt=INLA::inla.emarginal(function(x)exp(x),r$marginals.hyperpar$`Theta2 for idx`);y0=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta3 for idx`); dy=INLA::inla.emarginal(function(x)x,r$marginals.hyperpar$`Theta4 for idx`);rho = INLA::inla.emarginal(function(x)2/(1+exp(-x))-1,r$marginals.hyperpar$`Theta5 for idx`); sigma = INLA::inla.emarginal(function(x)1/sqrt(exp(x)),r$marginals.hyperpar$`Theta6 for idx`)
  #muvek = linramp(timepoints,t0=t0,dt=dt,y0=y0,dy=dy)

  t0mean = r$summary.hyperpar$mean[1]; t0lower = r$summary.hyperpar$`0.025quant`[1];t0upper = r$summary.hyperpar$`0.975quant`[1]
  #margt0 = inla.tmarginal(function(x)df$age[1]+x/(n-1)*(df$age[n]-df$age[1]),r$marginals.hyperpar$`Theta1 for idx`);
  margt0 = INLA::inla.tmarginal(function(x)df0$x[1]+x/(n-1)*(df0$x[n]-df0$x[1]),r$marginals.hyperpar$`Theta1 for idx`);
  z.t0 = INLA::inla.zmarginal(margt0,silent=TRUE)

  object$linramp$param$t0 = list(marg.t0=margt0,mean=z.t0$mean,sd=z.t0$sd,q0.025=z.t0$quant0.025,q0.5=z.t0$quant0.5,q0.975=z.t0$quant0.975)
  if((abs(r$summary.hyperpar$mean[2])>1000) || (r$summary.hyperpar$sd[2]>1000)){
    margdt=NA
    margdtpos=NA
  }else{
    margdt = INLA::inla.tmarginal(function(x)exp(x)/(n-1)*(df0$x[n]-df0$x[1]),INLA::inla.smarginal(r$marginals.hyperpar$`Theta2 for idx`))
    margdtpos = data.frame(x=-margdt$x,y=margdt$y)
    z.dt = INLA::inla.zmarginal(margdt,silent=TRUE)
    z.dtpos = INLA::inla.zmarginal(margdtpos,silent=TRUE)
    object$linramp$param$dt = list(marg.dt=margdt,mean=z.dt$mean,sd=z.dt$sd,q0.025=z.dt$quant0.025,q0.5=z.dt$quant0.5,q0.975=z.dt$quant0.975)
    object$linramp$param$dtpos = list(marg.dtpos=margdtpos,mean=z.dtpos$mean,sd=z.dtpos$sd,q0.025=z.dtpos$quant0.025,q0.5=z.dtpos$quant0.5,q0.975=z.dtpos$quant0.975)
  }
  margy0 = INLA::inla.tmarginal(function(x)x/rescale.y.factor,
                                INLA::inla.smarginal(r$marginals.hyperpar$`Theta3 for idx`))
  margdy = INLA::inla.tmarginal(function(x)x/rescale.y.factor,
                                INLA::inla.smarginal(r$marginals.hyperpar$`Theta4 for idx`))
  z.y0 = INLA::inla.zmarginal(margy0,silent=TRUE)
  z.dy = INLA::inla.zmarginal(margdy,silent=TRUE)
  object$linramp$param$y0 = list(marg.y0=margy0,mean=z.y0$mean,sd=z.y0$sd,q0.025=z.y0$quant0.025,q0.5=z.y0$quant0.5,q0.975=z.y0$quant0.975)
  object$linramp$param$dy = list(marg.dy=margdy,mean=z.dy$mean,sd=z.dy$sd,q0.025=z.dy$quant0.025,q0.5=z.dy$quant0.5,q0.975=z.dy$quant0.975)

  margsigma = INLA::inla.tmarginal(function(x)1/sqrt(exp(x)),INLA::inla.smarginal(r$marginals.hyperpar$`Theta5 for idx`))
  margtau = INLA::inla.tmarginal(function(x)exp(x),INLA::inla.smarginal(r$marginals.hyperpar$`Theta6 for idx`))
  z.sigma = INLA::inla.zmarginal(margsigma,silent=TRUE)
  z.tau = INLA::inla.zmarginal(margtau,silent=TRUE)
  object$linramp$param$sigma = list(marg.sigma=margsigma,mean=z.sigma$mean,sd=z.sigma$sd,q0.025=z.sigma$quant0.025,q0.5=z.sigma$quant0.5,q0.975=z.sigma$quant0.975)
  object$linramp$param$tau = list(marg.sigma=margtau,mean=z.tau$mean,sd=z.tau$sd,q0.025=z.tau$quant0.025,q0.5=z.tau$quant0.5,q0.975=z.tau$quant0.975)

#nsims = 30000
  if(nsims>0 ) time.startbonussample = Sys.time()

  if(nsims>0 && print.progress ) cat("Simulating ensemble of ", nsims, " samples for t1 = t0 + dt..."," and linear ramp\n",sep="")

  if(nsims>0){

    samps=INLA::inla.hyperpar.sample(nsims,r)

    hpars = matrix(NA,nrow = nsims,ncol=5)
    hpars[,1:2]=samps[,3:4]/rescale.y.factor #y0,dy
    n=length(timepoints)

    hpars[,3] = df0$x[1] + (samps[,1]-1)/(n-1)*(df0$x[n]-df0$x[1]) #t0
    hpars[,4] = exp(samps[,2])/(n-1)*(df0$x[n]-df0$x[1]) #Dt
  }

  if(nsims>0){
    t1sims=numeric(nsims)
    t1mean = mean(t1sims)
    for(i in 1:nsims){
      t01 = hpars[i,3]
      dt1 = hpars[i,4]
      t1sims[i] = t01+dt1

    }

    t1dens = density(t1sims)
    margt1 = cbind(t1dens$x,t1dens$y); colnames(margt1) = c("x","y")
    z.t1 = INLA::inla.zmarginal(margt1,silent=TRUE)
    object$linramp$param$t1 = list(marginal=margt1,samples = t1sims,mean=z.t1$mean,sd=z.t1$sd,q0.025=z.t1$quant0.025,q0.5=z.t1$quant0.5,q0.975=z.t1$quant0.975)
  }
  if(nsims>0){
    vekmat = matrix(NA,nrow=n,ncol=nsims)
    for(i in 1:nsims){
      t01 = hpars[i,3]
      dt1 = hpars[i,4]
      vekmat[,i] = bremla::linramprev(object$data$x,t0=t01,dt=dt1,y0=hpars[i,1],dy=hpars[i,2])
    }
    vek.quant0.025 = numeric(n)
    vek.quant0.5 = numeric(n)
    vek.quant0.975 = numeric(n)
    vek.mean = numeric(n)
    for(iter in 1:n){
      dens = density(vekmat[iter,])
      vek.quant0.025[iter]=INLA::inla.qmarginal(0.05,dens)
      vek.quant0.5[iter]=INLA::inla.qmarginal(0.5,dens)
      vek.mean[iter] = mean(vekmat[iter,])
      vek.quant0.975[iter]=INLA::inla.qmarginal(0.95,dens)
    }

    object$linramp$linrampfit = list(mean = vek.mean,q0.025=vek.quant0.025,q0.5=vek.quant0.5,q0.975=vek.quant0.975)
  }
  #object$time$linramp = list(inla=elapsedinla)
  if(nsims>0 ) object$time$t1_and_ramp = difftime(Sys.time(),time.startbonussample,units="secs")[[1]]
  if((nsims>0 ) && print.progress) cat(" completed in ",object$time$t1_and_ramp," seconds!\n",sep="")

  #time.total = difftime(Sys.time(),time.start,units="secs")[[1]]

  object$linramp$.args = list(nsims=nsims)
  #object$time$linramp=time.total
return(object)
  }

```

```{r linrampplot}
library(ggplot2)
linrampplot = function(object,mainlab="",sublab="", ref=NULL){

  ggd = data.frame(x=object$data$x,y=object$data$y, mean = object$linramp$linrampfit$mean, lower=object$linramp$linrampfit$q0.025,upper=object$linramp$linrampfit$q0.975)

  ggp = ggplot(data=ggd) + geom_line(aes(x=x,y=y),col="gray")+
    geom_ribbon(aes(x=x,ymin=lower,ymax=upper),fill="red",alpha=0.3) + geom_line(aes(x=x,y=mean),col="blue") + theme_bw() + xlab("Depth") + ylab("d18O")+
    labs(title=mainlab, subtitle=sublab) + scale_x_reverse()
  if(!is.null(ref)){
    ggref = data.frame(ref=ref)
    ggp = ggp + geom_vline(data=ggref,aes(xintercept=ref),color="black")
  }

  return(ggp)
}
```

# Fitting linramp model

```{r nextblock}
library(readxl)
library(bremla)
library(numDeriv)

finetunenames = function(strings,tolower=FALSE){
  cstrings = c()
  for(i in 1:length(strings)){
    str = strings[i]
    #if(str_detect(str," ")){
    str = stringcut(str,expr=" ",before=FALSE)
  #}
    if(str_detect(str,"ö")){
      str <- gsub("ö", "o", str)
    }
    cstrings[i] = str
  }

  if(tolower) return(tolower(cstrings))
  return(cstrings)
}

sheetnames = excel_sheets(path="corricdata.xlsx")[-1]
sheetnames = finetunenames(sheetnames)

rowskips = rep(2,length(sheetnames))
rowskips[c(3,8,15,21)] = 3
rowskips[c(11,38)] = 4

#names(df2list)

errorvec = numeric(length(df2list)) #index represents iterator 'i' below. Value represent which event the error occured

margt0_x = c()
margt0_y = c()
recordindices= c()
recordname = c()
eventindices = c()
eventnames = c()

counter = 0
for(i in 1:length(df2list)){

  nam = names(df2list)[i]
  cat("Iteration #",i," of ",length(df2list),": ",nam,"\n\n",sep="")

  locstring = stringcut(nam,"_",before=FALSE)
  groupstring = stringcut(nam,"_",before=TRUE)

  if(locstring %in% sheetnames){
    sheetid = which(tolower(sheetnames) == tolower(locstring))+1
  }else{
    warning(paste0("Could not find correct sheet. ",i,": ",locstring,"\n"))
  }
  groupnames = colnames(read_excel("corricdata.xlsx",sheet=sheetid,n_max=1))[-1]
  groupnames = groupnames[!str_detect(groupnames,"\\.\\.\\.")]
  if(groupstring %in% groupnames){
  }else{
    warning(paste0("Could not find correct group. ",i,": ",locstring," - ",groupstring,"\n"))
  }

  dd = read_excel("corricdata.xlsx",sheet=sheetid,skip=rowskips[sheetid-1],trim_ws=TRUE)
  emptycols = rep(FALSE,length(colnames(dd)))
  emptycolslist = gregexpr("\\.\\.\\.", colnames(dd))
  for(j in 1:length(emptycolslist)){
    if(emptycolslist[[j]][1] == 1)
      emptycols[j] = TRUE
  }
  dd = dd[,!emptycols]
  cnames = colnames(dd)

  agevecs = dd[,str_detect(tolower(cnames),"age_1950")]
  depthvecs = dd[,str_detect(tolower(cnames),"depth")]
  d18Ovecs = dd[,str_detect(tolower(cnames),"d18o")]

  inevents = nrow(df2list[[i]])
  ineventnames = rownames(df2list[[i]])

  for(eventid in 1:inevents){
    counter = counter+1
    ineventname = ineventnames[eventid]
    cat("\n\n",counter,") event #",eventid," of ",inevents,": ",ineventname,"\n\n",sep="")

    eventage = as.numeric(df2list[[i]]$age[eventid])
    colid = which.col(agevecs, eventage)
    if(length(colid)==0){
      errorvec[i] = eventid
      cat("Breaking inner for-loop\n")
      break
      cat("Didnt work!\n")
    }
    if(length(colid)>1){
      warning("Onset found in multiple records, choosing the one where event is closest to middle in terms of indices...\n")
      mincolid = colid[1]
      bestplacement = Inf
      for(ci in 1:length(colid)){
        agve = agevecs[[ colid[ci] ]][!is.na(agevecs[[ colid[ci] ]])]
        deve = depthvecs[[ colid[ci] ]][!is.na(depthvecs[[ colid[ci] ]])]
        if(deve[1] > tail(deve,1)){ #choosing depths that are increasing
        }else{
          #1/length(agve)*sqrt(sum((eventage-agve)^2))

        placement = which(sort(unique(c(agve,eventage)))==eventage) / length(agve)

        offcenter = abs(placement-0.5)

        if(offcenter < bestplacement){
          mincolid = colid[ci]
          bestplacement = offcenter
        }
        }
      }
      if(bestplacement == Inf){
        stop("Could not find suitable record (increasing depths\n")
      }
      colid = mincolid

    }

    depthvec = as.numeric(depthvecs[colid][[1]])
    d18Ovec = as.numeric(d18Ovecs[colid][[1]])
    agevec = as.numeric(agevecs[colid][[1]])

    removeind = c(which(is.na(depthvec)),which(is.na(d18Ovec)),which(is.na(agevec)))
    removeind = unique(removeind)
    if(length(removeind)>0){
      depthvec = depthvec[-removeind]
      agevec = agevec[-removeind]
      d18Ovec = d18Ovec[-removeind]
    }

    if(depthvec[1]>tail(depthvec,1)){
      warning("depth is decreasing. Stopping\n")
      errorvec[i] = colid
      break
    }

    #faultydepths = which(diff(depthvec)<=0)+1
    #faultydepths = which(abs(diff(depthvec))>1)+1
    is.true=TRUE
    while(is.true){
      faultydepths1 = which(diff(depthvec)<=0)+1
      #faultydepths2 = which(abs(diff(depthvec))>1)+1
      #faultydepths = unique(c(faultydepths1,faultydepths2))
      faultydepths = unique(c(faultydepths1))
      if(length(faultydepths)>=1){
        depthvec = depthvec[-faultydepths]
        agevec = agevec[-faultydepths]
        d18Ovec = d18Ovec[-faultydepths]
      }else{
        is.true=FALSE
      }
    }

    #par(mfrow=c(1,2))
    #plot(agevec, d18Ovec,type="l", xlim=rev(range(agevec)))
    #abline(v=eventage)

    ###
    ### insert segmentation ###
    ###

    n = length(agevec)
    eventind = which.index(eventage, agevec)
    upper = min(eventind + round(n/20), n)
    lower = max(eventind - round(n/10), 1)

    interval = rev(seq(from=lower,to=upper,by=1))
    xx = depthvec[interval]
    yy = d18Ovec[interval]
    #plot(xx, yy,type="l",xlim=rev(range(xx)))
    #abline(v=depthvec[eventind],lwd=2,col="blue")

    ### linramp fitting ###

    #remember inversion if needed

    rescale.y.factor = -40/mean(yy)

    results = linrampfunc(xx,yy, steplength=0.005,nsims=5000, opt.params=NULL,imp.fit=TRUE, rescale.y.factor=1, print.progress=TRUE)

    ## temporary placed plot here to give the other plot more time
    plot(agevec, d18Ovec,type="l", xlim=rev(range(agevec)))
    abline(v=eventage)
    #####

    eventname = rownames(df2list[[i]])
    ggp = linrampplot(results,mainlab=ineventname,sublab=nam,ref=depthvec[eventind])
    plot(ggp)
  ### ### ###

    margt0_x = cbind(margt0_x,matrix(results$linramp$param$t0$marg.t0[,1],ncol=1))
    margt0_y = cbind(margt0_y,matrix(results$linramp$param$t0$marg.t0[,2],ncol=1))

    recordindices= c(recordindices, i)
    recordname = c(recordname, nam)
    eventindices = c(eventindices, eventid)
    eventnames = c(eventnames, eventname)

    ggsave(filename=file.path("figs",paste0(counter,"_recordnumber",i,"_event",eventid,"_plot-4800x2400.eps")), device=cairo_ps,width=4800,height=2400,units="px",dpi=500,limitsize=FALSE)

  }

  plot(margt0_x[,1],margt0_y[,1])

}

```

### 2023-07-05 R/bremla.R

@@ -95,7 +95,7 @@
 #' depth = NGRIP_5cm$depth
 #' d18O = NGRIP_5cm$d18O
 #' proxy=d18O
-#' formula = dy~-1+depth2
+#' formula = dy~-1+depth2 + proxy
 #' depth2 = depth^2/depth[1]^2 #normalize for stability
 #'
 #' data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)

### 2023-07-05 R/bremla_synchronized_simulation.R

@@ -125,83 +125,66 @@ time.start = Sys.time()
   free_n = n-m
   free_indexes = (1:n)[-tie_indexes]
 
-  if(tolower(object$.args$control.fit$method) == "inla" && is.null(object$simulation)){ #if INLA is used
+  if(tolower(object$.args$control.fit$method) == "inla"){ # && is.null(object$simulation)){ #if INLA is used
+
     reg.model = object$.internal$lat.selection
-
-    if(print.progress) cat("Simulating ",nsims, " hyperparameters from INLA posterior...",sep="")
-
-    if(tolower(noise) %in% c("rgeneric","custom")){
-      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
-      param.names = object$.args$control.fit$rgeneric$param.names
-      for(i in 1:length(object$.args$control.fit$rgeneric$from.theta)){
-        paramsamp = object$.args$control.fit$rgeneric$from.theta[[i]](hypersamples[,i])
-        if(is.null(param.names[i]) || is.na(param.names[i])){
-          tempname = paste0("hyperparameter",i)
-          object$simulation[[tempname]] = paramsamp
-        }else{
-          object$simulation[[param.names[i] ]] = paramsamp
-
-        }
-      }
-    }else if(tolower(noise) %in% c(0,"ar(0)","ar0","iid","independent")){
-      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
-      object$simulation = list(sigma = 1/sqrt(hypersamples[,1]))
-    }else if (tolower(noise) %in% c(1,"ar1","ar(1)")){
-      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
-      object$simulation = list(sigma = 1/sqrt(hypersamples[,1]), phi=hypersamples[,2])
-
-    }else if (tolower(noise) %in% c(2,"ar2","ar(2)")){
-      hypersamples = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
-      p=2
-      hypersamplesar2 = INLA::inla.hyperpar.sample(nsims,object$fitting$inla$fit)
-      phii = hypersamplesar2[, 2L:(2L+(p-1L))]
-      phis = apply(phii, 1L, INLA::inla.ar.pacf2phi)
-      object$simulation = list(sigma = 1/sqrt(hypersamples[,1]),phi1=phis[1,],phi2=phis[2,])
-    }
[…]

### 2023-07-05 R/default_arguments.R

@@ -87,9 +87,8 @@ control.fit.default <- function(){
   return(list(
     noise="ar1",
     method="inla",
-    hyperprior = NULL,
     verbose=FALSE,
-    log.theta.prior=NULL,
+    hyperprior = NULL,
     improve.fixed=FALSE,
     ncores=1,
     transform="identity",

### 2023-08-01 inst/reproduce_results/results.Rmd
Fixed subsetting bug
Caused problems in windows

@@
 res3 = bremla_synchronized_simulation(res3,control.sim=list(synchronized=TRUE))
+
 ggd = data.frame(depth=res3$data$depth,
@@
 plot(ggplotarrange)
-ggsave("illustrative-all-3000x3000.eps",device=cairo_ps,width=3000,height=3000,units="px")
+
+#ggsave("illustrative-all-3000x3000.eps",device=cairo_ps,width=3000,height=3000,units="px")
+ggsave("illustrative-all-3000x3400.eps",device=cairo_ps,width=3000,height=3400,units="px",dpi=500,limitsize=FALSE)
+
+#ggsave("illustrative-all-3000x4000.eps", device=cairo_ps,width=3000,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
 
@@
     geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.8)+
-    xlab("Depth (m)")+ylab("Layers per 5cm")+ggtitle("(a) Fitted layer increment model")
+    xlab("Depth (m)")+ylab("Layers per 1m")+ggtitle("(a) Fitted layer increment model")
 
@@
   geom_linerange(data=ggdtie,aes(x=locations,ymin=lower,ymax=upper),color="magenta")+
-  xlab("Depth (m)") + ylab("Deviation from simulated reference (years)")+
+  xlab("Depth (m)") + ylab("Age deviation (years)")+
   ggtitle("(b) Simulated chronologies")
@@
 library(gridExtra)
+
 # ggsave("fitdiffsim-5500x4000.eps", device=cairo_ps,width=5500,height=4000,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+
+ggsave("fitdiffsim-3000x2200.eps", device=cairo_ps,width=3000,height=2200,units="px",dpi=500,limitsize=FALSE,arrangeGrob(ggp,ggp2))
+
 dig=3

### 2023-08-01 R/helpfulfunctions.R

@@ -641,7 +641,11 @@ Qsimmer = function(nsims, Q, muvek){
 
   for(i in 1:nsims){
     samples_z = rnorm(nn)
-    v = solve(La,samples_z)[,1]
+    v = solve(La,samples_z)
+    if(!is.null(dim(v))){
+      v = v[,1]
+    }
+
     samples[,i] = muvek[(length(muvek)-nn+1):length(muvek)] + v
   }
   return(samples)

### 2023-08-01 R/rgeneric.fitting.R

@@ -54,24 +54,23 @@
 #'
 #' # R-INLA framework example
 #' n = 300
+#' depth = 1:n
 #' ntheta = 1
-#' dy = rnorm(n)+1:n/40
+#' dy = rnorm(n)+depth/40
 #' model.rgeneric = inla.rgeneric.define(rgeneric.fitting, n=n,ntheta = 1,
 #'                                       Q = Q,
 #'                                       log.prior=log.prior)
 #'
 #' formula = dy ~ -1+depth+ f(idy, model=model.rgeneric)
 #'
-#' result = inla(formula,family="gaussian", data=data.frame(dy=dy,depth=1:n,idy=1:n),
+#' result = inla(formula,family="gaussian", data=data.frame(dy=dy,depth=depth,idy=1:n),
 #'               num.threads = 2,
 #'               control.family = list(hyper = list(prec = list(initial = 12, fixed=TRUE))) )
 #'
 #'
+#'
 #' ## bremla framework example
 #'
-#' n = 300
-#' depth = 1:n
-#' dy = rnorm(n) + depth/40
 #' y = cumsum(dy)
 #'
 #' formula = dy ~ -1+depth

### 2025-01-17 inst/reproduce_results_BA/appendix-stanexample.Rmd
Now supports systematic biases
Systematic biases can now be described using an RW2 process. Also added other stuff, including code to reproduce the results of the most recent paper

# Simulation example in Section 9
This codes generates synthetic layer increments and tie-points and fits the layer-increment and the age-discrepancy models jointly using Markov chain Monte Carlo. Specifically, we use the Stan implementation of the No-U-turn Sampler.

This approach is much more computationally intensive than using the simplification described in the paper along with R-INLA. Hence, the synthetic time series is much shorter.

## Simulate data

```{r}
library(rstan)
library(Matrix)
library(devtools)
library(ggplot2)
set.seed(123)
n = 100

set.seed(123)
y0 = 0
z0 = 0
phi=0.8; sigma=3

depth = z0+seq(from=1,n,length.out=n)
depth2 = depth^2#/z0 #normalize for stability

dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))

truevals = c(sigma,phi,       0.5, 10, 0.01, 20, 0.02, 0, 0.03)
namevals = c("\\sigma","\\phi","b_w","a_1","b_1","a_2","b_2","a_3","b_3")

locs = c(1,25, 75, n)

events=list(locations=locs)
segs = diff(locs); segs[3] = segs[3]+1
a1 = c(rep(1,segs[1]),numeric(segs[2]+segs[3]))
a2 = c(numeric(segs[1]),rep(1,segs[2]),numeric(segs[3]))
a3 = c(numeric(segs[1]+segs[2]),rep(1,segs[3]))
c1 = c(depth[1:(locs[2]-1)],numeric(segs[2]+segs[3]))
c2 = c(numeric(segs[1]),depth[locs[2]:(locs[3]-1)],numeric(segs[3]))
c3 = c(numeric(segs[1]+segs[2]),depth[locs[3]:locs[4]])
dy = dnoise + a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
truetrend = a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
#plot(dy)
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth); data=rbind(c(y0,z0,NA,NA),df)
formula=dy~-1+proxy#+depth2
# formula = dy ~ -1 + a1+a2+a3+c1+c2+c3 + proxy
nsims = 10000

# lmr = lm(formula,data=data.frame(dy=dy,proxy=proxy,a1=a1,a2=a2,a3=a3,c1=c1,c2=c2,c3=c3))

res = bremla_prepare(formula, data,nsims, events=events)

design_matrix <- model.matrix(res$.args$formula.ls, res$data)
betatrues = c(truevals[3],truevals[4],truevals[5],truevals[6],truevals[7],truevals[8],truevals[9])

fixedtrue = design_matrix%*%betatrues
#dy = design_matrix%*%betatrues + dnoise

design_matrix <- model.matrix(res$.args$formula.ls, res$data)

{
  ggd = data.frame(true=fixedtrue, dy=dy,depth=depth)
  ggplot(ggd, aes(x=depth)) + theme_bw() + xlab("Depth") + ylab("Layer increments")+
    geom_point(aes(y=dy),col="gray")+
    geom_line(aes(y=fixedtrue),col="black")
}

set.seed(1)
y_truth = cumsum(dy) # just to check where tie-points should be placed

bias = 500*sin((1:n)*3*pi/(2*n))
y_full = y_truth+bias

#which_obs = sort(sample(1:n,n/5))
#which_obs = sort(sample(1:n,n-50))
which_obs = sort(c(10,25,50,75,90))
is_missing = rep(TRUE,n)
is_missing[which_obs] = FALSE
which_missing = which(is_missing)
is_obs = !is_missing
y_full[which_missing]=NA
y_obs = y_full[which_obs]

x_y = seq(from=1,to=10,length.out=n)
stan_data <- list(
  N = n,
  K = ncol(design_matrix),
  X = design_matrix,
  dy = dy,
  N_obs = length(which_obs),
  N_missing = length(which_missing),
  which_missing = which_missing,
  which_obs = which_obs,
  y_obs = y_obs#,
)

```

## Fit model
```{r}

nsims = 4000
warmup = 3000
fit <- stan(
  file = 'AR1plusarima.stan',           # Stan model file
  data = stan_data,                  # Data list for Stan
  iter = nsims,                       # Number of iterations
  warmup = warmup,
  chains = 4,                        # Number of chains
  seed = 123                         # Seed for reproducibility
) 
samples <- extract(fit)

dim(samples$y)
ycolm = colMeans(samples$y)
ycolq = colQuantiles(samples$y, probs=c(0.025,0.975))
Xmucolm = colMeans(samples$X_mu)
Xmucolq = colQuantiles(samples$X_mu, probs=c(0.025,0.975))

{
  ggd1 = data.frame(x=depth,mean=Xmucolm, lower=Xmucolq[,1], upper=Xmucolq[,2], obs = dy, truetrend = truetrend)
  ggp1 = ggplot(data=ggd1, aes(x=x)) + theme_bw() + xlab("Depth") + ylab("Layers per unit of depth")+
    ggtitle("(a) Fitted layer increments") +
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    geom_point(aes(y=obs),col="black")+
    #geom_line(aes(y=obs),col="gray")+
    geom_line(aes(y=truetrend),col="blue")+
    geom_line(aes(y=mean))
}

{
  cumsumXmu = matrix(NA,ncol=ncol(samples$X_mu),nrow=nrow(samples$X_mu))
  for(i in 1:nrow(samples$X_mu)){
    cumsumXmu[i,] = cumsum(samples$X_mu[i,])
  }
  cumxm = colSums(cumsumXmu)
  cumxquant = colQuantiles(cumsumXmu, probs=c(0.025,0.975))

  #plot(y_noise)
  #y_full = y_truth+bias
  y_og = cumsum(dy)
  #ggd2 = data.frame(x=depth,mean=ycolm-y_og,lower=ycolq[,1]-y_og,upper=ycolq[,2]-y_og, bias=bias)
  ggd2 = data.frame(x=depth,mean=ycolm,lower=ycolq[,1],upper=ycolq[,2], bias=bias+cumsum(dy),
                    xmean = cumxm, xlower=cumxquant[,1],xupper=cumxquant[,2])
  ggp2 = ggplot(data=ggd2,aes(x=x)) + theme_bw()+xlab("Depth")+ylab("Estimated age") +
    ggtitle("(b) Synchronized chronologies") +
    geom_ribbon(aes(ymin=xlower,ymax=xupper),fill="black", alpha=0.2)+
    geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3)+
    geom_line(aes(y=bias),col="blue")+
    geom_line(aes(y=mean))+
    #geom_point(data=data.frame(x=depth[which_obs],y=y_obs-y_og[which_obs]),aes(x=x,y=y),col="black")
    geom_point(data=data.frame(x=depth[which_obs],y=y_obs),aes(x=x,y=y),col="black")

  print(ggp2)
}
library(ggpubr)
ggboth1 = ggarrange(ggp1,ggp2,nrow=2)
#ggboth2 = ggarrange(ggp1,ggp3,nrow=1)

#ggpall = ggarrange(ggp11,ggp222,nrow=2)
ggsave(paste0("MCMCexample-3000x2000.eps"),ggboth1,device=cairo_ps, width=3000, height=2200, units="px")

#ggsave("MCMCexample2-4800x2000.eps", device=cairo_ps,plot=ggboth2,width=4800,height=2000, units="px",dpi=500,limitsize=FALSE)

traceplot(fit)
```

### 2025-01-17 inst/reproduce_results_BA/results.Rmd
Now supports systematic biases
Systematic biases can now be described using an RW2 process. Also added other stuff, including code to reproduce the results of the most recent paper

# Code for Myrvoll-Nilsen et al.

This file includes the data and code needed to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Synchronization of layer-counted paleoclimatic proxy archives.

## Illustration example

The data used to generate the illustrative examples in the paper is as follows.

```{r illustrationexample}
# rm(list=ls())
require(stats)
library(ggpubr)
library(ggplot2)
set.seed(123)
n=800
y0 = 0
z0 = 0
phi=0.8; sigma=15
depth = z0+seq(from=1,to=n,length.out=n)

covariate = arima.sim(model=list(ar=c(0.99)),n=n)
noise = sigma*arima.sim(model=list(ar=c(phi)),n=n,sd=sqrt(1-phi^2))

dy = 10 + depth*5/n+covariate+noise
y = y0+cumsum(dy)

data=data.frame(age=c(y0,y),depth=c(z0,depth),dy=c(0,as.numeric(dy)),covariate=c(0,covariate))
#data=rbind(c(y0,z0,0,0),df)
formula=dy~ 1 + depth + covariate 
nsims = 10000

# Unsynchronized
res = bremla(formula,data,reference.label="simulated time scale",nsims=nsims,
             control.fit=list(noise="ar1"),
             x.label="Depth",
             y.label="Ensemble age - observed age",
             control.sim = list(synchronized=FALSE,
                                nsims=nsims))

library(ggplot2)
orichron = res$data$age
ggd1 = data.frame(depth=res$data$depth,
                 mean = res$simulation$summary$mean-orichron,
                 lower = res$simulation$summary$lower-orichron,
                 upper = res$simulation$summary$upper-orichron)
ggd1_2 = data.frame(depth=res$data$depth,
                 mean = res$simulation$summary$mean,
                 lower = res$simulation$summary$lower,
                 upper = res$simulation$summary$upper)
ggpunc = ggplot(data=ggd1,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,linewidth=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)
ggpunc2 = ggplot(data=ggd1_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,linewidth=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)
# plot(res)
#ggsave("illustrative-unsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

# Fixed tie-points
ntie = 3
tiesims = matrix(NA,nrow=nsims,ncol=ntie)
tiedepths = c(100,400,600)
#tieage = c(450,3000,8500)
tieage = c(450,2000,10000)

for(i in 1:ntie){
  tiesims[,i] = rep(tieage[i],nsims)
}

res2 = tiepointsimmer(res,synchronization = list(locations=tiedepths,locations_unit="depth",
                       samples = tiesims)
               )

res2 = bremla_synchronized_simulation(res2,control.sim=list(synchronized=TRUE))

ggdfix = data.frame(depth=res2$data$depth,
                 mean = res2$simulation$summary_sync$mean-orichron,
                 lower = res2$simulation$summary_sync$lower-orichron,
                 upper = res2$simulation$summary_sync$upper-orichron)
ggd_2 = data.frame(depth=res2$data$depth,
                 mean = res2$simulation$summary_sync$mean,
                 lower = res2$simulation$summary_sync$lower,
                 upper = res2$simulation$summary_sync$upper)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res2$tie_points$locations,
                  mean=res2$simulation$summary_sync$mean[tiedepths]-oritie)
ggd2_2 = data.frame(depth=res2$tie_points$locations,
                  mean=res2$simulation$summary_sync$mean[tiedepths])

ggpfix = ggplot(data=ggdfix,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.6)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_point(data=ggd2,mapping=aes(x=depth,y=mean),col="black",size=1)
ggpfix2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.6)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_point(data=ggd2_2,mapping=aes(x=depth,y=mean),col="black",size=1)

#plot(res)
#ggsave("illustrative-fixedsync.eps",device=cairo_ps,width=4800,height=2000,units="px")

#tiemeans = c(450,3000,8500)
tiemeans = c(450,2000,10000)
tiesds = c(400,500,800)
res3 = tiepointsimmer(res,synchronization = list(locations=tiedepths,
                                                locations_unit="depth",
                                                method="normal",
                                                params=list(mean=tiemeans,
                                                            sd=tiesds))
               )
res3 = bremla_synchronized_simulation(res3,control.sim=list(synchronized=TRUE))

ggd = data.frame(depth=res3$data$depth,
                 mean = res3$simulation$summary_sync$mean-orichron,
                 lower = res3$simulation$summary_sync$lower-orichron,
                 upper = res3$simulation$summary_sync$upper-orichron)
ggd_2 = data.frame(depth=res3$data$depth,
                 mean = res3$simulation$summary_sync$mean,
                 lower = res3$simulation$summary_sync$lower,
                 upper = res3$simulation$summary_sync$upper)

oritie = orichron[tiedepths]
ggd2 = data.frame(depth=res3$tie_points$locations,
                  lower=res3$simulation$summary_sync$lower[tiedepths]-oritie,
                  upper=res3$simulation$summary_sync$upper[tiedepths]-oritie)
ggd2_2 = data.frame(depth=res3$tie_points$locations,
                  lower=res3$simulation$summary_sync$lower[tiedepths],
                  upper=res3$simulation$summary_sync$upper[tiedepths])

ggprand = ggplot(data=ggd,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(data=ggd,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_segment(data=ggd2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
               col="black",size=0.7)
ggprand2 = ggplot(data=ggd_2,aes(depth))+theme_bw()+xlab("Depth")+ylab("Ensemble age - observed age")+
  geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower, ymax=upper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=lower,ymax=upper),col="red",fill="red",alpha=0.3,size=0.7)+
  geom_line(aes(y=mean),col="blue",size=0.7)+
  geom_segment(data=ggd2_2,mapping=aes(x=depth,xend=depth,y=lower,yend=upper), 
               col="black",size=0.7)

#plot(res3)
ylim = range(res$simulation$summary$lower-orichron,res$simulation$summary$upper-orichron,
             res3$simulation$summary_sync$lower-orichron,res3$simulation$summary_sync$upper-orichron)
ggpunc = ggpunc + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")
ggpfix = ggpfix + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")
ggprand = ggprand + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")
ggplotlist0 = c(); ggplotlist0[[1]] = ggpunc; ggplotlist0[[2]] = ggpfix; ggplotlist0[[3]] = ggprand
ggplotlist00 = ggarrange(plotlist=ggplotlist0,nrow=3,ncol=1)

ylim = range(res$simulation$summary$lower,res$simulation$summary$upper,
             res3$simulation$summary_sync$lower,res3$simulation$summary_sync$upper)
ggpunc2 = ggpunc2 + ylim(ylim) + ggtitle("(a) Unsynchronous chronologies")+ylab("Ensemble age")
ggpfix2 = ggpfix2 + ylim(ylim)+ ggtitle("(b) Synchronous chronologies: fixed tie-points")+ylab("Ensemble age")
ggprand2 = ggprand2 + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")+ylab("Ensemble age")
ggplotlist1 = c(); ggplotlist1[[1]] = ggpunc2; ggplotlist1[[2]] = ggpfix2; ggplotlist1[[3]] = ggprand2
ggplotarrange = ggarrange(plotlist=ggplotlist1,nrow=3,ncol=1)
plot(ggplotarrange)

# ggsave("illustrative-all-3000x3400.eps",device=cairo_ps,width=3000,height=3400,units="px",dpi=500,limitsize=FALSE)

#ggprand2 = ggprand2 + ylim(ylim)+ ggtitle("(c) Synchronous chronologies: uncertain tie-points")+ylab("Ensemble age")
# 
# ggprand22 =  ggprand2 + geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower,ymax=upper),fill="black",alpha=0.2)
# ggpfix22 =  ggpfix2 + geom_ribbon(data=ggd1_2,mapping=aes(x=depth, ymin=lower,ymax=upper),fill="black",alpha=0.2)

ggplotlist1 = c(); ggplotlist1[[1]] = ggpunc2; ggplotlist1[[2]] = ggpfix2; ggplotlist1[[3]] = ggprand2
ggplotlist2 = c(); ggplotlist2[[1]] = ggpunc2; ggplotlist2[[2]] = ggpfix22; ggplotlist2[[3]] = ggprand22
ggplotarrange = ggarrange(plotlist=ggplotlist1,nrow=3,ncol=1)
plot(ggplotarrange)
ggplotarrange2 = ggarrange(plotlist=ggplotlist2,nrow=3,ncol=1)

print(ggplotarrange2)

#ggsave("illustrative-all2-3000x3400.eps",device=cairo_ps,width=3000,height=3400,units="px",dpi=500,limitsize=FALSE)

```

# Simulation example

```{r}
require(stats)
library(ggplot2)
set.seed(123)
n=1000
y0 = 0
z0 = 0
phi=0.8; sigma=3

depth = z0+seq(from=1,1000,length.out=1000)
depth2 = depth^2#/z0 #normalize for stability

dnoise = sigma*arima.sim(n=n,model=list(ar=c(phi)),sd=sqrt(1-phi^2))
proxy = arima.sim(n=n,model=list(ar=0.9),sd=sqrt(1-0.9^2))

truevals = c(sigma,phi,       0.5, 10, 0.01, 20, 0.02, 0, 0.03)
namevals = c("\\sigma","\\phi","b_w","a_1","b_1","a_2","b_2","a_3","b_3")

events=list(locations=c(1,250,750,1000))
a1 = c(rep(1,249),numeric(751))
a2 = c(numeric(249),rep(1,500),numeric(251))
a3 = c(numeric(749),rep(1,251))
c1 = c(depth[1:249],numeric(751))
c2 = c(numeric(249),depth[250:749],numeric(251))
c3 = c(numeric(749),depth[750:1000])
dy = dnoise + a1*truevals[4]+a2*truevals[6]+a3*truevals[8]+c1*truevals[5]+
  c2*truevals[7]+c3*truevals[9] + proxy*truevals[3]
plot(dy)
age = y0+cumsum(dy)
df=data.frame(age=age,dy=as.numeric(dy),proxy=as.numeric(proxy),depth=depth); data=rbind(c(y0,z0,NA,NA),df)
formula=dy~-1+proxy#+depth2
nsims = 1000

control.sim = list(ncores=1, nsims=nsims, synchronized=TRUE)
# Specify tie-point distribution
control.fit=list(method="INLA", noise="ar1")
synchronization = list(locations = c(1, 200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(0, 2200,6800,19000),sd=c(0, 10,50,60)), 
                       agedisc=list(model="rw2",
                                    options=list(restart.fromlast=FALSE,
                                                 inla.options=list(num.threads=1,
                                                                   verbose=TRUE,
                                                                   control.compute=list(openmp.strategy="huge")))))

synchronization = list(locations = c(1, 200,400,800),locations_unit="index",method="gauss", nsims=nsims,params=list(mean=c(0, 2200,6800,19000),sd=c(0, 10,50,60)), 
                       agedisc=list(model="rw2"))

```

```{r}

## set up model
res = bremla_prepare(formula=formula, data=data, nsims=nsims, events=events,
               control.fit=control.fit,
               control.sim=control.sim,
               synchronization=synchronization)

## fit model
res = bremla_modelfitter(res)

## sample unsynchronized chronologies

res = bremla_chronology_simulation(res)

## generate tie-points

res = tiepointsimmer(res)

## sample synchronized chronologies
res = bremla_synchronized_simulation(res, print.progress=TRUE)

## To do all these steps in one call use:
# res = bremla(formula=formula, data=data, nsims=nsims, events=events,
#                control.fit=control.fit,
#                control.sim=control.sim,
#                synchronization=synchronization,
#                print.progress=TRUE)

```

# Real data example: NGRIP/GICC05

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). 
We also use tie-points described by probability distributions obtained from Adolphi et al. (2018) and Muscheler et al. (2020). Both the NGRIP/GICC05 data and the Adolphi/Muscheler tie-points are included in the package and can be imported as follows.

```{r}
library(INLA)
set.seed(1991)
require(stringr)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
d18O = NGRIP_5cm$d18O
proxy=d18O
formula = dy~-1+depth2
depth2 = depth^2/depth[1]^2 #normalize for stability

# set up data set with n+1 rows. Must include 'age', 'depth' and all covariates in 'formula'. In the first row, only 'y0' and 'z0' are collected:
data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth2,proxy=proxy)

nsims=1000 #number of chronologies to be sampled

# specify the events which separates piecewise predictor in layer increment model. See '?events.default' for more details:
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

# Simulation options. Specify 'synchronization=TRUE' so it doesnt only create unsynchronized samples. See '?control.sim.default' for more details
control.sim=list(synchronized = TRUE)

# Fit options. These specify options for the model and fitting procedure in the layer-increment model. We use AR(1) noise and the INLA method. See '?control.fit.default' for more details
control.fit=list(method="inla", noise="ar1")

# Options for synchronization approach. Here, tie-point samples, or model specifications must be included. locations should also be included. To use the Adolphi tie-points, specify 'method="adolphi"' and use 'locations' to specify which tie-points to be included as demonstrated below:
synchronization=list(method="adolphi", #Choose which Adolphi tie-point to use (the first one takes place after the Holocene and is therefore omitted here)
                     locations=c(FALSE,TRUE,TRUE,TRUE,TRUE)
                     )

```

## Fit models with bremla

The bremla package will set up the bremla object with all options specified in 'events', 'control.fit', 'control.sim' and 'synchronization', using the 'bremla_prepare' function. It also contains functions to perform least squares to find good initial values for an INLA fit, and simulate 
It will also set up the bremla object with all

### Assuming no systematic biases
If we assume there are no systematic biases, then the tie-points can be considered as realizations of the cumulative of the inferred layer increment model. If this model is Gaussian we can do this very efficiently

```{r}

 object0 = bremla(formula=formula, data=data, nsims=nsims, events=events,
                control.fit=control.fit,
                control.sim=control.sim,
                synchronization=synchronization)

```

#### Plot results

```{r}

ggdreal0 = data.frame(depth=object0$data$depth, age=object0$data$age, agemean= object0$simulation$summary$mean, agelower = object0$simulation$summary$lower, ageupper=object0$simulation$summary$upper, sagemean=object0$simulation$summary_sync$mean, 
                 sagelower=object0$simulation$summary_sync$lower, sageupper=object0$simulation$summary_sync$upper,
                 agediffmean=object0$simulation$summary$mean-object0$data$age,
                 agedifflower=object0$simulation$summary$lower-object0$data$age, 
                 agediffupper=object0$simulation$summary$upper-object0$data$age,
                 sagediffmean=object0$simulation$summary_sync$mean-object0$data$age,
                 sagedifflower=object0$simulation$summary_sync$lower-object0$data$age,
                 sagediffupper=object0$simulation$summary_sync$upper-object0$data$age)

#library(matrixStats)
tiequants = colQuantiles(object$tie_points$samples, probs=c(0.025,0.975))
ggdtiereal = data.frame(locations=object$data$depth[object$tie_points$locations_indexes],
                        lower = tiequants[,1], upper = tiequants[,2],
                        dlower = tiequants[,1] - object$data$age[object$tie_points$locations_indexes], dupper = tiequants[,2] - object$data$age[object$tie_points$locations_indexes])

library(ggplot2)

ggpreal0 = ggplot(data=ggdreal0,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agedifflower, ymax=agediffupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagedifflower,ymax=sagediffupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagediffmean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=dlower,ymax=dupper),color="black")+
  xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")

```

### Using RW2 bias
Here we assume there could be systematic biases, and model the resulting discrepancy between the number of inferred layers and the true age, as represented by the tie-points, using a 2nd order random walk model

```{r}

## set up model

synchronization$agedisc = list(model="rw2") #choose the 2nd order random walk for the age discrepancy model

object = bremla_prepare(formula=formula, data=data, nsims=nsims, events=events,
               control.fit=control.fit,
               control.sim=control.sim,
               synchronization=synchronization)

## fit model
object = bremla_modelfitter(object)

## sample unsynchronized chronologies

object = bremla_chronology_simulation(object)

## generate tie-points

object = tiepointsimmer(object)

## sample synchronized chronologies
object = bremla_synchronized_simulation(object, print.progress=TRUE)

## To do all these steps at the same time use
# object = bremla(formula=formula, data=data, nsims=nsims, events=events,
#                control.fit=control.fit,
#                control.sim=control.sim,
#                synchronization=synchronization,
#                print.progress=TRUE)

```

#### Plot results

```{r}
ggdreal = data.frame(depth=object$data$depth, age=object$data$age, agemean= object$simulation$summary$mean, agelower = object$simulation$summary$lower, ageupper=object$simulation$summary$upper, sagemean=object$simulation$summary_sync$mean, 
                 sagelower=object$simulation$summary_sync$lower, sageupper=object$simulation$summary_sync$upper,
                 agediffmean=object$simulation$summary$mean-object$data$age,
                 agedifflower=object$simulation$summary$lower-object$data$age, 
                 agediffupper=object$simulation$summary$upper-object$data$age,
                 sagediffmean=object$simulation$summary_sync$mean-object$data$age,
                 sagedifflower=object$simulation$summary_sync$lower-object$data$age,
                 sagediffupper=object$simulation$summary_sync$upper-object$data$age)

#library(matrixStats)
tiequants = colQuantiles(object$tie_points$samples, probs=c(0.025,0.975))
ggdtiereal = data.frame(locations=object$data$depth[object$tie_points$locations_indexes],
                        lower = tiequants[,1], upper = tiequants[,2],
                        dlower = tiequants[,1] - object$data$age[object$tie_points$locations_indexes], dupper = tiequants[,2] - object$data$age[object$tie_points$locations_indexes])

library(ggplot2)
ggpreal1 = ggplot(data=ggdreal,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agelower, ymax=ageupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagelower,ymax=sageupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagemean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=lower,ymax=upper),color="black")+
  xlab("Depth (m)") + ylab("Age (yr b2k)")+
  ggtitle("(a) Absolute time scale")

ggpreal2 = ggplot(data=ggdreal,aes(x=depth)) + theme_bw()+
  geom_ribbon(aes(ymin=agedifflower, ymax=agediffupper), fill="black", alpha=0.2)+
  geom_ribbon(aes(ymin=sagedifflower,ymax=sagediffupper), color="red", fill="red", alpha=0.3)+
  geom_line(aes(y=sagediffmean),color="blue") +
  geom_linerange(data=ggdtiereal,aes(x=locations,ymin=dlower,ymax=dupper),color="black")+
  xlab("Depth (m)") + ylab("Age deviation (years)")+
  ggtitle("(b) Age deviation from GICC05")

```

### 2025-01-17 inst/reproduce_results_CP/myrvollnilsen2022.Rmd
Now supports systematic biases
Systematic biases can now be described using an RW2 process. Also added other stuff, including code to reproduce the results of the most recent paper

--- 2022-12-06 inst/Examples/Paper_results/Paper_results_1.Rmd
@@
 
-<!-- Paper_results.md is generated from paper_results.Rmd. Please edit that file -->
+<!-- myrvollnilsen2022.md is generated from myrvollnilsen2022.Rmd. Please edit that file -->
 
@@
   comment = "#>",
-  fig.path = "man/figures/results-",
+  fig.path = "man/figures/reproduce/myrvollnilsen2022/results-",
   dev = "png",
-  out.width = "80%"
+  out.width = "100%"
 )
@@
                 control.fit=list(noise="ar1"),
-                control.sim=list(synchronized=FALSE) ,
-                print.progress=TRUE
+                control.sim=list(synchronized=FALSE) 
                 )
@@
                                   biasparams = matrix(c(0.98,1.02,0.96,1.04),nrow=2)
-                                  ),
-                print.progress=TRUE
+                                  )
                 )
@@
 
+  #rescale logdust proxy to avoid having to retune priors again
+  rescale.y.factor = -20
   control.linramp_dust = list(proxy=logdust,interval=interval_dust,interval.unit="index",
@@
                          priorparams=priorparams,
-                         rescale.y.factor=-20,
+                         rescale.y.factor=rescale.y.factor,
                          imp.fit=TRUE,
@@
 rownames(ageresults)=rownames
+
+cat("Onset depth estimates:\n",sep="")
+print(depthresults)
+cat("Onset age estimates:\n",sep="")
+print(ageresults)
 ```

### 2025-01-17 R/bremla_prepare.R

@@ -96,6 +96,7 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
     control.fit=list(noise="ar1")
 
   }
+
   control.fit = set.options(control.fit,control.fit.default())
 
   if(!is.null(control.sim)) control.sim = set.options(control.sim,control.sim.default())
@@ -216,6 +217,7 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
   if(!is.null(control.fit)){ #add random effect to formula string for use in INLA
     if(tolower(control.fit$method) %in% "inla" ){
       if(!is.null(control.fit$rgeneric)){ #rgeneric model is specified
+
         control.fit$noise = "rgeneric"
 
         ntheta = length(control.fit$rgeneric$from.theta)
@@ -231,6 +233,7 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
 
 
       }else{
+
         hyperprior = control.fit$hyperprior
         if(!is.null(hyperprior)){
 
@@ -278,16 +281,19 @@ bremla_prepare = function(formula,data,nsims=NULL,reference.label=NULL,
     formula_inla=as.formula(formulastring)
   }
 
+
   object= list(data=data_obj,formula=formula_inla,
                initials=initials,
                original.chron = data.frame(depth=depth,age=age))
   if(!is.null(events)) object$events=events
   #store input arguments
   str = cleanstring(format(formulastring))
+
   tildepos = str_locate(str,"~")[1]
   responsename = str_sub(str,start=1L,end=tildepos-1)
 
   lat.selection = lat.selector(format(formulastring))
+
   object$.internal=list(formula.ls = as.formula(formulastring),
                         lat.selection=lat.selection)
   object$.args=list(formula.ls=as.formula(formulastring),

### 2025-01-17 R/bremla_synchronized_simulation.R

@@ -63,6 +63,7 @@
 #' @keywords simulation synchronization tiepoint
 #'
 #' @export
+#' @importFrom utils write.table
 bremla_synchronized_simulation = function(object,control.sim,print.progress=FALSE){
 
 time.start = Sys.time()
@@ -70,7 +71,7 @@ time.start = Sys.time()
 
     if(!is.null(object$.args$control.sim)){
       if(print.progress){
-        cat("'control.sim' missing. Importing information from 'object'.",sep="")
+        # cat("'control.sim' missing. Importing information from 'object'.",sep="")
       }
       control.sim = object$.args$control.sim
     }else{
@@ -127,122 +128,395 @@ time.start = Sys.time()
 
   if(tolower(object$.args$control.fit$method) == "inla"){ # && is.null(object$simulation)){ #if INLA is used
 
-    reg.model = object$.internal$lat.selection
-    latentselection = object$.internal$lat.selection
 
-
-    timecoef.start = Sys.time()
-    ncores_postsamp = max(1,object$.args$control.sim$ncores)
-    latentsamples = INLA::inla.posterior.sample(nsims,object$fitting$inla$fit,
-                                                selection=latentselection,verbose=FALSE,
-                                                add.names=FALSE,num.threads = ncores_postsamp)
-    #int_hyper = latentsamples[[i]]
-
-    #int_hyper = data.frame(matrix(NA, nrow=nsims,ncol=length(latentsamples[[1]]$hyperpar))) #get hyperparameters from joint distribution
-
-    #colnames(int_hyper) = names(latentsamples[[1]]$hyperpar)
-    timecoef.end = Sys.time()
-    if(print.progress) cat(" completed in ",difftime(timecoef.end,timecoef.start,units="secs")[[1]]," seconds.\n",sep="")
-    samples = matrix(NA,nrow=n,ncol=nsims)
+    ## Collect unsynchronized samples
 
 
 
+    if(print.progress){
+      cat("Preparing synchronization...\n",sep="")
+    }
+    if(is.null(object$.args$synchronization$agedisc)){
 
-  if(print.progress) cat("Sampling fixed coefficients...",sep="")
+      ## Old
+      if(print.progress){
+        cat("Assuming no structural biases...\n",sep="")
+      }
+
+      reg.model = object$.internal$lat.selection
+      latentselection = object$.internal$lat.selection
 
 
[…]

### 2025-01-17 R/default_arguments.R

@@ -2,7 +2,7 @@
 #'
 #' Sets the default variables in the list \code{events} used to specify the climatic
 #' periods and separating events used in the linear predictor. The list contains the following arguments:
-#' \itemize{
+#' \describe{
 #'   \item{\code{locations} }{Numeric describing the location of the transitions between climatic periods.
 #'    Default value is \code{NULL}, and must be provided by user.}
 #'   \item{\code{locations_unit} }{Character describing on which axis the locations are given. Can be \code{age},
@@ -38,43 +38,21 @@ events.default <- function(){
 #'
 #' Sets the default variables in the list \code{control.fit} used to specify the
 #' fitting procedure. The list contains the following arguments:
-#' \itemize{
-#'   \item{\code{noise} }{Character which noise process is assumed for the residuals.
-#'   \code{"iid"}, \code{"ar1"} and \code{"ar2"} are currently. However, if synchronization is to
-#'   be performed only \code{"ar1"} is currently supported. Default input is \code{"ar1"}.}
-#'   \item{\code{method} }{Character describing which method should be used to fit the process.
-#'   Only \code{inla} is currently supported, but more will hopefully come in the future.}
-#'   \item{\code{verbose} }{Boolean telling whether or not the \code{inla}-program should run in
-#'   verbose mode. Default \code{FALSE}.}
-#'  \item{\code{hyperprior}}{List corresponding to the \code{hyper} object in INLA's \code{f}
-#'   function which specifies the priors of the hyperparameters. Default is \code{NULL}
-#'   which sets INLA's default priors.}
-#'   \item{\code{improve.fixed} }{Boolean indicating whether or not the least squares estimate of
-#'   the fixed values should be used as initial values for the optimization procedure in INLA.
-#'   This feature has not been extensively tested so it might be unstable. Default \code{FALSE}}
-#'   \item{\code{ncores} }{Integer used to set the number of cores to be used in the \code{inla} fitting
-#'   procedure. Default value is \code{1}.}
-#'   \item{\code{transform} }{Character indicating which transformation should be applied to
-#'   the observations. Can be \code{"log"} or \code{"identity"}. If synchronization is
-#'   intended then only \code{"identity"} is currently supported, which is also the
-#'   default value here.}
-#'   \item{\code{rgeneric}} List containing necessary arguments for specifying a custom
-#'   model for the noise component of the layer increments using the \code{rgeneric}
-#'   framework of \code{INLA}. Note that this alternative generally performs less well
[…]

### 2025-01-17 R/rgeneric.test.R

@@ -1,0 +1,96 @@
+
+rgeneric.test = function( #specifies necessary functions for INLA to define the linear ramp model
+  cmd = c("graph", "Q","mu", "initial", "log.norm.const", "log.prior", "quit"),
+  theta = NULL)
+{
+
+  #envir = environment(sys.call()[[1]])
+  envir = parent.env(environment())
+  #TAU = exp(12)
+
+  interpret.theta = function() { #helpful function to transform back from internal parametrization
+    if(!is.null(envir)){
+      dMCEE=get("dMCE",envir)
+    }
+
+    kappa = exp(theta[1])
+    sigma = 1/sqrt(kappa)
+    sigmas = sigma*dMCEE/2 #+1/TAU
+    kappas = 1/sigmas^2
+cat("sigma:",sigma,"\n")
+cat("rangesigmas:",range(sigmas),"\n")
+    return(list(sigma=sigma,sigmas=sigmas,kappa=kappa,kappas=kappas))
+  }
+
+  mu = function() { #mean vector defined as a linear ramp
+
+    return(numeric(0))
+  }
+
+
+  graph = function(){ #graphs of conditional dependence structure. 1 where Q[i,j] != 0, 0 where Q[i,j] = 0
+    G = Q()
+    G[G != 0] = 1
+    return (G)
+  }
+
+  Q = function(){ #inverse covariance matrix: AR(1) that allow for unequal spacing
+    if(!is.null(envir)){
+      #n=get("n",envir)
+      #timepoints=get("timepoints",envir)
+    }
+    param=interpret.theta()
+
+
+    kappas = param$kappas
+    n=length(kappas)
+    ii = c(1:(n-1), n, 1:(n-1))
+    jj = c(1:(n-1), n, 2:n)
+
+    xx = c(kappas[1:(n-1)]+kappas[2:n], kappas[n], -kappas[2:n])
+
+    Q = Matrix::sparseMatrix(
+      i = ii,
+      j = jj,
+      x = xx,
+      symmetric = TRUE
+    )
+    cat("Qrange:",range(Q),"\n")
+    cat("Qdet:",det(Q),"\n")
+    #diag(Q)=diag(Q)
+    return (Q)
+  }
+
+  log.norm.const = function(){ #INLA computes this automatically
+    return(numeric(0))
+  }
+
+  log.prior = function(){
+    if(!is.null(envir)){
+      #tslutt=get("tslutt",envir)
[…]

### 2025-01-20 README.Rmd
Small changes to text

@@
 
-This repository contains the bremla package (**B**ayesian **Re**gression **M**odeling of **L**ayer-counted **a**rchives) for uncertainty quantification and synchronization for layer-counted proxy archives, including the data and code to reproduce the results for
+This repository contains the bremla package (**B**ayesian **Re**gression **M**odeling of **L**ayer-counted **a**rchives) for uncertainty quantification and synchronization for layer-counted proxy archives.
+This repository also includes, in the 'inst/' folder, the data and code to reproduce the results for
 
@@
 
-Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Tie-point paper (title TBD)
+Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (2025). Synchronization of layer-counted paleoclimatic proxy archives using a Bayesian regression modeling framework. Bayesian Analysis. In review.
 

### 2025-01-20 inst/reproduce_results_BA/results.Rmd
Small changes to text

@@
 #### Plot results
+This code generates the plots included in the paper.
+The important results can be illustrated simply by using 'plot(object0)', but this code also includes the shaded area representing the unsynchronized chronology ensemble.
 
@@
 
+#plot(object0)
 ggdreal0 = data.frame(depth=object0$data$depth, age=object0$data$age, agemean= object0$simulation$summary$mean, agelower = object0$simulation$summary$lower, ageupper=object0$simulation$summary$upper, sagemean=object0$simulation$summary_sync$mean,
@@
 ```{r}
+
+#plot(object)
 ggdreal = data.frame(depth=object$data$depth, age=object$data$age, agemean= object$simulation$summary$mean, agelower = object$simulation$summary$lower, ageupper=object$simulation$summary$upper, sagemean=object$simulation$summary_sync$mean,
